\documentclass[11pt]{article}
\usepackage{html, epsf}
\begin{document}
\title{\bf AIPS++ Quarterly Report: 1998 Q2}
\author{T.J. Cornwell, NRAO}
\date{1998 July 17}
\maketitle

\tableofcontents

\section{Introduction}

In this report, we cover the personnel changes and progress in AIPS++
over the last quarter (1998 Q2). We also discuss the issue of ``Thin
Path'' development raised by the STAG and mentioned in the previous
quarterly report.

\section{Personnel changes}

\begin{itemize}

\item Dan Briggs of the NCSA/BIMA AIPS++ group was tragically killed
in a sky-diving accident on July 5, 1998. Dan joined AIPS++ last
year to work on synthesis processing and parallelization. Many of us
knew Dan closely and valued him as an excellent friend and colleague.
His death is a sore loss for both AIPS++ and the radio-interferometry
community. A memorial page for Dan can be found at:

\htmladdnormallink{{\em http://www.nmt.edu/\~{}pio/dan/dan-play.html}}
{http://www.nmt.edu/\~{}pio/dan/dan-play.html}.

\item Brian Glendenning has been chosen to be Head of Computing for
the NRAO Millimeter Array project. This is a honor for Brian and
an opportunity not be be missed, but we will miss his presence in
AIPS++. Brian has been a member of the project during it's entire
lifetime. He is the key architect of the system and has contributed
ideas and code to nearly all parts of the system. He is our best
pinch-hitter, able to work productively in all areas. We wish him
well in his new position, and note that it augurs well for the
cooperation between AIPS++ and the MMA. Brian will remain located 
in Socorro and will be available for occasional consultation.

\item Athol Kemball will replace Brian as Deputy Project Manager,
effective July 1. Ger van Diepen will be Acting Technical Leader
until the end of 1998, at which point the question of the requirements 
for the Technical Leader will be revisited.

\item Mark Holdaway will join the AIPS++ Project on July 13, filling
the position vacated by Brian Glendenning. Mark will be located in
Tucson, thus providing a local contact for 12m users. Mark has
previously worked within AIPS++ and is thus already familiar with both
the basic AIPS++ library and programming in C++. In addition, Mark has
great expertise in many areas of interferometry, including mosaicing,
and the design of the MMA. His first task will be to work on mosaicing
applications.

\item Jeff Uphoff is leaving the NRAO AIPS++ group in Charlottesville
to take a position with TransMeta Corp. Joe McMullin will move to
Charlottesville to replace Jeff, and NRAO will hire a replacement
for Jeff in Green Bank.

\item John Pixton will leave the NCSA/BIMA AIPS++ group to take a
position in Norway.

\item David Barnes joined the AIPS++ Project in Epping, and
Peter Barnes (no relation) joined the AIPS++ Project in Socorro.
\end{itemize}

As a result of all of these changes, the next quarter will require
lot of work on hiring. In addition to replacing Dan Briggs and
Jeff Uphoff, we have to hire a post-doc to work on wide-field
imaging, and someone to work on parallel i/o for the parallelization
effort.

The loss of Dan Briggs will slow down our work on wide-field imaging,
and the departure of Jeff Uphoff will further freeze our work on
system support.

\section{Developments in 1998 Q2}

In {\bf Single Dish support}, we added the following to the Dish
program: an operation to save results back to disk, the ability to
save the state of Dish and its plotter between executions of the
program, and use of the same pgplotter widget in the dish plotter as
used elsewhere in Aips++.  The full SDFITS convention is now supported
in the fits2table application.  At Green Bank, we added support for
the Holography backend to the existing fillers.  The RFI data base
developed last quarter is now in use in Green Bank and we are
responding to the initial user feedback.  We support GBT staff in
on-going testing of GBT hardware and software.  We have begun
development of the continuum analysis tools necessary for GBT
commissioning.

In {\bf Synthesis support}, we finalized the definition of the second
version of the MeasurementSet format. This will be implemented following
the third beta release. We developed a glish-based visibility
plotter using the ms module and the pgplotter interactive plotter.
We expect that this will meet visibility examination needs for the
short term but will have to be revised or replaced when the display
library is available for application programmers. Only about a weeks
investment of time was needed for the development of a very functional
initial version. The {\tt sky} synthesis imaging application was subject
to testing by various testing and as a result a number of (mostly 
subtle) bugs were fixed. The corresponding {\tt cal} synthesis 
calibration application is still under development and has not
been placed in the system. Wim Brouw developed a very promising
approach to wide-field imaging in which some of the most time-consuming
operations (the image coordinate system conversions) are no longer
required. The Measures system now supports this via the UVWMachine
class, and we expect to implement the full algorithm in the next
quarter.

In {\bf Glish}, a number of important modifications were made:

\begin{itemize}

\item Basic file I/O was completed; this allows Glish users to access
        ASCII files line-by-line. The user can also get basic information
        about files (similar to what is returned by the Unix stat() function).

\item printf(), sprintf(), and fprintf() were added to Glish as built-in
        functions. This addition was complicated by the fact that stdarg
        argument lists cannot be created dynamically at runtime. As a
        result, the (publicly available) source code for the shell
        version of printf() had to be used.

\item A TCL/Tk based client was developed. This client can be used instead
        of the current version of the TCL/Tk widgets which is linked into
        the Glish interpreter. This client replaces the current widgets
        without any modification to existing scripts which the Tk widgets.
        This required additional classes built upon the Client class
        along with changes to the internals of the Glish interpreter.
        These changes make it easy for those developing Glish clients
        to have a client act as a repository for "proxy clients". These
        proxy clients are similar in functionality to the AIPS++ distributed
        objects (DOs).

        There were several reasons for making this change:

\begin{itemize}
\item The library (Rivet) which Glish used for the linked-in
                        version of the widgets is no longer supported.

\item Because the Rivet library didn't include the TCL
                        interpreter, access to all of the "mega-widgets"
                        is impossible.

\item Portability to other operating systems is limited
                        because the Rivet library is built upon out-of-date
                        versions of TCL and Tk.

\end{itemize}

This new client has been checked into the AIPS++ system, and after a
long period of testing, the Tk widgets will no longer be linked into
the Glish interpreter.

\item Many bugs/minor improvements were fixed/made:

\begin{itemize}
\item TCL/Tk bugs with popup frames and deleting frames
      within a canvas were fixed

\item TCL/Tk additions for raising frames, controlling
                        scales and disabled buttons, and setting title for
                        toplevel frames were made

\item Bugs with regular expressions, scope of missing(),
                        loss of shape attribute with conj() and complex(),
                        loss of non-shape attributes when slicing an
                        array, and problems with request/reply were fixed

\end{itemize}

\item Analysis of Glish's memory usage and speed continued, but no major
        improvements were made
\end{itemize}

In {\bf Measures}, after ample discussions about the proper
inheritance structure and format, we wrote (and got code-reviewed)
QuantumHolder and MeasureHolder classes, which are generic holders of
any Measure or Quantum, and aimed at providing a general interface to
the Tasking system. We included the VLBI reference source list in the
'data' system, which is accessable from both the Glish (measures gui)
and c++ code.  We wrote the MBaseline Measure (and its coordinating
classes): no Earth tide or other secondary effects included as yet,
Muvw Measure (and its coordinating classes), added an angular shift
method (in both true and projected angles) to the MDirection Measure,
added de-projection of an 'off-axis' field to UVWMachine, improved
some of MDirections forward/backward conversion precisions, finished
EarthMagnetic field Measure using the IGRF reference magnetic field,
wrote a FieldMachine to easily calculate the magnetic field at a
certain direction in e.g. the ionosphere, wrote a VelocityMachine to
convert between velocities and frequencies in a simple manner, and
added new ITRF and TOPO codes to MDirection.  We also wrote an article
for newsletter on measures gui.
 
In {\bf AIPS++ Infrastructure}, mostly minor changes were made, mostly
in response to bugs. The tablebrowser was improved significantly,
adding user-controlled formatting, smart formatting of special column
types such as time and direction, {\em etc.}. The catalog browser was
substantially re-written in response to comments and bug reports from
the alpha testers. Finally, we developed a prototype of a command-line
parameter-setting shell, similar in principle and details to various
systems now in use in astronomy (Standard commands package, AIPS,
MIRIAD).

In {\bf FITS support}, Harro Verkouter of JIVE continued his work on
re-implementation of some of the key FITS classes. This is needed
prior to the writing of an improved UVFITS reader and writer.

In {\bf Image Analysis}, work has largely revolved around regions of
interest, masks, the classes which hide the details of optimal
iteration through images, and the integration of these things with our
high level applications. Neil Killeen and Ger van Diepen made good
progress in these directions, and now have in place:

\begin{itemize} 

\item A wide range of C++ classes to handle lattice
coordinate based regions and masks (for example, boxes, polygons,
ellipsoids, extensions of these shapes along axes, unions of them,
differences, intersections and so on)

\item A much smaller set of C++ classes to handle world coordinate
base regions. So far we just have a box and polygon.  These class
objects know how to convert themeselves to the equivalent lattice
coordinate based objects whereupon the union, complement, difference
etc classes can be invoked 

\item The first steps towards improving the user interface from the
simple lattice coordinate blc/trc regions has been taken.  This was
done by creating Glish functions which create GlishRecords defining
the desired region definition.  The Glish record is then passed to the
image DO via an application where it is reconstituted into the C++
class object for use by the classes which access the lattice data in
the image.

\item The image analysis applications that we do have have been
integrated with these new classes as much as is possible.

\end{itemize}

In {\bf Display} work, David Barnes has joined the ATNF aips++ team
and reinvigorated the Display Library (DL) work following Tom
Oosterloo's departure last year; he is working closely with John
Pixton of NCSA. David spent time familiarizing himself with the DL 
design and current implementation.  Then, he

\begin{itemize}
\item worked through the existing high-level display library classes
   and identified capabilities that the library was lacking.
\item Constructed a class design to encompass the many different 
   possibilities for handling and display of astronomical data,
   based on the existing single-purpose class already provided
   in the DL.
\item Commenced implementing parts of the above design.  At the time
   of writing, the initial design is of order 30\% implemented.
\item Learnt how to build a Tcl/Tk widget to contain a DL 
   WorldCanvas.
\end{itemize}

In the {\bf System} area, the major change this quarter was the
addition of a system to "bless" development updates as having met
internally-defined criteria for stability and usability; this now
enables AIPS++ sites that require stable installations to confidently
track development progress between public releases.  Other system
changes include: continued incremental improvements to the code
management, distribution, compilation, testing, and documentation
systems; improved support for providing some of the third-party
software packages required by AIPS++ to end-users, particularly those
running Linux; and an investigation into building AIPS++ with the free
Lesstif library suite, attempting to eliminate its dependence on the
commercial Motif library suite.

Some future plans include: investigating replacing the current code
management and distribution systems with a free, "off-the-shelf"
alternative such as the widely-used CVS package; continued incremental
improvements to project "infrastructure" components; elimination of
AIPS++'s reliance on Motif; and simplified installation procedures for
AIPS++ on the major supported architectures.

In the area of {\bf Parallelization}, work proceeded on a number of
fronts. In April, Wes Young completed the parallelization of the spectral
line Clark CLEAN application in AIPS++.  Using the spectral line
deconvolution Doug Roberts has spent much of April, benchmarking the
task to measure the speedup across processors on the NCSA SGI Origin
2000.  The speedup achieved is encouraging; the application executed
22 times faster than uniprocessor speed on 32 processors.  An ideal
speedup on 32 processors would be 32.  The results of the benchmarking
were presented by Doug Roberts at the Alliance98 conference, held in
Urbana at the end of April.  Brian Glendenning and Ruth Milner from
NRAO along with Dan Briggs, Dick Crutcher, and Doug Roberts from NCSA
attended the Alliance'98 conference.  The on-line version of the
Alliance'98 presentation is available at:

\htmladdnormallink{{\em http://pacont.ncsa.uiuc.edu/A98/poster/roberts/}}
{http://pacont.ncsa.uiuc.edu/A98/poster/roberts/}.

Several potential
connections were initiated at the conference.  An NCSA team which has
developed tools to analyze I/O usage is interested in helping to
instrument our code.  NRAO is looking to hire someone for this in
Socorro who would collaborate closely with the NCSA team to identify
I/O problems.  The parallel-I/O component of the MPI-2 standard is
available now and will be investigated by the new hire to determine if 
it can help the I/O performance of AIPS++.

In {\bf Documentation}, we continued with publication of the
AIPS++ Newsletters:

\htmladdnormallink{{\em http://aips2.nrao.edu/aips++/docs/html/navpages/learnmore/newsletters.html}}{http://aips2.nrao.edu/aips++/docs/html/navpages/learnmore/newsletters.html}. 

Publication of a new edition of the newsletter is now announced via a
mailing-list (aips2news@nrao.edu).  We have installed a new system of
web pages that have improved layout and navigation aides. Our alpha
testers (particularly Rupen and Anantharamaiah) have started
re-writing the introductory documentation, with a view to making the
concepts in the system comprehensible to astronomers, and to defining
terminology to be used elsewhere in the documentation.

In {\bf Management}, Brian Glendenning left his position as Deputy
Project Manager to join the NRAO MMA project. Athol Kemball will
replace Brian as Deputy Project Manager, effective July 1. Ger van
Diepen will replace Brian as Acting Technical Leader, effective
September 1.  Tim Cornwell and Brian Glendenning gave presentations at
the NRAO Visiting and Users' Committee and Tim Cornwell gave a
presentation to the AUI Board of Trustees. We also performed an
analysis of the our success at keeping target dates. During the last
three years, we have identified and met 433 targets. The quartiles are
75\% are late by 1 week or less, 50\% are late by two weeks or less,
and 25\% are late by 5 weeks or more. The tail of late dates ranges up
to a few hundred days but these are mainly targets that have been
deferred following a change of plans. There is no systematic trend in
target date lateness with absolute time ({\em i.e.} we have not got
worse at meeting targets).  Since the average target takes 4-6 weeks
to finish, the fractional delay is about 33-50\%. This noise will
clearly make planning on longer time scales problematical.  The cause
of lateness is hard to identify but our belief is that much of the
delays are due to time being taken up by other, unscheduled activities
such as bug-fixing and making unforeseen changes to nominally
completed parts of the system. One might then regard the lateness as
the cost of complexity. As a result of this analysis, we are now
tracking time spent on- and off-target, with a view to improving our
overall performance in meeting targets, and thus easing planning
within the project. In the future, we will collect statistics on times
for fixing bug-reports, as this is more important as a quality measure
from the point of view of users.

\section{Beta Release}

Preparation of the third beta release continues. The pacing items now
are the release of the {\tt cal} synthsis calibration application, the
continuing testing of all modules, and the documentation being written
by our testers (particularly Michael Rupen and Anantharamaiah). The
planned documentation consists of a rewrite of the introductory
material and the establishment of astronomer-preferred
nomenclature. Retrofitting of the new terminology to the existing
extensive documentation will also be needed.

Our expectation for the third beta release is that the vast
improvements in functionality and interface over the previous (second)
release will attract comments from both new users and our existing
beta testers. We believe that continued tuning of the user interfaces
will be the most important issue arising from the release, and with
that in mind, we plan to re-allocate some fraction of Darrell
Schiebel's time to replace Brian Glendenning as our key worker on user
interfaces.

We expect that at least one more beta release will required following
this third release, putting the probable time for the first public
release some time early in 1999. 

\section{Thin path development in AIPS++}

This section discusses the question of so-called "thin-path" development
in AIPS++, and the implications for current priorities and resources.
This topic was raised by the STAG in the recent meeting, and discussed
extensively both within the project and with various advisory bodies
such as the NRAO Visiting and Users' Committees. Here we give our
final statement on our approach to thin-path development.

\subsection{Current AIPS++ synthesis development priorities}

At the current stage of development of the project, and for a
planning interval of 6-12 months, the following key synthesis
development priorities are identified:

\begin{description}
\item[Broader scientific use]

Significant synthesis capabilities exist within the package at
present, but are not sufficiently widely used in the scientific
community. We identify the following reasons: 
\begin{enumerate}
\item[i] the system is
inherently, but necessarily, complex, due to the sophisticated model
adopted for both calibration and imaging; 
\item[ii] the current user
interface does not hide the details of this complexity sufficiently
well from the user; 
\item[iii] documentation in a form suitable for
beginning scientific users is not in an accessible form, and iv) some
capabilities are still missing which would allow greater integrated
use of the package.
\end{enumerate}

We note that significant progress has been made on many of these
items over the course of this year, and the primary focus of the
project at this time is application development for scientific
use. Thus the balance has been shifted away from infrastructure
development, where prudent and possible.

\item[Finalize interfaces]

The internal synthesis interfaces, which form an important part of
the synthesis design as a whole, need to be finalized for the core
synthesis capabilities, so that more advanced applications can be
developed by a broader range of developers. Such applications include
three-dimensional imaging and mosaicing, as examples. The interfaces
are more efficiently tested by developing advanced prototype
applications of the type considered, by evaluating the interface
design on this basis, and also by ongoing general design review and
testing.

\item[Time-critical local priorities]

It is important that the project meet time-critical targets required
for successful AIPS++ use at consortia sites, particularly where the
use of AIPS++ is closely integrated into the critical path required
for instrument use or operation.

\end{description}

\subsection{Development strategy}

An overall synthesis development strategy has been adopted in which
the key strengths of the AIPS++ system are used to achieve the
priorities listed above. These strengths include the programmability
and flexibility offered by the command line interpreter (Glish), and
the system architecture overall, as well as the fully general
calibration and imaging model. AIPS++ offers both improved user and
system programming as the command line, and greater communication
between the command line environment and applications.

The facility for greater programmability can be used to package the
existing synthesis capabilities in a manner which allows greater
scientific use. This includes the production of scripts to automate
basic calibration and imaging. Further user interface improvements can
and have been achieved through a customized graphical user interface
for important synthesis components (e.g. sky), and the implementation
of a command line parameter-setting shell. A separate requirement for
greater access to the package is that of targeted documentation for
scientific users.

The general calibration and imaging model allows new and unique
scientific capabilities to be implemented as a result of the more
sophisticated formalism. The strategy for adding new capabilities can
take two directions, namely: 
\begin{enumerate}
\item[i] development of basic capabilities
covering the entire reduction path as a priority (hereinafter the
"thin-path" strategy); 
\item[ii] the development of more sophisticated
capabilities across a broader front, moving incrementally into areas
already covered by existing packages. 
\end{enumerate}

In the second option, data are transferred to the new package at a
well-defined point, for further reduction, with this point moving,
over time, to cover a larger fraction of the existing packages. In
this scheme, new scientific capabilities not already available in
other systems appear earlier, and have higher precedence, than in
option (i). Most data reduction packages for radio-interferometry
developed to date have followed development path (ii), and this has
been the general AIPS++ synthesis strategy.

The question considered here is the change required to the overall
synthesis development plan to accommodate a thin-path capability for
connected-element data as a short-term priority. The components and
resources required are summarized next.

\subsection{Thin-path development}

The primary changes in the current synthesis development plan to
implement a thin-path capability lie in the following areas:

\begin{enumerate}
\item[1] Enhanced uv-data display 
\item[2] Interactive editors
\item[3] Data fillers for the VLA 
\item[4] Associated utilities used in early processing
\end{enumerate}

Calibration and cross-calibration have been the subject of active
AIPS++ development this year and are consequently not considered as
part of the thin-path components. The core components of these
capabilities form part of the planned beta-3 release. Imaging and
self-calibration has been available for some time in AIPS++, and are
similarly not included.

Thus, the thin-path capability requires a change in the priority of
the elements required in early data reduction. Primary amongst these
would be the implementation of a VLA filler. Note that other
connected-element instruments including WSRT and the ATCA already have
preliminary fillers, written by local consortia members. In some cases
these are simpler to implement, due to a smaller number of archive
formats or related restrictions. Experience at NRAO has shown that
data fillers in other packages supporting a full range of archive
formats, and multi-volume input media have been a significant effort
for the VLA and VLBA. This is mostly related to the evolution of the
on-line output formats and the mismatch between the on-line and
off-line data format requirements. In addition, concatenation of
multiple tape volumes to existing files on disk adds further
complexity, as information is revealed to the filler on a per-record
basis and not for the dataset as a whole.

Data display utilities include those required for all the customary
graphical displays of uv-data (eg. visibility time series, spectra,
closure data), and the capability to list data in tabular form.
Interactive editors are those applications which allow uv-data to be
edited in a fully-featured, interactive graphical environment
including data edited per baseline in line-graphic form, or stacked in
axes of time, frequency or baseline in a raster TV display.

The associated utilities listed above, refer to those applications
required in support of early calibration or editing. Examples include
utilities to set and determine absolute flux density scales, or
correct for antenna or source position errors.

The definition of what constitutes a thin-path is problematic in
practice. It is highly dependent of the nature of the observations
being processed, and may differ significantly even within a single
category of observations (eg. continuum). It requires the basic
capabilities listed above. If the thin-path capabilities are
inadequate there is the real risk of alienating users or pulling the
project in directions away from fundamental work which needs to be
completed in imaging and calibration. This would limit the overall use
of the package if neither option received sufficient resources to
achieve critical objectives as a result. The implementation of
makeshift thin-path capabilities would ultimately prove expensive, as
they would need to be rewritten in the future.

A hybrid development plan to address the thin-path question is
described next.

\subsection{Development plan revision}

The project is opposed to a full-scale diversion of effort into early
processing capabilities at this stage, due to the risk this poses to
fundamental work underway in calibration and imaging, and consequently
the overall scientific use of the package. However, an approach has
been adopted and is described here which meets key thin-path
requirements without compromising broader development.

The key elements of this strategy include: 
\begin{enumerate}
\item[i] immediate work on a
basic VLA filler; 
\item[ii] work on uv-data display and editing modules
which can be used in automated data processing and by basic
interactive editors and display applications; 
\item[iii] simultaneous work
on higher-level calibration and imaging currently underway within the
project consistent with the original development plan.
\end{enumerate}

This plan can be achieved as a result of the recent expansion in the
number of synthesis application programmers in the project, consistent
with the increased emphasis of applications over infrastructure, and
some diversion of effort from VLBI priorities. The new synthesis
programmers are Mark Holdaway and Peter Barnes. It is unclear on
what timescale Dan Briggs can be replaced.

In this approach, Ralph Marson (25\%), assisted by Athol Kemball, will
start immediate work on a basic VLA filler for the limited public
release. The design will be approached such that components can be
re-used for a VLBA filler at a later time. The development of basic
data display modules consistent with the image display library
requirements, allows work to proceed on automated data reduction
scripts and simple interactive editors using the same display
components. The interactive editors will not offer all features
available in other packages but should meet the thin-path
requirement. Most importantly this approach allows the goal of scripted
processing to proceed unimpeded, and to offer a new, attractive
scientific capabilities which are part of the current synthesis goals
listed earlier. Peter Barnes will assist in the area of display
utilities.

Utilities associated with early processing, such as flux density
scaling and baseline corrections, will be scheduled at a higher
priority within calibration development.

High level work in calibration (A. Kemball/M. Wieringa/W. Brouw) and
imaging (3D-D.Briggs' replacement; Mosaicing-M.Holdaway/W.Brouw;
General imaging and solvers-T.Cornwell) will proceed essentially as
originally planned, thus preserving the original goals in synthesis
development.

The primary cost in this approach is the slower development of more
advanced calibration capabilities, especially in VLBI. Note however,
that VLBI development will not be terminated in its fundamental form
(such as fringe-fitting). Work on more advanced capabilities such as
phase-referencing or geodesy will have to be deferred however, as will
more advanced calibration solvers in connected-element reduction.

\section{Developments planned for 1998 Q3}

In {\bf Single Dish}, the Dish program will be included in the next
beta release.  This will fully support single dish FITS data into
Dish.  Existing Dish operations will be made to work on
multi-polarization data.  It is expected that both Parkes and Arecibo
staff will being testing this version of Dish and providing the first
real-user feedback of Dish.  An initial set of GBT commissioning tools
for continuum data will be ready by the end of July.  The internals of
Dish will be rewritten to make it easier to use Dish through the
command line i nterface as well as to make it easier to add operations
and site-specific features.  The existing gbt fillers will be reworked
to improved their speed and usefulness with on-line data.  Work will
start late in the quarter on Single Dish imaging using the same
imaging scheme as used in Sky.

In {\bf Synthesis}, we will continue development along the lines
spelt out in the synthesis development plan, and described above
in the section on  Thin Path development.

In {\bf Glish} support, In the next quarter, we hope to perform the
following, relatively straightforward modifications:

\begin{itemize}

\item prepare Glish for the next AIPS++ beta release and release another
        public version of Glish

\item switch over completely to the client based Tk widgets

\item add dynamic loading of TCL scripts and shared libraries to the
        TCL/Tk client

\item add the option for "eval()able" strings to Glish's as\_string()
        function
\end{itemize}

In {\bf Image Analysis}, the major task is to cement the region and
mask handling infrastructure and propagate it to Glish with a good
interface for users.  We plan to:

\begin{itemize}
\item Extend the C++ region classes to offer more world coordinate based
    classes, in particular classes to handle the extension, union, 
    difference, complement etc of any combination of lattice coordinate
    and world coordinate region objects
\item Implement the persistence of regions of interests and masks.  Currently
    these are both ephemeral.
\item Continue to develop the Glish-based user interface of region handling.
    This means
\begin{itemize}

\item design a special syntax for the Lattice Expression Language classes
     (these classes provide the support for the image calculator application)
     so that they can flexibly handle regions.  This syntax might also
     be generically available to other applications
\item Develop a GUI application which "manages" regions of interest.  That
      is, it keeps track of regions you create, allows you to manipulate
      them, transfer them and save them.

\end{itemize}
\end{itemize}
If there is time available, probably the next major application
    to be written will be that of image component fitting.

In the {\bf Display Library}, the next quarter should see the DL
evolve into a "very-close-to-useable" aips++ module, with capabilities
of displaying rasters and contours of 2-dimensional slices of
n-dimensional Lattices taken along any of the pixel axes.  Support
will be available with such displays for colour table manipulation,
zooming, and cursor feedback. We will also develop a schedule for the
generation of applications thereafter.

David Barnes will visit John Pixton at NCSA during this quarter.
Unfortunately, John leaves the NCSA to work in Norway in early August.
David's visit is partly aimed at maximising their coordinated output,
but also to maximize knowledge transferrence to David.  It is hoped
that DL infrastructure will be largely in place by the time John
leaves, but no doubt there will be some loose ends.  John's
considerable talents will be missed.
  
In {\bf Infrastructure}, Darrell Schiebel will take over Brian
Glendenning's work on the user interface, adding a number of
facilities to the GUI and also developing special purpose widgets,
{\em e.g} for inputting Measures such as Directions.

In {\bf Measures}, we will continue to improve the implementation
in various ways and add some new features as needed for applications:

\begin{itemize} 
\item unify some calculations (notably conversions in directions, 
baselines, uvw, magnetic field)
\item include some kind of Earth tide calculation in baselines and directions
(also raises related questions about retarded baselines and the like)
\item include an option to have tables of external position calculations
(e.g. for solar system bodies) as a basis for direction conversions
\item extend positions and directions to include derivatives (plate motion, 
p.m., radial velocity, distance)
\end{itemize}

Wim (our sole contributor to Measures) will have to spend some time on
SOFA work in the next quarter. Up till now it has mainly been limited
to comments on work by others, but some coding and/or testing is
imminent.

\section{Appendix: ATNF contribution {\em  Neil Killeen}}

\subsection{General}

This quarter the ATNF had 4 people working in AIPS++.  These are Neil
Killeen (25-75\%, also local manager), Wim Brouw (75\%), Mark Wieringa
(30\%) and now David Barnes (75\%) as of April 6 1998.   

We welcome David to the ATNF and aips++.  He is stepping admirably into
the role of responsibility for the Display Library and its applications. 
The DL has been largely dormant since Tom Oosterloo's departure last year.
Note that Wim's nominal contribution of 90\% has been decreased to 75\%
owing to growing 1KT committments. 


\subsection{System}

Our systems (Solaris/Gnu and Digital Unix/Gnu) were generally stable
this quarter and required close to zero maintenance.  However, the
documentation system became unstable towards the end of the quarter (for
all consortium sites), and we have been several weeks missing the users
reference manual, the glish manual and the notes and memos series.  The
problems arose through the non-trivial interaction of TeX, LaTeX, dvips
and latex2html. 

NCSA now appears to be making regular native IRIX (SGI) inhales.
We can probably now consider another attempt to build on
our SGI machine.


\subsection{Visits}

Athol Kemball visited us for three weeks.  He spent two weeks in
Narrabri working with Mark Wieringa, and one week in Epping mainly
working with Wim, but also with David and Neil.  We all found Athol's
visit very rewarding. 

Neil Killeen spent two days in Dwingeloo working with Ger van Diepen
on issues to do with regions of interest in lattices.


\subsection{Testing}

There has been no significant new system testing this quarter at the
ATNF (of course the multibeam people still process their data with their
Glish scripts built on aips++, and Glish/Tk forms the basis of the
Parkes Telescope Control System user interface).  We do have three users
who have expressed strong interest in using the single dish part of
aips++ to analyse their data (ATNF single dish software is not overly
impressive).  We are awaiting the final connections in "dish" to handle
reading and writing Single Dish FITS files, and multiple polarization
handling.  This work is expected complete by mid July 1998. 

\subsection{Individuals}

Wim Brouw has spent his time this quarter on:


Measures:

\begin{itemize}

\item after ample discussions about the proper inheritance structure and format,
    wrote (and got code copped) QuantumHolder and MeasureHolder classes, which are
    generic holders of any Measure or Quantum, and aimed at providing a general
    interface to the Tasking system. 
\item included the VLBI reference source list in the 'data' system, which is
    accessable from both the Glish (measures gui) and c++ code.
\item wrote the MBaseline (for interferometer baselines on Earth) Measure and its 
    coordinating classes: no Earth tide or other secondary effects included as yet.
\item wrote the Muvw Measure (uvw calculations) and its coordinating classes
\item added an angular shift method (in both true and projected angles) to the
    MDirection Measure
\item added de-projection of an 'off-axis' field to UVWMachine
\item improved some of MDirections forward/backward conversion precisions
\item finished EarthMagnetic field Measure using the IGRF reference magnetic
    field
\item wrote a FieldMachine to easily calculate the magnetic field at a certain
    direction in e.g. the ionosphere
\item wrote a VelocityMachine to convert between velocities and frequencies in a
    simple manner
\item added new ITRF and TOPO codes to MDirection
\end{itemize}

Various:

\begin{itemize}
\item wrote a scrollable 'menu' widget (listselector)
\item made popuhelp more general and containing 'long' and 'short' help.
\item discussions about calibration and synthesis with Athol Kemball
\item code copped RecordTransformable
\item some action (on behalf of QAG) on moving code from apps directories: more
    action needed
\item updated the IERS database in the system: should be done automatically
\item wrote article for newsletter on measures gui
\end{itemize}

Bugs and problems:

\begin{itemize}
\item traced an inconsistency in some Linux calculations (hardware circular
    functions near pi/2) and in Sun mathematical library (abs(+0)->-0).
\item problems with Solaris 2.6: decision about the aips++ default compiler should
    be made soon, with proper consultation about implementation
\item network problems on tarzan meant large time losses
\end{itemize}

\subsection{David Barnes}

David's time was spent on

\begin{itemize}
\item Working through existing high-level display library (DL) classes,
  in order to appreciate the extent of the existing code, and to
  identify capabilities that the library was lacking.
\item Constructing a class design to encompass the many different 
  possibilities for handling and display of astronomical data,
  based on the existing single-purpose class already provided
  in the DL.
\item Commencing to implement parts of the above design.  At the time
  of writing, the initial design is of order 30\% implemented.
\item Learning how to build a Tcl/Tk widget to contain a DL 
  WorldCanvas, and writing a very simple (and useless) widget
  to show "proof of concept."
\end{itemize}

\subsection{Neil Killeen}

Neil spent one month away in Bologna having some science time. Neil's
aips++ time was spent on:

\begin{itemize}
\item Ger van Diepen produced some new classes handling image masks.
He integrated these with the C++ image applications.
This work is ongoing, mask handling is not yet complete (we
can apply them, but not yet save them in an image)
\item Ger also produced initial classes handling lattice coordinate based
regions of interest for lattices.  Neil integrated these with the 
C++ image applications (this work is ongoing,
not all the required functionality  is there yet).
\item Worked on implementing some simple region of interest C++ classes 
which take their input as world coordinates rather than lattice coordinates.
\item Worked on the handling of image regions of interest within Glish.
This requires interfacing of the C++ classes to Glish, and providing 
a mechanism for users to specify and handle them.  
\item Did some designing on extra Coordinate class functionality that is required.

\item Spent some time evaluating and commenting on the new aips++ web pages
\item Evaluated Michael Rupen's document of suggestions for improvements to
the user's reference manual and sent reply.  The changes to the
documentation are what is delaying the beta release.  This is going
slowly owing to a lot of meetings happening at NRAO this month.
\item Constructed quarterly report
\end{itemize}

\subsection{Mark Wieringa}

Mark had little time this quarter for aips++ as a result of Gamma ray
bursters, work on local Year 2000 issues and other local work. He has
spent his aips++ time on:

\begin{itemize}
\item Some work on the MeasurementSet Distributed Object
\item The new ecgs compiler installation on Narrabri's Digital alpha
\item A lot of discussions over two weeks with Athol about synthesis issues
\end{itemize}

\section{Appendix: BIMA/NCSA contribution {\em  Doug Roberts}}

During the past quarter we had 5 people (2.8 FTE) working on AIPS++:
Dick Crutcher (5\%), John Pixton (50\%), Harold Ravlin (50\%), Doug Roberts
(75\%), and Dan Briggs (100\%). In addition, NCSA is funding the salary of
Wes Young at NRAO-Socorro and will shortly fund the salary of a second
NRAO person. Accomplishments during the past quarter are described below
under each person.

\subsection{Dick Crutcher}

He served as local AIPS++ project manager and as BIMA
director in charge of our participation in AIPS++.  The job of local
AIPS++ project manager is largely being taken over by Doug Roberts.

\subsection{John Pixton} 

He has focused on the AIPS++ display library, under the direct
supervision of Neil Killeen. He will be leaving NCSA and the AIPS++
project this summer.

\subsection{Harold Ravlin} 

He administers our AIPS++ workstations and maintains and improves
aipsview. His aipsview work included the following:
\begin{itemize}
\item Redid the way aipsview handles colormaps to allow it to run even if a
program like Netscape or the CDE window manager grabs all the colors of
the root colormap.
\item Fixed a problem with aipsview that would occasionally cause a crash.
\item Fixed a problem where the data set objects weren't being destroyed
when the "Close Data" menu item was chosen. This resulted in file
locking problems with AIPS++ data sets.
\item Answered a number of email questions about aipsview.
\item Several warnings generated by egcs and the new IRIX compilers were  
removed, although many other warnings seem related to system or compiler
files.
\item Stand alone aipsview compiles and runs with egcs on our Solaris 2.5,
IRIX 6.2 and Linux systems.
\end{itemize}
 
His system admin duties included the following:
\begin{itemize}
\item Installed the egcs compiler on Solaris, Irix, and Linux platforms.
\item Installed a new Sun Ultra 30.
\item Installed ZIP drives and a new VCR.
\item Upgraded memory on several machines.
\item Made a query on newsgroup about stuck pixels on Octanes. (Use gclear).
This led to the discovery of a bad frame buffer on an Indy. Our SGI Onyx
crashed on 5/22/98 with a bad power supply. A significant amount of time
was spent dealing with its after effects. We may still be having disk
troubles.
\item Installed xemacs-20.4 and other software on SGI systems.
\item Installed lots of security and general system patches.
\item Discussed and set up disk drives for NRAO's new Octane.
\item Made kernel changes on some of our IRIX machines to accommodate
AIPS++. These changes open up potential security holes so, hopefully,
the changes can be removed in the future.
\item Began installation of new DLT tape drive.
\end{itemize}

He also attended the NCSA Alliance '98 conference.

\subsection{Doug Roberts}

Doug Roberts has been working on using the optimized FFT routines in
the SGI/Cray Scientific Library (SCSL), available on the SGI Origin
2000.  The SCSL may become available for other platforms in the next
year or so.  The results of the FFT implementations is that on a single
processor the the SCSL is faster than the FFTPACK version by an
average of 50% over a variety of problem sizes.  The SCSL is available
in a multiple processor version as well, which should show
significantly more speedup on large problem sizes.  The SCSL has
built-in multidimensional FFT's 2D, 3D, and N-dimension.  Transforming
a 2-dimensional matrix using the built in 2D transform does not appear
to be significantly faster (within 10-20%) than decomposing the matrix
into 1D FFT's.  Doug is going to change the FFTServer code to use the
SCSL on the SGI's.

Doug Roberts has installed AIPS++ on the 4-processor SGI Onyx at the
astronomy department and on his SGI Octane.  The SGI systems have
built using the native SGI MIPSpro compilers.  On the Origin system, a 
red/black system has been set up to allow parallel development on a
stable system while code fixes are tested in a separate version.

\subsection{Dan Briggs}

As noted above Dan Briggs was killed in a sky-diving accident on July
5, 1998. We decsribe here his work over the last quarter.  Dan Briggs
continued to investigate parallel gridding.  Dan Briggs started design
work on extending the aips++ task 'sky' to include full wide field
imaging capability.  Dan rewrote chapters 6 and 8 of ``Synthesis
Imaging in Radio Astronomy'', one of the fundamental references for
new AIPS++ users.  Dan traveled to Socorro June 16 through June 23 to
present imaging lecture at NRAO summer school and to work with Brian
Glendenning and Tim Cornwell on wide field imaging design.  Dan
installed Classic AIPS on NCSA SGI workstations for comparison use
against AIPS++.  Dan made minor bug fixes involved with the port to
the Origin 2000.  He took a more active role in assisting with AIPS++
system administration, in support of getting a reliable port and
installation on the Origin 2000.  Worked on Local Organizing Issues
for the upcoming ADASS meeting.

\section{Appendix: NFRA contribution {\em  Jan Noordam}}

Local project members: Ger van Diepen (GVD), Michael Haller (MXH), Jan
Noordam (JEN, local manager).


\subsection{General}

The elevation of Ger van Diepen to Interim Technical Leader from
September will make things more difficult locally at NFRA. However,
its is probably the most effective thing to do in the short term, and
various Good Things may come from it.

Athol Kemball has made a very fruitful visit to Dwingeloo in June. 
All the more so since it coincided with the yearly visit of Wim Brouw.

Bob Campbell (postdoc at NFRA) has made a large contribution to the
definition of ionospheric calibration in AIPS++. He will remain
involved in the near future.

This quarter, the NFRA contribution in FTE's was as follows:
\vskip 0.5in
\begin{tabular}{|l|c|c|c|}
\hline \
&                     Global AIPS++ &     Local AIPS++ &   Other\\
\hline \
Ger van Diepen &     40\%    &             50\%        & 10\%    \\
Michael Haller &     50\%    &             40\%        & 10\% \\
Jan Noordam &        30\%    &             40\%        & 30\%\\
\hline
\end{tabular}
\vskip 0.5in

\subsection{WSRT upgrade}

Commissioning of the new WSRT backend (DZB) and on-line software (TMS)
is going more slowly than expected. This has made it easier for JEN to
keep pace with developments in providing local AIPS++ applications.

Actual data reduction still takes place in other packages (NEWSTAR,
AIPS, DIFMAP). The various data conversion programs are sufficient for
the moment, but would have to be upgraded for more advanced observing
modes. Therefor, processing in AIPS++ itself remains a high priority.

\subsection{Applications}

In the context of writing local applications for the WSRT upgrade, JEN
has created a number of Glish scripts that might be useful for other
applications. They include a 'standard' application interface
consisting of a menubar, a textwidget, a statusline and a dismiss
button. This has proved a rather useful concept in practice,
espacially because of the various parameter-input services of the
menubar object. In addition, there is a general tool for inspecting
Glish variables of any shape or type (including tables), and an
on-line help summary of the Glish language.  

MXH's work on DatalineTool (user of pgplotter for visualiser) will
make use of the application interface sketched above. It will have
many more possibilities for user-interaction with the plot than its
gplotid-based predecessor xyftool. 

The progress of the HOLOG (holography) program by GVD was hampered by
the necessity to add support of system temperatures in the conversion
programs.


\subsection{Fillers and converters}

The conversion program ms2scn has been extended by GVD to take the
system temperatures into account.

Class UVFITSWriter has been changed by GVD to optionally write the
system temperatures and gain curves to a UVFITS file.  It makes it
possible to import these aspects of a WSRT measurementset in classic
AIPS.

Harro Verkouter (JIVE) has taken on responisbility of a more permanent
version of the UVFITS writer that will alos allow uv-data calibration
in AIPS (rather than just imaging).

\subsection{Infrastructure software}

GVD has worked with Killeen to fully support regions/masks in the
lattices and images. On 15-16 April Neil paid a very fruitful visit to
Dwingeloo.  The amount of email messages exchanged must be
enormous. It's a good thing Neil works late, so our working days
overlap a great deal. A class MaskedImage has been introduced to make
the inheritance tree for SubImage and ImageExpr better.  The LCRegion
classes have been changed a bit (e.g. for better round-off)
LatticeApply and associated classes have been changed to make use of
regions/masks possible by operating on MaskedLattice instead of
Lattice. Also a few other improvements have been made to these
classes. Neil uses them in his Moments and Statistics classes.  The
LatticeExpr classes have been changed similarly. They operate as best
as possible by only taking a mask into account when present.

Apart from a few small bug fixes and optimizations, not much has been
done on the table system.  The most notable bug fix is the .nfsNNN
problem. When a table was accessed on an NFS-mounted disk, sometimes a
.nfsNNN file appeared which prevented the table from being
deleted. Finally I learnt that NFS renames an open file to .nfsNNN
when it is deleted before being closed. Indeed a file was not closed
in special circumstances.  The table client got some new functionality
to make it possible to handle table locking in glish scripts.  To
improve the performance of access to reference tables several
XXXColumnCells were added to the data managers. They greatly reduce
the calling and sometimes locking overhead.  A class TableLocker has
been introduced to make automatic release of locks possible,
especially in the case of exceptions. For this purpose it has been
derived from Cleanup.

On 4-5 June GVD has attended a PAO-course on ODMG 2.0, the new
standard for object databases. It was very interesting, especially the
subject of OQL (its query language). Ger is planning to introduce some
OQL concepts in the table query language.

On 22 June GVD attended a day devoted to Corba.  It was rather
interesting. It was good to hear that Corba will support DCOM.

\subsection{AIPS++ Site in Dwingeloo/Westerbork}

At least 20\% of Michael Haller's time was spent on either normal
maintenance or a variety of problems.  One major problem here has been
the aips++ documentation.  In almost all cases the source of
documentation problems has been differences here in documentation
package versions to those in Socorro.  Identifying these problems has
been a rather long-winded process but the recent movement to
latex2html version 98.1 inspired a focused effort to fix these
problems and with the help of Henk Vosmeijer and Wes Young the we
appear to have succeeded.

MXH also set up binary distribution of AIPS++ on linux machines at
Westerbork and JEN's laptop.

\section{Appendix: NRAO contribution {\em  Tim Cornwell}}

The core NRAO AIPS++ group is now Barnes (100\%), Cornwell (100\%),
Garwood (100\%), Holdaway(100\%), Kemball(100\%), Marson (100\%),
McMullin (100\%), Schiebel (100\%), Uphoff (100\%), Weatherall (50\%),
and Young (100\%).

\subsection{Tim Cornwell}

Tim:
\begin{itemize}
\item Managed AIPS++.
\item Conducted extensive discussions connected with Brian's departure from
the Project.
\item Directed the alpha testing at the AOC.
\item Made presentations and demonstration concerning AIPS++ at the 
AUI Board of Trustees and at the NRAO Visiting and Users' Committees.
\item Continued testing and optimization of sky synthesis application.
\item Made radical changes to the graphical catalog browser.
\item Developed a new graphical visibility visualizer based on Glish/tk,
the ms module, and the pgplotter interactive plotter. The development of
this was seen as a critical priority by the alpha-testers at the AOC.
\end{itemize}

\subsection{Bob Garwood}

Bob Garwood's primary responsibility is to oversee the single dish work which 
currently is focussed on the DISH environment as well as supporting the GBT.

Over the past 3 months (April through June) he has done the following work: 

\begin{itemize}
\item Enhanced fits2table (and the underlying classes) so that the full SDFITS
        convention is now supported.
\item Added a save operation to DISH so that individual results can be saved to
        any opened data source.
\item Enhanced the file open menu in DISH so that a new data file can be 
        created  and files can be opened for reading and writing as well as 
        just read-only.
\item Made a number of changes to the file chooser GUI tool to support the 
        new file open options in DISH.
\item Started work on reorganizing the internals of DISH so that it will be 
        easier to use it without the GUI and to make it easier for others to
        add new features and site-specific features.
\item Attend the regular GBT coordination meetings. 
\item Visited Green Bank 2 times. 
\item Supported the various data fillers used by the GBT staff through bug 
        fixes and new support for the GBT Holography backend.
\end{itemize}

\subsection{Brian Glendenning}

Brian Glendenning attended Alliance '98 at NCSA, and had many
discussions about parallelizatin in AIPS++. It was decided to attempt
a Windows NT port if a suitable candidate can be found, to enable
AIPS++ to run in a parallel fashion on a Windows NT cluster. We also
decided to hire a person to look into parallel I/O for AIPS++. Brian
helped define an applicator/algorithm set of classes which should
allow lattice based computations to parallelize very readily.

Brian made a number of small fixes to the tasking system, and in general
started winding down his AIPS++ activities.

Brian gave presentations at the NRAO Visiting Committe, and at the Clark
Symposium.

\subsection{Athol Kemball}

Time has been spent this quarter on synthesis planning and coordination.
In particular Athol visited ATNF for three weeks and NFRA for 10 days
in this regard. This has allowed detailed synthesis planning at all
levels and coordination with the management at both consortium sites.
A draft synthesis development plan incorporating these changes has
been developed.

Significant effort has continued in cross-calibration development, as
part of the continuing implementation of this capability for the third
beta release. Discussions of calibration questions have been held
throughout the project.

The MS v2.0 definition is essentially finalized and will be implemented
soon after the third beta release. A convention for data selection in
synthesis has been proposed and discussed.

A command-line parameter-setting shell has been developed as a prototype,
and will be made available in the next release. This user interface uses
the same meta-information as the existing GUI components.

\subsection{Ralph Marson}


In the last quarter Ralph has spent most of my time in the cleaning up,
inproving, and testing the code in the ComponentModels module. This
module is used to represent components of the sky brightness using a
functional form. Some of the things that have been done are:

\begin{itemize}
\item Implement a major structural change which allows a much more flexible 
  mix of spectral ans spatial variation. As far as I know this module
  is the first one that models both spectral as well as spatial structure.
\item Implement a new structure for the way the component Tables are
  written to disk. This new structure is incompatible with the older
  one but future changes are expected to be compatible.
\item Make much all of the new functionality avaliable to the user. This
  includes the doubling of the number of methods in the componentlist
  distributed object and documentation for every method.
\item testing of the module at the user level and the use of components in 
  the sky module (for moddelling of visibility data).
\item The addition of a disk component. This was relatively simple given
  the structural changes mentioned in point one.
\end{itemize}

He has also spent some time this quarter examining some speed issues in 
the Clark clean and testing the egcs compiler as a alternative to the
gnu compiler. Code cop duties typically take 10\% of Ralph's time.

\subsection{Joseph McMullin}

Joe's main responsibilities are managing the Green Bank
installation of AIPS++ and supporting AIPS++-related GBT needs.
 
\begin{itemize}

\item Developed some basic functions required for commissioning of the GBT
(scan info, Tsys, Power Spectra). Worked on compiling information
on the DCR and writing a guide to taking and analyzing DCR data (in
progress).
 
\item Worked with the Monitor and Control group to define and add to the definition of the GBT FITS
and Scan Log.
 
\item Implemented a weather station monitoring system through AIPS++.
 
\item Currently working on software for commissioning requirements and in
support

\end{itemize}

\subsection{Darrell Schiebel}

Darrell Schiebel implemented all of the Glish changes and additions
described in the section on progress in the last quarter. He also
worked on the GBT observer table parser client.

\subsection{Jeff Uphoff}

Jeff worked in a wide range of areas, from supporting dishplot to
supporting the AIPS++ system.

He looked into accuracy questions for cos() under Linux/Intel that Wim
reported.  Answer: it's apparently a known problem.

He continued work on Dish/Dishplot:
\begin{itemize}
\item         Bug-fixes as needed.
\item         Range-setting tweaks.
\item         Baseline statistics behavior.
\item         Restrict which spectra can be un-overlaid.
\item         Work toward better plot/menu color synchronization.
\item         Changed to use pgplotwidget.g for PGPLOT interface.
\item         Started adding state-saving to plotter.
\item        Started on plot-surface tiling.
\end{itemize}

He continued work on AIPS++ infrastructure/system:
\begin{itemize}
\item         Makefile system, makedefs tweaks.
\item         version.o dependency issue.
\item         ai/ao/etc. replacement system for Linux and Digital Unix.
\item         Tweaks to glish-mode.el; a Glish editing mode for Emacs.
\item         latex2html debugging, upgrading, testing.
\item         Testsuite issues.
\item        Debugging, fixing, etc. of inhales/sneezes all over the place.
\item        RCS system.
\item         Linux RPM building for AIPS++-related packages: EGCS, binutils,
latex2html, GDB, PGPLOT (for Lesstif).
\item         Started investigating CVS client/server protocol.
\item         Created and documented new "blessed" inhale system.
\item         Hacks to exhale to make debugging easier.
\item         exhale no longer expires old RCS information unless requested.
\item        Tweaks to testdefs, avers scripts.
\end{itemize}

He also worked on other miscellaneous tasks:

\begin{itemize}
\item Fixed problem with ENOTEMPTY under Linux in tDirectory.

\item Started code-copping Guiutils/progress.*...not done yet.

\item Some minor fixes/enhancements to Brian's pgplotwidget.g for use by
Dishplot, e.g. added button-specific callbacks.  Looking into postscript
printing/file-saving problems, which haven't yet been resolved.

\item Created a relatively simple state-saving mechanism for use by Glish
scripts.

\item Helped several people through (often problematic) AIPS++ installations
\begin{itemize}
\item         Stephen Myers, Caltech.
\item         Tony Willis, DRAO.  (Raised issues such as Motif/Lesstif.)
\item         Mark Holdaway, NRAO-TUC.  (Required some sysadminning too.)
\end{itemize}

\item Set up Usenet mail2news gateway for EGCS mailing lists, at Brian's
request...required some local hacking of gateway software, etc.

\item Worked on building AIPS++ against Lesstif (various versions).  Verdict:
it builds, but doesn't yet run correctly.

\item Updated FTP import area on Tarzan with new versions of AIPS++
prerequisite packages.

\item Investigating weird compilation problems at NFRA.

\end{itemize}

\subsection{Kate Weatherall}

\begin{itemize}
\item Writing documentation on Revision Control System.  Handed off to Peter
Barnes to see if it is useful.

\item Copyedited and produced  AIPS++ Online Newsletter 
for April and June. 

\item Restructured and changed web site as suggested by Killeen, Brouw, and 
others.

\end{itemize}

\subsection{Wes Young}

\begin{itemize}
\item Conducted discussions and roughed out Applicator/Algorithm classes to integrate  parallelization work into the AIPS++ system.

\item Provided support for SGI port of AIPS++.

\item Made the aips++ on-line help AIPS++ table based.

\item Worked on improvements to the "new table browser".  These include 
enable field editing and improving the rendering speed.

\item Modified runtests to produce summary results by package and send 
aips test results to module bosses and trial results to aips2-inhale.
 
\item Provided documentation support.  

\item Ordered and installed an SGI Octane for AIPS++ development at the
AOC.
\end{itemize}

\section{Appendix: Summary of AIPS++ Personnel Changes}

In this section, we give the names of people in the various AIPS++
groups and the nominal fraction of time allocated to AIPS++.
These are as expected for the next quarter (and thus do not
include the expected replacements for Briggs or Uphoff, or the
new hires in Socorro).

The ATNF group is: Neil Killeen (25-75\%, also local manager), Wim
Brouw (75\%), Mark Wieringa (30\%) and now David Barnes (75\%) as of
April 6 1998.

The BIMA/NCSA group is: Dick Crutcher (5\%), 
John Pixton (50\%), Harold Ravlin (50\%), Doug Roberts (75\%).

The NFRA group is: Ger van Diepen, Michael Haller, and Jan Noordam.
The allocations are:

\vskip 0.5in
\begin{tabular}{|l|c|c|c|}
\hline \
&                     Global AIPS++ &     Local AIPS++ &   Other\\
\hline \
Ger van Diepen &     40\%    &             50\%        & 10\%    \\
Michael Haller &     50\%    &             40\%        & 10\% \\
Jan Noordam &        30\%    &             40\%        & 30\%\\
\hline
\end{tabular}
\vskip 0.5in
The NRAO group is: Tim Cornwell (100\%), Bob Garwood (90\%), Mark
Holdaway (100\%), Athol Kemball (100\%), Ralph Marson (100\%), Joe
McMullin (100\%), Darrell Schiebel (100\%), Kate Weatherall (50\%) and
Wes Young (100\%). We have one open position in Green Bank and two
new positions in Socorro (funded by NRL and NCSA).

Thus, in aggregate, we have 21 people contributing just over 17 FTEs
to the AIPS++ Project. Of these, 9 are employed by NRAO, and contribute
9.4 FTEs. The numbers for the other partners are: ATNF 4 and 2.80,
BIMA/NCSA 4 and 1.8, NFRA 3 and 1.2.

\end{document}

