\documentclass[11pt]{article}
\usepackage{html, epsf}
\begin{document}
\title{\bf AIPS++ Quarterly Report: 1999 Q2}
\author{T.J. Cornwell, NRAO}
\date{1999 August 11}
\maketitle

\tableofcontents

\section{Introduction}

This quarter, we have focused almost exclusively on preparing
the first release. This included testing the system, fixing
defects, and preparing and testing the actual CROM release. 

Other than testing and fixing defects, little new work has been
performed. Most of the new work has been in the application areas of
single dish, synthesis and visualization.

\section{Developments in 1999 Q2}

In {\bf Single Dish support}, Jim Braatz joined the single dish group
in aips++ in April.  He will be working out of Green Bank, replacing
Joe McMullin who is now based in Charlottesville.  Garwood, McMullin
and Braatz continue to be involved in supporting the GBT.  This
includes regular visits by Garwood and McMullin to Green Bank and by
Braatz to Charlottesville.

The revised GBT filler was used in a planned test of the GBT
holography backend in April.  The test went well exposing one bug and
several shortcomings of the filler.  The filler is being revised in
time for another planned holography backend test in late July.  This
revision will include the necessary framework to handle on-line data.

Much of the additional GBT support was focussed on transferring 
knowledge and responsibility of the GBT commissioning support,
tipper support, and other routine Green Bank support tasks
from McMullin to Braatz.  This has been mostly completed.

Much of the non-GBT focus during this quarter was on finding and 
fixing bugs in the single dish software. The dish internal rewrite
was postponed in order that dish be ready and debugged in time
for the upcoming release. The dish plotter 
continues to be improved.  A command-line Gaussian component
fitting operation was added to dish using available tools.
A plug-in for the dish plotter was written to use the JPL
line list for quick line identifications.  The documentation for
dish was also improved.

Work was begun this quarter on a suite of single dish imaging
tools to be used in the dish environment.  Three tools have 
been developed so far.   lookmap reads a dish data set and plots
the spatial locations for all of the data in the set relative to
a selected scan.  spectramap reads a dish data set and plots the
spectra at their relative spatial locations.  contourmap reads
a dish data set and plots contours of integrated intensity.
These tools will be improved upon and used in conjunction with
a single dish on-the-fly imaging tool to be developed next
quarter. 

In {\bf Synthesis support}, in keeping with the rest of the package, a
significant focus in synthesis this quarter has been the stabilization
of current capabilities in preparation for the release. This has
included extensive testing to isolate defects, and their subsequent
correction.  This has been a necessary and beneficial process in
preparation for wider deployment of the system in the scientific
community.

The context of the development that has taken place, however, remains
the same as in previous quarters. The strategy remains that of
developing advanced synthesis capabilities in key areas to test design
completeness and sufficiency, while also developing a "thin-path"
reduction sequence for connected-element interferometry. As previously
noted, the resources for thin-path development have primarily been
assigned at the expense of some VLBI development. This is expected to
be reversed after the first release, in preparation for the second
release which will contain greater VLBI support.

High-level synthesis development efforts have been concentrated
primarily in the area of wide-field imaging (T. Cornwell), and in
mosaicing, primary beam effects and general deconvolution (M. Holdaway
and T. Cornwell). Kumar Golap, who arrived as a Visiting Scientist at
the NRAO during this quarter, has provided significant testing of the
wide-field imaging capability on an multi-processor SGI system in
Socorro, in preparation for his future planned work in parallelization
of these algorithms. This has included testing against simulated
wide-field data, and the development of basic automated scripts to
extract and test derived image component positions. He has also
written a cookbook chapter on how to use wide-field imaging
capabilities in AIPS++. The wide-field development work has included
an application, called dragon, to replace and extend the capabilities
previously available in SDE under this name. The NRL wide-field
imaging group have agreed to assist in testing the new features using
74 MHz VLA data. The mosaicing and deconvolution work has included
improved primary beam weighting, the migration of multi-resolution
CLEAN deconvolution to the imager tool, and the implementation of a
multi-field Hogbom CLEAN for comparison against the existing Clark
CLEAN implementation. Various other changes and defect corrections of
benefit to mosaicing have also been implemented.  Time has also been
spent on preparations for the Mosaicing Techniques meeting, planned
for mid-July in Socorro, which is being arranged by M. Holdaway and
M. Rupen, and is a collaborative effort between AIPS++ and the MMA
project.

General defect correction and testing has continued for the imager
and calibrater tools (T. Cornwell and A. Kemball), including some new
features designed to improve ease of use. Synthesis has also benefited
overall from the improved capabilities and features in the graphical
user interface. The MS v2.0 definition has been stable in this
quarter, and M. Wieringa has continued the propagation of changes
required for this format revision throughout the affected classes in
the library. It is planned to adopt the new format, which will allow
new single-dish and synthesis reduction approaches, after the first
release. User impact will be minimized in this process. There has been
a significant contribution to the format specification from across the
project as a whole, and we have benefited greatly from the diverse
instrumental experience and opinions brought to bear in the discussion
of this matter.

NFRA synthesis development has been focused on utilities required to
support the commissioning of TMS at WSRT (J. Noordam and G. van
Diepen). Development for WSRT will expand in the future with the
hiring of new scientific staff to assist work in this area.

Thin-path development has continued during this quarter. R. Marson
has further refined the VLA data filler, and expanded the supported
scientific capabilities. It has been tested in the supported modes and
is being compared to existing data fillers for the VLA. P. Barnes has
continued work on uv-data visualization in the framework of the
Display Library, in collaboration with D. Barnes, N. Killeen and
A. Kemball. A basic pgplotter implementation for some of these
capabilities has been completed in the interim.

The parallelization effort forms an important part of synthesis
development, to address the largest, most computationally demanding
applications. Work on a test mosaic dataset taken on M33 at the VLA
have continued, as we build parallelized capabilities in this
area. 

Specialized design work in this quarter has included a holography
requirements document (A. Kemball), which was submitted and accepted
as part of the ALMA holography design review. Work on integrating
single-dish and synthesis reduction in a common design framework has
continued in the framework discussed in previous quarters, but at a
slower pace this quarter.

In {\bf Glish}, garbage collection was added as an optional feature to
Glish. This collector cleans up memory which reference counting fails
to delete. Typically this memory is allocated as part of records which
are self-referencing through functions and other records.

In addition to bug fixes, some clean-up work was also done with Glish
shared library creation to attach version numbers to the libraries.

In {\bf Parallelization}, most of the second quarter Roberts has been
working with the Pablo group at UIUC CS department to identify the IO
bottlenecks in large image runs.  We found a few files that did not
have the proper instrumentation.  Also in order to trace where IO was
being carried out some event tracing needed to be introduced into the
imager and pimager applications.  The trace calls (IO and event) are
checked into the code and can be turned on by an makedefs variable
(PABLO\_IO) We now have a stable version of pimager which we have been
using for testing.  We have been carrying out tests on dedicated
machines and intend to work on a single machine for a long test in
late July.

In {\bf Measures}, no substantial changes were made.

In {\bf AIPS++ Infrastructure}, little has happened beyond bug fixes
and some minor integration improvements.

In {\bf Image Analysis}, we added various new convolution
capabilities, improved the interfaces for various parts of the image
tool, and improved the performance of the image fitter.

In {\bf Documentation}, we finished the revision of {\em Getting
Started in AIPS++}, started on a cookbook, and implemented a FAQ.

In {\bf Management}, we made demonstrations and presentations at the
Chicago AAS, and at the Tucson meeting on Sub-Millimeter Imaging. We
held discussions with the JCMT/ACSIS group on their requests for
assistance. This amounts to 6-7 weeks of time spread over about
9 months.

In {\bf System}, we directed nearly all of our effort to production of
the initial CDROM. Some planning for the post-release phase has been
completed. We expect to move towards using CVS for code management and
distribution.

In June, our server tarzan was the subject of a breakin. This led to
the machine being out of commission for about 6 days, during which
code development at the sites was hindered. Following this attack,
we have made a number of changes designed to improve the security
of tarzan. The principal ones are that remote NFS is no longer
permitted, and the secure shell must be used to connect to tarzan.
The resulting changes in working habits for remote sites are
acceptably small.

In the {\bf Quality Assurance Group}, we embarked on a review
of the compliance with various rules, resulting in, for example,
closer adherence to our guidelines for writing Graphical User
Interfaces.

\section{Preparation for the First Release}

The first release was originally planned for the Chicago AAS meeting
in early July. As a result of the testing results, we realized that
this was too aggressive and decided instead to opt for a round of
pre-release testing. This pre-release testing will last for 6-8 weeks
and will start with the distribution of a candidate release to
consortium sites and beta testers. We expect the pre-release to
happen this week and the actual release to be performed by the end
of September.

The major task in preparing for the release has been the extensive
testing of the package. This has been performed primarily by
Project Staff and accounts for most of their work in this last
quarter. Bugs reports (now known as defects) are managed using
the ClearDDTS system that NRAO purchased for Socorro. This enables
tracking of the history of each defect report, and relays all
events to the submitter, and allows the submitter to verify that
the defect has indeed been fixed. Many types of statistical
reports can be generated, and we use such reports to help
direct work. In particular, the current scoresheet of
defects arrival and repair rates is circulated at the end of
evey week. We can also look at the number of assigned defects
per person and even out the load by delegating some repairs
to other people.

At the time of writing, the defect arrival and repair rates per
week were:

\vskip 0.5in
\begin{tabular}{|c|c|c|c|c|c|}
\hline
     Week &        Date   &     \# New &   \# Resolved &     Diff&    \#
Unresolved \\
\hline
       1  &    4-25-1999  &        9  &        0     &       9 &
9 \\
       2  &     5-2-1999  &        5  &        2     &       3 &
12 \\
       3  &     5-9-1999  &        3  &        0     &       3 &
15 \\
       4  &    5-16-1999  &       15  &       22     &      -7 &
8 \\
       5  &    5-23-1999  &       35  &       13     &      22 &
30 \\
       6  &    5-30-1999  &       39  &       18     &      21 &
51 \\
       7  &     6-6-1999  &       41  &       11     &      30 &
81 \\
       8  &    6-13-1999  &       18  &        7     &      11 &
92 \\
       9  &    6-20-1999  &       21  &       11     &      10 &
102 \\
      10  &    6-27-1999  &       37  &       24     &      13 &
115 \\
      11  &     7-4-1999  &       60  &       35     &      25 &
140 \\
      12  &    7-11-1999  &       45  &       44     &       1 &
141 \\
      13  &    7-18-1999  &       18  &       27     &      -9 &
132 \\
      14  &    7-25-1999  &       46  &       32     &      14 &
146 \\
\hline
\end{tabular}
\vskip 0.5in

The above list includes all defects, ranging from severity 1
(catastrophic) to severity 5 (cosmetic). If we include the defects in
the old GNATS system, in the history of the project we have processed
about 1500 defects, of which about 10\% are still open. Most of the
unresolved defects are now of severity 4 and 5, and are relatively
minor.  For the severity 1 and 2 errors (the most serious), the rates
are:


\vskip 0.5in
\begin{tabular}{|c|c|c|c|c|c|}
\hline
     Week &        Date   &     \# New &   \# Resolved &     Diff&    \#
Unresolved \\
\hline
       1  &    4-25-1999  &        6  &        0     &       6 &
6 \\
       2  &     5-2-1999  &        4  &        0     &       4 &
10 \\
       3  &     5-9-1999  &        3  &        0     &       3 &
13 \\
       4  &    5-16-1999  &       12  &       20     &      -8 &
5 \\
       5  &    5-23-1999  &       19  &       10     &       9 &
14 \\
       6  &    5-30-1999  &       22  &       14     &       8 &
22 \\
       7  &     6-6-1999  &       26  &       10     &      16 &
38 \\
       8  &    6-13-1999  &        4  &        5     &      -1 &
37 \\
       9  &    6-20-1999  &        6  &        7     &      -1 &
36 \\
      10  &    6-27-1999  &       19  &       19     &       0 &
36 \\
      11  &     7-4-1999  &       28  &       21     &       7 &
43 \\
      12  &    7-11-1999  &       17  &       26     &      -9 &
34 \\
      13  &    7-18-1999  &        5  &       10     &      -5 &
29 \\
      14  &    7-25-1999  &        3  &        8     &      -5 &
24 \\
\hline
\end{tabular}
\vskip 0.5in

The decision to hold off on the planned late June release was based in
part on these statistics and on a close perusal of the nature of the
most severe defects.

The key remaining area of general concern identified in the last
quarterly report was the robustness of the system to user error.  The
robustness issues have been addressed and a number of key bugs
identified and fixed. The operational stability of Glish alone is
excellent, and the stability of Glish plus the AIPS++ capabilities is
now acceptable, although continuing improvement is still expected.

In documentation, we have made moderate progress:

\begin{description}
\item[Checking of the Reference Manual]: Material in the reference
manual has been partially checked for completeness and correctness,
but more remains to be done.
\item[Completion of the Revised {\em Getting Started in AIPS++}]:
This is now done.
\item[Development of a cookbook]: Individual chapters have 
and are being written.
\item[Provision of processing recipes]: The provision of 
processing recipes has not occured on the time-scale that
we hoped. We will probably devote some Project Staff to this
endeavor in the next quarter.
\item[Provision of a FAQ]: The FAQ is in place and being extended
as necessary.
\item[Updating of the Glossary]: The glossay has been updated
to reflect new usages.
\end{description}

Our major new general concern is the memory usage of the system. In a
typical use of the system, the memory usage is between 110 and 120
MBytes, including about the components of Glish (about 40 Mbytes), and
about 8 servers. This usage is about twice what we had projected and
planned for. We have done a lot of testing and development to
determine the cause of this extra usage. The use of shared libraries
did not save much, and now we are experimenting with a mega-server
approach.  We expect that with some work like this, we can save
between 30 and 60 MBytes, thus bringing the number down to a
reasonable number. This work can be completed before the actual
release but the pre-release will be affected by this excessive memory
use.

The production of the CDROM master is now well in hand. In retrospect,
it is clear that we underestimated the time needed to iron out the
production difficulties. Testing on a variety of platforms takes time,
especially when CDROMS must be FedEx'ed around the world. However, we
have now have learnt how to make a distribution for each of Linux and
Solaris that can be run from the CDROM or installed to disk and run
from there. This knowledge should remain current for future releases.

Finally, we have started to prototype some of the operational systems
needed. We have considered how to set up helpdesk capabilities in a
what that will spread the load around and will not overwhelm the
Project Center with the handling of issues that are better addressed
locally.  Handling of questions and defects is a priority for the
whole Project. In the first instance, a defect or question will be
handled by on a local and regional basis ({\em i.e.}  North America,
Europe, and Australia). Once a defect is determined to be a true error
and not {\em e.g.} a configuration or user error, it is forwarded to
the Project Center for allocation to a Project Staff member for
repair. Similarly, questions beyond the competence of the regional
center will be referred to the Project Center. All of these steps are
handled using the NRAO installation of ClearDDTS.

\section{Upcoming issues}

There is an increasing interest on the part of various groups to
collaborate with AIPS++. As reported last quarter, we expect MERLIN to
apply formally to join the consortium. In addition, there is a
possibility that a collaboration between MERLIN and two groups at the
MPIfR-Bonn will acquire about 4 programmers to work with AIPS++.
Currently we are collaborating informally with USNO/NPOI, and
with JCMT.

From these examples, and other expressions of interest that we have
heard, it is to be expected that the number of groups actively using
AIPS++ will expand over the next few years. It is therefore important
that the consortium plan any expansion carefully. The following
questions will have to be answered:

\begin{enumerate}
\item On what basis are new members of the consortium to be accepted?
\item How should we manage the contributions and requests of non-consortium
members? 
\item How should an expanded consortium be managed? The current model
of a small Executive Committee overseeing a stream-lined Project
Management may not work in future.
\end{enumerate}

\section{Developments planned for 1999 Q3}

Most work for the next quarter will be focused on the release.
After the post-release, we expect to update our plans for the
next 6 month period. Here we give details of work for which
plans are already clear.

In {\bf Single Dish}, work for quarter three will focus on imaging and
calibration.  A single dish tool capable of gridding on-the-fly data
will be developed in July.  An initial single dish calibration tool
will be developed shortly thereafter.  Work within dish will consist
of rounding out the functionality to include multi-component fitting
and to allow direct interaction with an aips++ MeasurementSet.

Support for GBT will expand to begin to provide near real-time
display of GBT data.  The purpose of these types of display is to
give a quick look at the data so that data quality can be monitored.
A second holography backend test is scheduled for July.
The format for the spectrometer backend will be finalized and
the classes written to allow the filler to fill GBT spectrometer
data.

In {\bf Synthesis}, we will retain the overall strategy outlined
above. A particular focus will be on outreach to the scientific
community, through internal workshops and external visits, and in
ensuring that the first release is widely used and accepted by the
scientific community. Greater emphasis will be placed on VLBI
applications after the first release, in completing thin-path
capabilities, such as uv-data visualization, and in finalizing and
expanding automated data reduction scripts and utilities.

\section{Appendix: ATNF contribution {\em  Neil Killeen}}

\subsection{General}

This quarter the ATNF had 3 people working in AIPS++.  These are Neil
Killeen (75\%, also local manager), Mark Wieringa (30\%) and David
Barnes (75\%).  Wim Brouw continued on secondment to ATOMS and spent
only about 20\% of his time on aips++ this quarter.  Wim will return
at a higher level to aips++ in the next quarter.


\subsection{System}

Our systems (Solaris/Gnu, Solaris/egcs, and Digital Unix/egcs) were
generally stable this quarter.  We now have a running Linux/egcs system
as well at Epping.  We have some remaining file locking problems that we
hope will be solved by an operating system upgrade. 

\subsection{Visits}

Peter Barnes spent a few days here whilst on leave.  He and David
discussed the Display Library with which Peter is working. 

Frederic Badia from ATNF/Narrabri spent a week working with Neil and David
on protyping some ATCA online imaging components with aips++.

\subsection{Testing}

Two people are doing testing for us (as well as the aips++
group itself).  They are working with dish, images and
quanta/measures documentation.

\subsection{Individuals}

\subsubsection{David Barnes}

David's time was spent on

\begin{itemize}

\item Provision of PostScript and X11 Pixmap output from the viewer
  application.

\item Documenting the programmer interface to the viewer application.
  This is nearly complete.

\item Fixing a few bugs in the viewer.

\item Reporting more than 30 defects through testing of the system.

\item preparing for and installing a development AIPS++ system on 
  our new linux server.

\end{itemize}


\subsubsection{Wim Brouw}

Wim has spent his time this quarter on:

\begin{itemize}

\item QAG related tests of software and bug testing

\item reloading of IERS related data

\item loading spectral line data base

\item minor requested changes to a.o. super galactic coordinates; IERS predict
  renewal

\item documentation quanta and measures revision based on referee questions

\item FITS and units

\item AIX and SUN\_NATIVE dependencies; cxx2html; Linux problems etc

\item purify problems for egcs

\item discussion about fuzzy calculations and secure shells


\end{itemize}



\subsubsection{Neil Killeen}

Neil's time went on:

\begin{itemize}
\item convolution functionality from image tool

\item improved custom GUIs for image tool

\item improve documentation and testing of images and regionmanager

\item improve imagefitter performance (quality of fits mainly).  This involved work
  from user interface to support classes

\item defect fixes in images, lattices, FITS, coordinates

\item testing various components of the system and submitting defect reports

\item many email discussions

\end{itemize}




\subsubsection{Mark Wieringa}

This quarter Mark's time has been spent on

\begin{itemize}

\item MS v2.0 - test programs, debugging, TableMeasures integration

\item work on UVFITS filler

\item gDisplay.so - system work to get it going on the alpha

\item rpfitsfiller - merged back with on-line version, fixed gui input, code cleanup

\item bug fixes for release: msDO - problems with source IDs, documentation,
filler, added max \# row argument for iterator.

\item on-line imaging - discussions with Fred, and some debugging.

\end{itemize}


\section{Appendix: BIMA/NCSA contribution {\em  Doug Roberts}}

\subsection{Display Library PostScript Driver}

Harold Ravlin has put in some bug fixed into the PostScript driver for 
the Display Library as well as a few additional features.  Harold added
bounding box calculations for EPS, but these do not do the work the
way they should, since the one of the first things the Display Library
does is to clear the entire writable area.  Harold did bug fixes on
plotting dashdash lines.  Harold looked into a request to be able to
select that paper tray, however, this is a device dependent feature
and could make a file difficult to print on other printers; Ravlin
will look further into this after the release.  The PostScript driver
no longer appends a showpage when in EPS mode; it had been doing this
as a debugging aid.  Harold began looking into adding support for
registering colormaps.

\subsection{Aipsview}

Aipsview will now print a warning message and exit if the default
visual is not PseudoColor.  Numerous changes were made to Aipsview so
it will look for a PseudoColor visual if the default is something
else. It appears to work as long as Aipsview is running on an
SGI. There may be problems with Sun's implementation of Motif. The
changes will be checked in after the release.  Harold added missing
copyrights and RCS information to some files. 

\subsection{System}

Ravlin and Roberts spent time to work on various check in and check
out problems and how the remote commands rai/rao/rau worked under
secure shell.

\section{Appendix: NFRA contribution {\em  Ger van Diepen}}


\subsection{General}

Due to the pressures of the WSRT upgrade, the NFRA contribution to the
global AIPS++ package has regrettably sunk to an absolute minimum. The
only substantial contribution was made by Ger van Diepen in his role
as Technical Leader, and as the man responsible for a number of
essential infrastructure modules (tables, lattices, ms2uvfits, etc).

This quarter, the NFRA contribution in FTE's was as follows:
\vskip 0.5in
\begin{tabular}{|l|c|c|c|}
\hline
&                     Global AIPS++  &    Local AIPS++  &  Other \\
\hline
Ger van Diepen:  &    50\%           &       30\%       &  20\%    \\
Jan Noordam:   &      10\%           &       60\%       &  30\%    \\
\hline
\end{tabular}
\vskip 0.5in

\section{Appendix: NRAO contribution {\em  Tim Cornwell}}

The core NRAO AIPS++ group is now Barnes (100\%), Braatz (100\%),
Cornwell (100\%), Garwood (100\%), Golap (100\%), Holdaway(100\%),
Kemball(100\%), Marson (100\%), McMullin (100\%), Schiebel (100\%),
Weatherall (50\%), and Young (100\%, Golap and Young are funded by
NCSA-NRAO collaboration.

%\subsection{Hardware}

\subsection{Individuals}

\subsubsection{Peter Barnes}

Completed initial implementation of Table/MS side of Display Library
inheritance tree, concluding this stage of the DL work from the last
quarter.  This consisted of writing 8 classes within the tree: 4
DisplayDatas classes (TablePADD, TableAsXYPlot, MSAsXYPlot,
MSAsTimeSeries) and the corresponding DisplayMethods classes; and a
utility class, MSPlotter.  This latter was a new class, and
considerable effort was made to ensure adherence to the ideals of the
PlotStyles design document, despite the fact that it is only a
temporary class until the MultiWorldCanvasHolder is implemented.  Most
of the time was spent defining the proper parameter handling for the
required plot functionality.  Also wrote a test program for this class.

Began duties as AOC Solaris site manager.  Checking sneeze logs,
mostly, with other small duties as they arise.

Made extensive use of assay and imagetest, esp during May when Anantha
was away.  Most of these results were reported to aips2-lib.  Began
testing visplot, ms, \& pgplotter, generating bug reports.

Attended Imaging '99 Conference in Tucson, June 6-9.
  
\subsubsection{Jim Braatz}

Jim Braatz was hired on to the AIPS++ project in April 1999.  His primary
responsibilities are to manage the Green Bank installation of AIPS++, to
support GBT operations, and to work on single-dish applications.
Specifically, Jim worked on the following tasks during this quarter:

\begin{itemize}
\item Became acquainted with glish and the AIPS++ system, so far concentrating
  on the environment for software development.  Learned the software
  distribution methods, and began to take on the responsibility as the
  Green Bank AIPS++ manager, which involved learning the inhale process.
  Visited Charlotteville on a roughly bi-weekly basis to work with
  the AIPS++ team there, as part of this training.  Updated the Glish
  User's Manual.

\item Developed some glish routines for reading and performing simple analysis
  on data from the holography receiver.  Participated in the holography
  tests in April 1999.

\item Became acquanted with the gbtlogview software, which is an engineering tool
  used to view data from GBT-related devices.  Started to rewrite the plugins
  to this tool to accomodate some requests from engineering.

\item Learned the Green Bank tipper software, transferred the programs, data
  files and web pages to my own account, and made several improvements and
  bug fixes to that software, primarily related to the display of the 
  opacity results.

\item Observatory-related duties this quarter included:
\begin{itemize}
\item   Coordinator of the weekly Science Lunch Talk in Green Bank
\item   Mentor for the Chatauqua week-long undergraduate program
\item   Lecture on "The Age of the Universe - A New Controversy" to the RARE CATS Institute
\item   Mentor to an NRAO REU summer student, who is working on VLA data
\item   Public Radio appearance on West Virginia Mountain Radio, discussing 
     radio studies of supermassive black holes
\end{itemize}
   
\end{itemize}
  
\subsubsection{Tim Cornwell}

In addition to his project management duties, Tim spent a large amount
of time on managing, assigning and fixing defects. He worked with
Kumar Golap on the provision of wide-field imaging capabilities within
AIPS++, and with Mark Holdaway on Mosaicing, and Multi-scale clean.

\subsubsection{Bob Garwood}
Bob Garwood's primary responsibility is to oversee and contribute
towards the single dish work in aips++.  This work remains focussed on the
"dish" environment and the support of the GBT.  His contribution
in support of the GBT is primarily through the GBT fillers which
convert the GBT FITS data files to an aips++ MeasurementSet.

Over the past 3 months (April through June) he has done the
following:

\begin{itemize}
\item        Resolved several bugs involving dish and the FITS classes.

\item        Improved (with Joe McMullin) the documentation for dish, including 
        better documentation for the graphical and command line interfaces.  
        The single dish FITS binary table convention is now documented as 
        part of the dish documentation.

\item        Made a quick review of all existing GUIs in order to assess how well
        they follow the GUI guidelines.  Bugs have been submitted for
        all of the GUIs which do not follow these guidelines.

\item        Prepared the GBT filler for use with the holography backend in time
        for a planned test of that backend.  Bugs and shortcomings of the
        filler found during this test are being fixed.  A revised GBT
        filler will be available for another planned holography test in
        July.

\item        Worked with the GBT staff on several iterations of the format for
        the GBT spectrometer backend FITS files.

\item        Visited Green Bank several times.
\end{itemize}

\subsubsection{Kumar Golap}

In April, Kumar joined NRAO as a Visiting Scientist for 2 years. He
has worked with Tim Cornwell on testing of the wide-field imaging
capabilities within AIPS++. He is also responsible for testing of
the SGI port of AIPS++.

\subsubsection{Mark Holdaway}


Mark Holdaway has been working on mosaicing integration in the imager
tool.  The voltage pattern editor vpmanager, which allows the user to
create a wide variety of voltage pattern and primary beam models, is
complete.  The table of voltage patterns it generates can be read by
imager, which uses the voltage patterns when mosaicing.  Default
voltage patterns for each telescope and observing band can be used
without accessing the vpmanager.  (Presently, only data with a single
voltage pattern can be correctly imaged with imager due to a current
limitation in the visibility iterator.)

Mosaicing has required many minor changes to the SkyEquation/SkyModel
design.  One of the most important has been "Sault weighting".  In
Sault weighting, the noise does not flare up at the edge of the mosaic
sensitivity pattern.  However, under Sault weighting, the flux scale
is not constant over the image, and inverse-Sault weighting must be
applied for some operations.

It is very imortant for mosaicing observations to accurately image
extended structure.  While maximum entropy is generally the algorithm
of choice for imaging extended structure, it's positivity bias often
results in an overestimation of the flux, and its residuals are
correlated with the true emission.  The multi-scale CLEAN algorithm is
the answer to this problem, succesfully imaging weak extended
structure with noise-like residuals.  Mark has integrated the
multi-scale CLEAN algorithm into imager, including masking.
(This is not checked into the system due to the "code freeze"
prior to the public release.)

Another aspect of mosaicing integration is modifying the SkyEquation
to perform minimum sized FFT's.  When minimum-sized FFT's are in
place, the cpu time required for a mosaic will vary approximately
linearly with the number of pointings, a great improvement over the
current code.  Mark has done exploratory work on how to implement
minimum-sized FFT's via the MFSkyEquation (multi-field sky equation)
class.

On July 14-16, AIPS++ and the ALMA project cosponsored a Mosaicing
Techniques Meeting in Socorro, New Mexico.  Mark was the main
organizer for this meeting, bringing about 30 researchers together for
discussions on new mosaicing algorithms and the problems they are
having with the existing algorithms.  

\subsubsection{Athol Kemball}

Athol's work is described above, principally in the section on
Synthesis development.

\subsubsection{Ralph Marson}

In the last quarter Ralph has worked on:

\begin{itemize}
\item VLA filler
\begin{itemize}
\item  verify that the data is correct
\item  speed it up so that it is limited by I/O rates.
\item  improved to read multiple tape files
\item  improved to read only data from selected project id's
\item  removed the channel zero data
\item  rewrote the vlatoms glish code to work asynchronously.
\end{itemize}
\item Lattice convolver
\begin{itemize}
\item investigated the speed and memory performance of this class and
    improved the latter lots and the former a bit.
\end{itemize}
\item testing: 14 bugs submitted     
\item bug fixing: ten bugs resolved,  six postpones, one active.
\end{itemize}


\subsubsection{Joseph McMullin}

Joseph McMullin's main responsibilities are work on single dish 
applications within AIPS++.

\begin{itemize}

\item General Improvements and bug fixes to DISH: 
Improvements to DISH plotter. Standardized buttons, added popup help, fixed
resizing problems, improved state saving. Implemented a focus parameter so that
the active plot will have its statistics represented. Modified behavior of 
baseline operation to default to an overplot of the fit despite the state of
the overlay. Fixed plotting of multi-channel data. Implemented a better 
clear overlays to fix the lapse in the color table. Fixed initial character
height setting. Reworked some header information to make overwrites very rare.
Added a command line gaussian fit operation to DISH.
Added "Print to File", "Print to Clipboard" utilities for statistics operation.
Improved Write spectrum operation so that the default file is tagged by the
scan number (essential for script loops).
Added JPL line list data to DISH for quick line identifications. It currently
only knows about 2 types of axes (frequency and velocity).

\item Wrote (with Bob Garwood) documentation for DISH (Note 225).

\item Generated a list of realtime display targets for Green Bank (with Jim Braatz
and Bob Garwood).

\item Fixed bugs 1208, 1015.

\item Added a plugin file for the pgplotter to allow the old gplot1d commands to
be executed. This enabled old scripts to be used with only minor changes and
provides added functionality for Green Bank engineering use.

\item Used new viewer and images module to reduce BIMA images for paper. Submitted
to ApJ.

\item Presented AIPS++ demos at the AAS in Chicago, IL. Wrote automated scripts for
demonstrating DISH functionality (reduces NRAO 12-m data).

\item Developed an imaging tool for DISH. Three tools were developed initially: 1)
lookmap which reads a working set and plots the spatial locations of where
data were obtained (and optionally the scan numbers) relative to a selected
scan, 2) spectramap which reads a working set and plots the spectra at
the spatial locations where they were obtained relative to a selected scan,
3) contourmap which reads a working set and plots conours of integrated 
intensity for a selected velocity range (currently this is quite limited as
it does no interpolations).
\end{itemize}

\subsubsection{Darrell Schiebel}

Darrell performed the work on Glish described above.

In addition to the Glish development accomplished during the quarter, Darrell
got shared library builds of AIPS++ libraries working, and he worked to get
things in order for producing AIPS++ CDROMs.

\subsubsection{Kate Weatherall}

\begin{itemize}
\item	Worked on dangling links and fixed incorrect or broken paths to documentation
\end{itemize}	

\subsubsection{Wes Young}

\begin{itemize}


\item Installed ClearDDTs and upgraded it once.  
\item Updated existing html pages to use ClearDDTs.

\item General ClearDDTs adminstation - creating projects - assigning
permissions - finding lost bugs

\item Fixed some defects assigned to me.

\item Added copyright to generated docs files

\item Provided SGI support for AIPS++. 
\begin{itemize}
\item Identify and fix (or send them to someone who could) build problems
\item Helped identify where the IO is done in AIPS++ so we could instrument it
    with Pablo
\item Looked at the core dump on exit problem
\item Baseline testing of imager
\end{itemize}

\item Provided documentation support.
\begin{itemize}
\item Split the reference manual in to "package sized" books
\item Added a makefile target for resolving links
\item Abortive attempt to add javascript based menus to web pages
\end{itemize}

\item Added a registration feature so user can register aips++
\end{itemize}

\section{Appendix: Summary of AIPS++ Personnel Changes}

[Preliminary]

In this section, we give the names of people in the various AIPS++
groups and the nominal fraction of time allocated to AIPS++.

The ATNF group is: Neil Killeen (75\%, also local manager), Wim
Brouw (20\%), Mark Wieringa (30\%) and David Barnes (75\%).

The BIMA/NCSA group is: Harold Ravlin (20\%), Doug Roberts (50\%).

The NRAO group is: Tim Cornwell (100\%), Jim Braatz (100\%), Bob
Garwood (100\%), Kumar Golap(100\%), Mark Holdaway (100\%), Athol
Kemball (100\%), Ralph Marson (100\%), Joe McMullin (100\%), Darrell
Schiebel (100\%), Kate Weatherall (50\%) and Wes Young (100\%).  Young
and Golap are funded by NCSA-NRAO collaboration.

Thus, in aggregate, we have 18 people contributing about 13 FTEs to
the AIPS++ Project. Of these, 9 are employed by NRAO (one funded by
NCSA), and contribute 8.5 FTEs. The numbers for the other partners
are: ATNF 4 and 2, BIMA/NCSA 2 and 0.55, NFRA 3 and 1.9. There are
three unfilled positions, one at BIMA/NCSA (Briggs) and two at NRAO
(GBT scientist, parallel applications developer: Kumar Golap due in
April).

\end{document}

