\documentclass[11pt]{article}
\usepackage{html, epsf}

\newcommand{\searchURL}{../../html/aips2search.html}
\newcommand{\aipsviewURL}{../../user/aipsview/aipsview.html}
\newcommand{\GlishmanualURL}{../../reference/Glish/Glish.html}
\newcommand{\devplanURL}{../../notes/226/226.html}
\newcommand{\GlishtutorialURL}{../../notes/195/195.html}
\newcommand{\gsURL}{../../user/gettingstarted/gettingstarted.html}
\newcommand{\cookbookURL}{../../cookbook/cookbook/cookbook.html}
\newcommand{\userrefmanualURL}{../../user/Refman/Refman.html}
\newcommand{\specsURL}{../../specs/specs.html}
\newcommand{\betaURL}{../beta/beta.html}
\newcommand{\consortiumURL}{../consortium.html}
%
\begin{document}
\title{\bf AIPS++ Quarterly Report: 1999 Q3}
\author{T.J. Cornwell, NRAO}
\date{1999 November 23}

\maketitle

\tableofcontents

\section{Introduction}

This quarter, we have focused almost exclusively on preparing the
first release. This included testing the system, fixing defects,
issuing two patches, preparing and testing the actual CDROM release,
and distributing the final release. The official release date was
October 4, and was made at the ADASS meeting in Hawaii.

We are now attempting to establish a 6 month plan-develop-test-release
cycle. To start the next cycle, the development plan for the
next release (the next one: 1.3 targeted for April 2000) was
revised, and targets assigned.

\section{Developments in 1999 Q3}

In {\bf Single Dish support}, work was slowed this quarter due to
absence of Joe McMullin from the group for over two months.  The work
done by Jim Braatz, Joe McMullin and Bob Garwood continues to be
focussed on support of the GBT and development of the DISH environment
for general single dish analysis in AIPS++.  There continue to be
regular visits by Garwood and McMullin to Green Bank and Braatz to
Charlottesville.

Jim Braatz spent much of this quarter learning his way around
AIPS++ and learning about his Green Bank specific duties from Joe
McMullin.  In particular, the tipper software was transferred from
McMullin to Braatz as was the pointing software.  Braatz and McMullin
continue to work on transferring responsibility for the commissioning
tools.

There was a second test of the GBT holography backend this quarter.
Garwood fixed several defects reported in the GBT filler after the
first test in preparation for this test.  In addition, Braatz
developed AIPS++ scripts to verify the integrity of the data during
these test.  The tests went well as far as AIPS++ was concerned
although as of that test, the majority of the post-observation data
reduction was done in UniPOPS.  Braatz has worked on converting the
UniPOPS scripts to AIPS++ and the results of this work (which is
nearly complete) was presented at ADASS in October.  Garwood continues
to work to enhance the GBT filler.  In addition to the fixed defects
noted above, the filler was turned into a true AIPS++ distributed
object in preparation for its use in an on-line setting.  The
remaining Green Bank work was concentrated on weaning a few engineers
off of a very old installation of AIPS++.  The interface to several
utilities, especially plotting, has drifted somewhat over the past
year such that these older scripts did not work with the current
installation.  This work has been completed and the older installation
removed from Green Bank.  Finally, the format for the GBT spectrometer
backend FITS files was finalized.

The non-GBT focus this quarter was primarily on responding to reported
defects and enhancement requests from users of DISH in the AIPS++ 
pre-release.  The current state of DISH was summarized in a poster
presented at ADASS in October.

Work continued on single dish imaging tools.  NRAO 12m On-the-fly data
was successfully converted from UniPOPS to an AIPS++ MeasurementSet.
This data set was used by Tim Cornwell in the first successful imaging
of single dish OTF data in AIPS++.  These results were presented at
the October ADASS meeting.

A poster summarizing the single dish FITS binary table convention was
prepared this quarter for presentation at ADASS.

In {\bf Synthesis support}, this quarter is unique in that it has been
significantly concerned with preparations for the first release which
was issued on October 4. As a result, general synthesis development
was tapered off over the quarter in favor of CDROM testing, the
identification and correction of code defects, the preparation of
high-level documentation and an associated public outreach effort to
scientific users. These efforts have been successful and represent a
significant transition in the focus of the project, and the
operational status of the synthesis code in particular. This is in
keeping with previous planning in this area in which this was
identified as a key objective.  We have reached more scientific users
in this quarter, and this focus will continue as a high priority for
the foreseeable future. Synthesis development has continued in
targeted areas in this quarter, however, as discussed below, and will
resume aggressively at the start of the current quarter to meet
planned requirements for the second release.

Release preparations were extensive and encompassed careful testing
by developers, and an increasing number of outside users. The latter
component is expected to increase from this point on, but in keeping
with the general project philosophy, internal testing will remain a
clear responsibility and priority of the project. All identified
defects were filed in the standard ClearDDTS defect tracking system,
with associated defect severities, and were addressed with special
focus on the severity one and two categories. This system has worked
well in synthesis, as in other areas, to track and provide user
feedback on defect status. AIPS++ has unique capabilities to provide
automated testing due to the powerful command line (Glish) and the
high-level connection between Glish and the underlying C++ code. The
automated test scripts have been reviewed and their completeness
assessed as part of an effort to revitalize and modernize this
capability. This will enhance daily feedback on the synthesis code
correctness, as well as improve the speed of release and release patch
testing.

The public outreach campaign to reach new scientific users was
started in this quarter at the AOC in Socorro, and constituted six,
hour-long tutorials conducted once per week. Two focused exclusively
on synthesis capabilities in imaging and calibration. As an adjunct,
the first edition of the cookbook was completed, which will be the
highest-level documentation for scientific users. It augments the
already extensive user reference manual, but does not replicate the
lower-level material. It is indexed on scientific data reduction
activities, and contains synthesis chapters on general imaging
(T. Cornwell), wide-field imaging (K. Golap and T. Cornwell), and
synthesis calibration (A. Kemball), in addition to important chapters
in other areas. We expect to expand the cookbook significantly in
response to user feedback, and to add more material for the second
release. The public outreach campaign will be taken to Europe in
November (Cornwell and Kemball), as well as other consortium sites in
the US and elsewhere in the short term. Our intent is to publicize
general capabilities, including synthesis, to potential scientific
users as widely as possible and for all consortium instruments, and we
expect this to continue to some time as a high priority to introduce
new users to the system.

The general synthesis development which has been undertaken during
this quarter has been in areas consistent with previous planning
priorities, namely: i) the development of "thin-path" capabilities for
as complete end-to-end connected-element reduction as possible; and,
ii) advanced synthesis development in targeted areas to finalize key
interfaces.

Thin-path development was focused in the areas of data fillers and
uv-data visualization using the Display Library framework. The VLA
data filler was advanced significantly during this period (R. Marson);
more advanced capabilities will be added during the current
quarter. The ATCA filler has also been revised to add new
features. The uv visualization work in the Display Library moved more
slowly, but initial code and a design review has been completed. This
will move more aggressively in the current quarter.  Basic editing
capabilities were added to the {\tt visplot} tool by T. Cornwell, as
well as general optimization. Work on the outstanding code required
for full adoption of the MeasurementSet v2.0 format moved ahead,
beyond the {\tt ms} tool and into the synthesis MS access code
(M. Wieringa). We plan final implementation in the full package early
in this development cycle, well in advance of the second release.

A mosaicing techniques meeting arranged by M. Holdaway, T. Cornwell
(AIPS++) and M. Rupen (for ALMA) held in July was a significant
success, with important contributions by many participants. Work in
mosaicing has continued in this quarter (M. Holdaway and T. Cornwell)
in the areas of multi-resolution CLEAN and MEM deconvolution for
mosaicing. In addition, work has been undertaken in optimized FFT
gridding and primary beam weighting for mosaicing. Planning for
simulation work in support of ALMA has advanced and will be a new
focus for M. Holdaway once some final mosaicing capabilities have been
completed. K. Golap has tested the wide-field imaging capabilities
extensively in this quarter, using both real low-frequency VLA data
and simulated data. This has primarily involved use of the {\tt
dragon} tool for wide-field imaging, but support utilities to measure
the positions of simulated image components have also been developed
in support of this testing effort.  K. Golap and A. Kemball are now
parallelizing the wide-field imaging capabilities as part of the
parallelization effort with NCSA. This holds significant promise as
these are very computationally challenging problems for all realistic
cases. A basic on-the-fly (OTF) gridding capability has been added to
{\tt imager} (T. Cornwell and B. Garwood) in the same framework as
synthesis gridding, in keeping with the overall philosophy of combined
synthesis and single-dish processing in AIPS++. This OTF effort is
part of the strategy of finalizing key high-level code interfaces. A
related effort concerns planning for image-plane solvers (K. Golap and
T. Cornwell), a final interface which will be established by
prototyping.

Synthesis calibration has been extended in this quarter through the
addition of data selection and display capabilities, the improvement
of basic infrastructure, and through the provision of flux
bootstrapping utilities (A. Kemball). The cross-calibration
capabilities are in place, as planned for the first release, but there
is still work to be done expanding capabilities in this area, and it
will be undertaken in the current quarter.
 
NFRA has continued work on utilities in support of WSRT/TMS
commissioning (J. Noordam and G. van Diepen). Some assistance has been
provided with debugging the transfer of data through AIPS++ by
A. Kemball.

Parallelization and high-performance computing, a collaborative
effort between NCSA and NRAO (A. Kemball and D. Roberts), has a close
connection with synthesis development, in the areas of wide-field
imaging, mosaicing and the planned image-plane solvers. This effort is
summarized elsewhere in this report.

A new planning cycle has been completed in this quarter to define the
synthesis development targets for the second release. The primary
focus will lie in expanding the connected-element thin-path in areas
in which it is weak at present (uv-display and editing), and in
providing a basic VLBI capability (data filler and
fringe-fitting). Automated reduction and imaging pipelines will also
be a key focus in this development cycle.

In {\bf Glish}, no substantial changes were made.

In {\bf Parallelization}, work continued consistent with the previous
planning in this area. The broad goals of this effort are to
demonstrate, and make scientifically available, a capability to
process the most challenging data reduction problems in radio
astronomy, and to introduce the parallelization infrastructure in
AIPS++ to make these high-level applications possible.

A continuing infrastructure development area which received attention
in this quarter has been a research effort into the use of parallel
I/O in AIPS++.  In the first two quarters of 1999, a graduate student
with the computer science department at the UIUC added I/O
instrumentation calls to the AIPS++ library.  All reads, writes,
opens, closes, and seeks have been instrumented.  During execution,
the counts and times of the I/O calls are written to a trace file in
Self-Defining Data Format.  The Pablo analysis tools are used to
investigate the resulting trace file.  D. Roberts proposed and was
granted access to an NCSA Origin 2000 before it was introduced into
production.  The parallel group had access to the machine for 10 days,
during which 8 experiments of various sizes were performed.  In
preparation for this experiment, D. Roberts and W. Young worked on
validating correct execution of pimager (the MPI-based, parallel
version of imager) on large images.  Several problems were identified
including a 1 GB limit on memory allocation with 32-bit programs under
IRIX.  After validating the correct image generation, D. Roberts
carried out the experiments and analyzed the resulting I/O trace
files.  The results were that several files were opened and not
closed.  More importantly, for large images, I/O to the TempLattice
file that was created in the imaging process dominated the total I/O
by a large factor.  For runs that created 2 GB images, the total I/O
to the TempLattice was 500 GB, with frequent small size I/O requests.
The parallel group is looking into optimizations to address this
performance bottleneck.

After the I/O experiment W. Young visited NCSA and worked on porting
AIPS++ to the newest generation of the SGI MIPSpro compilers (7.3).
This effort has been completed. The 7.3 compiler has the newly defined
OpenMP binding for C++.  D. Roberts started a project to parallelize
GridFT using OpenMP and the new compiler.  This should be complete in
the next quarter.

After porting AIPS++ to MIPSpro 7.3, D. Roberts and W. Young
investigated what would be necessary to port AIPS++ to a 64-bit
compiler in order to address the 1 GB memory limit discussed above.
W. Young is summarizing the specific required changes to the code and
will submit the details to the project in October for consideration of
their broader impact. W. Young has been using a code verification and
memory tracking utility, Insure++, in this work.

D. Roberts did some site administration duties and installed AIPS++ on a
new Linux system in order to be able to test Linux changes on a local
machine.  During a visit to NCSA, A. Kemball and D. Roberts worked on
a document specifying NCSA computational requirements and revision of
the plan for parallelization and optimization for the next 6 months.

An initiative in the parallelization of wide-field imaging within
AIPS++ has been started by A. Kemball and K. Golap. The existing
parallelization infrastructure has been refined to address some
requirements of this initiative as part of this work. The wide-field
parallel capability is expected to be available in this
quarter. Performance profiling of the existing serial wide-field
imaging capability was completed this quarter.

The port of a snapshot of AIPS++ to Windows NT has proceeded as part
of the MS research project of P. Cortes under supervision of
A. Kemball at NRAO, and a reasonable fraction of the {\tt aips}
library has been compiled under Microsoft Visual C++ in this
quarter. The C++ syntax differences are being tracked between the NT
and UNIX compilers, along with differences in template instantiation
and related issues. The current plan calls for completing the NT port
for the {\tt aips} and {\tt trial} computational libraries, and then
running parallel applications on the NT super-cluster at NCSA. Porting
of Glish and the Display Library is deferred until later. It should be
emphasized that this is not a final port of AIPS++, but instead the
trial port of a snapshot. Part of this work will be a report detailing
the implications of a full port. The related work by P. Cortes
concerns the performance of AIPS++ parallel applications on cluster,
or loosely coupled, parallel architectures. We have obtained an
allocation of time on the Linux super-cluster of the Albuquerque High
Performance Computing Center as part of this work, and have started to
install AIPS++ on this system. We expect I/O and memory access to be
significantly different in this environment as compared to more
tightly-coupled architectures (eg. SGI), and the effect this will have
on radio astronomy reduction algorithms in part of this research
effort. Clusters are likely to play a strong role in radio astronomy
reduction in future years.

Work continues in ensuring build stability between the SGI AIPS++
versions at the AOC and NCSA. We have undertaken system administration
work to synchronize operating systems and compiler versions between
the two sites in this quarter, as in previous quarters. The SGI build
stability has improved significantly as a result of these and related
measures.

In {\bf Measures}, no substantial changes were made. One amusing
exception was that ALMA was added to the database of observatory
locations.

In {\bf AIPS++ Infrastructure}, most AIPS++ time has been spent on
testing the system and resolving defects. Little new development
has been done.  Defects were mainly resolved in the tools catalog, os,
and logger to make them adhere to the GUI guidelines, to have better
interaction between GUI and command line, and to add new options.  A
few defects in cxx2html have been fixed to make it process header
files where summary or class declaration is split over multiple lines.

In {\bf Image Analysis}, we fixed defects, considerably augmented the
test scripts, added some new capabilities (concatenation of images,
improved viewing), and improved the documentation.

In {\bf Documentation}, the first version of the Cookbook was
assembled by Athol Kemball, with contributions from Cornwell, Kumar
Golap, David Barnes, and Neil Killeen.

In {\bf Management}, at the end of the quarter, we revised and updated
the \htmladdnormallink{development plan}{\devplanURL}.  We intend to
settle down on a 6 month plan-develop-test-release (PDTR) cycle. The
time scale is as follows, counting from the date of the previous major
release.

\begin{description}
\item[+0 weeks] Revise planning documents
\item[+2 weeks] Finalize planning, assign targets
\item[+2 weeks] Start development, testing 
\item[+2 weeks] Initiate completion of testing of existing capabilities
\item[+4 months] Freeze new checkins, start dedicated testing
\item[+5 months] Hard freeze, uprev to next versions of release 
and develop masters.
\item[+5.5 months] Cut final CDs
\item[+6 months] Ship CDs
\end{description}

In {\bf Outreach}, we conducted a series of tutorials talks by Tim
Cornwell and Athol Kemball at the AOC. These were based on various
pieces of documentation, including ``Getting Started in AIPS++'' by
Anantharamaiah, and the AIPS++ Cookbook assembled by Athol Kemball.
These talks were well received with about 20-25 people attending each
of the 6 talks. Our first impressions of the response are:

\begin{itemize}
\item We learned to go slowly. Many aspects of the package
are subtle and need careful explanation. We settled on a successful
format in which one person demonstrated one area of the package
per talk, with a talk lasting 60 minutes. Use of an overhead
projector driven by a laptop computer helps everyone sit comfortably
and see the screen easily.
\item The series was spread out over 6 weeks mainly because of
our other heavy load at the time. It could in principle be
condensed into one week of one talk per day or even a few
days with a three talks per day.
\item The attendees basically welcomed the new package and looked 
forward to doing new science with it. The ability to do new
types of science with AIPS++ seemed to be apparent to many, though
not all, attendees.
\item The package is seen as hard-to-learn but the underlying power 
is very apparent.
\item The response to the revised user interfaces was generally
positive.  Lots of work remains in smoothing off rough edges in the
user interfaces, and we had many insightful and helpful comments 
about this aspect in particular.
\item As we knew, several key areas of functionality are still missing. In
particular, visibility editing remains a substantial problem and must
be addressed in the next development cycle.
\item The object-based approach has not yet been a large apparent
obstacle. We cannot detect whether our switch to avoid using
words like tool and method was a contributing factor.
\item Many attendees seemed to think that the scripting power of
the package would be very advantageous.
\item A number of attendees clearly grasped the importance of the
overall system design in future applications such as the VLA
upgrade project.
\item The tutorial series was a clear success and bridged some
of the historical divisions between the NRAO AIPS++ staff and
the NRAO scientific staff.
\end{itemize}

AIPS++ had a considerable presence at ADASS in Hawaii:

\begin{itemize}
\item The public release was made there on October 4. Those
project members attending celebrated in the evening.
\item AIPS++ was demonstrated using a laptop and a projector.
\item Darrell Schiebel gave an invited talk on the use of Glish for
Real-Time programming
\item Braatz, Cornwell, Garwood, and Roberts presented posters on
various aspects of AIPS++. Tony Willis of HIA/DRAO presented a poster
on ACSIS, in which AIPS++ plays a leading role.
\item A Birds-Of-a-Feather (BOF) session on AIPS++ was held.
About 40 people, mostly new faces, attended this session.
Discussion was excellent.
\end{itemize}

In {\bf System}, a lot of effort has been put into preparing the
release, especially by Darrell Schiebel and Tim Cornwell.  The system
tree has been split into a release and development tree.  Patches are
checked into the release system, which can only be done by Tim.  New
developments have been started in the development tree.  Most notable
are:

\begin{itemize}
\item Adoption of egcs 1.1.2 as the official project compiler.
\item Changes (by Wim Brouw) to get rid of typical gcc 2.7.2
  workarounds (like typedefs to make templates known and the
  ac() and at\_c() functions).
\item A change proposal to handle ifdef-s in a more structured way.
\end{itemize}

It appeared that the Rational products like purify do not support egcs
1.1.2 and it looks like it takes quite some time before it gets
supported. The beta versions of purify, etc. support egcs 1.0.3 to a
greater degree.  Therefore it has been decided that AIPS++ code will
add egcs 1.0.3 as a secondary compiler in the interim. Where needed,
macros or other constructs have to be used to make it compile.  In
Socorro it is investigated whether Insure++ is a possible replacement
for purify and purecoverage. A good replacement for quantify does not
seem to be readily available.

In the {\bf Quality Assurance Group}, we continued on a review of the
compliance with various rules. This resulted in a substantial number
of minor changes here and there in the system. We expect this work to
continue at a lower rate for the next development cycle.

In {\bf Visualization}, we concentrated on fixing defects and
improving documentation. NRAO was just awarded an NSF grant (from
Computing, Info Sci, Engineering) for visualization development in
AIPS++. This amounts to two positions for three years. This will
augment staffing at NCSA and ATNF to double the total effort.  The
goal of the proposal was to allow visualization and editing in two
different domains while transforming between the two using AIPS++
capabilities. We are currently working to fill these positions.

\section{First release}

In the preparation for the release, we issued two patches (binary
fixes) to the CDROM pre-release. These patches were in response to the
defect reports submitted by our testers.

In the previous quarterly report, the number of total defects was
still rising linearly with time. In this quarter, the number of
defects saturated, indicating either that the package was becoming
more robust and stable, or that our testers were exhausted! The common
experience amongst our testers (mostly project staff) affirms the
former, though the latter is probably also true.  The mean time
between failure of AIPS++ is typically a few hours.

\vskip 0.5in
\begin{tabular}{|c|c|c|c|c|c|}
\hline
     Week &        Date   &     \# New &   \# Resolved &     Diff&    \#
Unresolved \\
\hline
       1  &    4-25-1999 &         9   &       0   &         9    &
9 \\
       2  &     5-2-1999 &         5   &       2   &         3    &
12 \\
       3  &     5-9-1999 &         3   &       0   &         3    &
15 \\
       4  &    5-16-1999 &        15   &      22   &        -7    &
8 \\
       5  &    5-23-1999 &        35   &      11   &        24    &
32 \\
       6  &    5-30-1999 &        39   &      18   &        21    &
53 \\
       7  &     6-6-1999 &        41   &      11   &        30    &
83 \\
       8  &    6-13-1999 &        18   &       7   &        11    &
94 \\
       9  &    6-20-1999 &        21   &      11   &        10    &
104 \\
      10  &    6-27-1999 &        37   &      24   &        13    &
117 \\
      11  &     7-4-1999 &        60   &      35   &        25    &
142 \\
      12  &    7-11-1999 &        45   &      43   &         2    &
144 \\
      13  &    7-18-1999 &        18   &      25   &        -7    &
137 \\
      14  &    7-25-1999 &        53   &      35   &        18    &
155 \\
      15  &     8-1-1999 &        17   &      24   &        -7    &
148 \\
      16  &     8-8-1999 &         3   &       4   &        -1    &
147 \\
      17  &    8-15-1999 &        90   &      19   &        71    &
218 \\
      18  &    8-22-1999 &        20   &      11   &         9    &
227 \\
      19  &    8-29-1999 &        63   &      32   &        31    &
258 \\
      20  &     9-5-1999 &        36   &      30   &         6    &
265 \\
      21  &    9-12-1999 &        21   &      32   &       -11    &
254 \\
      22  &    9-19-1999 &        19   &      17   &         2    &
256 \\
      23  &    9-26-1999 &        28   &      15   &        13    &
269 \\
      24  &    10-3-1999 &        12   &       4   &         8    &
277 \\
      25  &   10-10-1999 &         6   &      10   &        -4    &
273 \\
      26  &   10-17-1999 &         6   &      12   &        -6    &
267 \\
      27  &   10-24-1999 &         0   &       1   &        -1    &
266 \\
\hline
\end{tabular}
\vskip 0.5in

The highest severity defects run about 10\% of the total. We expect to
act on those as quickly as possible, and the lower severity defects as
a background task consuming about 25\% of our available time.  As the
number of defects saturated, we became confident that the package was
ready for the first releease. The first release was made on Oct 4 at
the ADASS meeting in Hawaii.  Approximately 225 CDROMS were
distributed at ADASS, 100 being for the most recent version of Linux,
and the rest split equally between the previous edition of of Linux,
and Solaris. Our mailing list of names gathered at meetings prior to
ADASS has over 150 names, and we continue to receive new requests
about 1 per day. Interestingly, many requests are coming from
non-academic addresses such as aol.com, indicating that we may soon
have to consider the pros and cons of restricting distribution.
However, for the moment, the attention is welcome.

The release delivered the following capabilities:

\begin{itemize}
\item There is a guide to \htmladdnormallink{Getting Started in
AIPS++}{\gsURL}, and an \htmladdnormallink{AIPS++ cookbook}{\cookbookURL}.
\item All user capabilities of AIPS++ are documented via the
\htmladdnormallink{AIPS++ User Reference Manual}{\userrefmanualURL}.
\item The heart of AIPS++ is a powerful command line interpreter
called Glish (see Appendix A for copyright information). Glish has both a
\htmladdnormallink{tutorial}{\GlishtutorialURL} and a
\htmladdnormallink{reference manual}{\GlishmanualURL}.
\item Apart from the Glish command line interface, we also have
graphical user interfaces for most user-visible objects. These
represent a large investment of effort in direct response to
our earlier beta releases where we learnt that users needed simpler
interfaces.
\item In AIPS++, users do things using tools. The primary way to 
start, control and stop objects is via the Tool Manager ({\tt
toolmanager}), whereas files are handled by the File Catalog
({\tt catalog}).
\item In AIPS++, all data are stored in Tables, and are accessible
from inside Glish via the {\tt table} tool.
\item The synthesis imaging capabilities are found in the two tools
{\tt imager}, and {\tt calibrater}, the principal tools for imaging
and calibration. {\tt imager}, in particular, has many features not
found in other packages, such as multi-field processing in
user-selected coordinate systems. For VLA wide-field imaging, the
{\tt dragon} tool can be used.
\item To demonstrate the possibilities of stream-lined processing
in AIPS++, we have provided the {\tt simpleimage} tool. This
walks the user through straightforward imaging of synthesis
data, first making a tapered image at low resolution and then
imaging a user-specified region at full resolution. The user
has only to select the data to be imaged, and to delimit the
full-resolution region. This is an example of the type of
packaging that we plan to incorporate in the future.
\item The {\tt image} tool has many important capabilities
for the manipulation and display of images. One key is the use of the
Lattice Expression Language (developed by Ger van Diepen and Neil
Killeen) for arithmetic operations on images. A very powerful image
calculator based on the LEL is present in the release.
\item Display of images is possible using the {\tt viewer}
tool developed using the Display Library (David Barnes). The DL is a
C++ toolkit for the display of AIPS++ data types, such as Arrays,
Images, Tables, and MeasurementSets. The {\tt viewer} tool is the
Glish interface to the DL capabilities, and can itself be used
elsewhere as a component, such as in the {\tt imagefitter} tool.
\item The Measures system developed by Wim Brouw is to our
knowledge unparalleled in the types of operations it allows
on measured quantities with units and coordinate systems.
The functionality is available via both the Glish command line
(the {\tt measures} tool), and a specialized graphical user interface.
\item The single dish environment {\tt dish} is a carefully designed 
interactive tool for processing of single dish data, most particularly
spectra but other one-dimensional data formats are supported.
\item A very capable plotting tool, {\tt pgplotter}, based
on the PGPLOT subroutine library, is available and is used
by many other tools. For example, an interactive  visibility plotting
tool, {\tt visplot}, has been built on top of {\tt pgplotter}.
\item Numerous general purpose utilities for handling AIPS++ data
are present. For example, the tablebrowser (available as a function
{\tt browse} of the {\tt table} tool) can edit, plot, query and
select data, and can be configured in many ways to display data as the
user desires.
\item Introductory documentation has been written by early testers
of AIPS++, and terminology established for subsequent use throughout
the project. 
\item Extensive reference documentation is available for all
user-visible tools.
\item Programmability at the Glish level.
\end{itemize}

The following were missing:

\begin{itemize}
\item Polished interface for calibration of synthesis data
\item Filling of VLBA data apart from via UVFITS. Currently we expect 
you to apply AIPS++ to VLBA data that has been initially calibrated
elsewhere.
\item Efficient processing of many channel ({\em e.g.} more than 64)
spectral images. Current processing is optimized towards spectral-line
observations of moderate numbers of channels, and large continuum processing.
\item Efficient processing of large mosaic images. 
\item Interactive flagging of visibility data.
\item A robust way to archive the MeasurementSets to FITS. The best
is to use UVFITS for synthesis data ({\tt ms.tofits}) and 
{\tt ms2sdfits} for single dish data. These would not be loss-less
since these FITS formats do not map directly to the full complexity
of the MeasurementSet.
\item We do not expect people to program at the C++ level using the
release.
\end{itemize}


\section{Project overview}

After passing a very important milestone, the first release, it is a
good time to consider the overall health of the project, and to 
make any required changes:

\begin{description}
\item[The core package] We have delivered a working system that
contains much, but not all of what we hoped to. The basic tools and
command language (Glish) are acceptably robust and stable for the
first release, though improvements must still be made. The graphical
user interface is now useable, though improvements will be required as
we get user feedback. The Measures and Quanta system functionality is
superb. We have a first version of what will grow to be a very capable
display package. Image handling is good. Synthesis imaging is
excellent in many categories, with major improvements yet to be made
only in mosaicing and single dish imaging.  The agreed calibration
``thin-path'' is present. Overall, the core package is about a year
away from being complete enough that other competitive packages can be
decommissioned. {\bf Required action: The consortium should maintain a
focus on the core package for at least one more year.} We note that
this is as stated in the first quarterly report (1995 Q2) under the
current management:

\begin{quote}
{\em The long-term goal of the Project has been defined to be the
achievement of functional equivalence to AIPS by 2000. At that point,
AIPS will be a very small subset of AIPS++ and most applications areas
will look quite different from the corresponding areas in AIPS.}
\end{quote}

\item[External programmability] We have not been able to devote any
significant resources to a developer's release. The earliest that this
can occur based on current planning is early 2001. This relative
priority is in accord with various advice that we have received.
We can, of course, give a limited number of groups the required
access to develop in the system.

\item[The project staff] We have assembled a very high quality
team that works well together even spread across the globe. This
investment is substantial. While productivity in Glish is usually good
almost immediately, training a programmer to be productive in C++
typically takes 6-12 months. Every effort should be made to keep this
team together. One key deficiency is the lack of programmers to work
within the synthesis calibration package. To expand the synthesis
calibration package beyond the current thin path, we still need all
the coordinated effort that we can muster. Given this, it is worrying
to see that members of the consortium are planning to use the
synthesis calibration package in imaging pipelines. We believe that
while such use is close to being possible, first some key areas of the
calibration must be finished, and second, coordination is needed to
avoid unnecessary duplication of core capabilities. {\bf Required
action: Explicit coordination by the project center is needed. We
request that the various people dedicated to these imaging pipelines
first be utilized to finish the synthesis calibration package under
the direction of the project center.}

\item[Project process] For the most part, the Project has established
well-documented processes for many activities. For example,
substantial changes are proposed by a standardized {\em change
proposal} process. Defects are handled using a standard package and
protocols. The package health is monitored by documented tests run on
a regular basis. Coding and documentation standards are enforced on
review of the code. The latter slipped in priority during the
preparations for the first release, but is now getting more
attention. An informal comparison of our processes with those employed
by ESO show them to be comparable, with different areas of strength
and weakness but with overall parity.
\item[Internal communications] The project is effectively run by 
consensus amongst the various managers, with ties and day-to-day
decisions being made by the project management in Socorro.  The
unclear demarcation between these categories leads to in some cases
excessive communication over trivial matters, and in others, missed
consultation over areas that merit more discussion.  {\bf Required
action: To address these points, we plan to institute more open
notification and record keeping of decisions. In addition, a planned
get-together in April will help the entire group communicate.}
\item[Planning, Scheduling, {\em etc.}] The actual release date 
was about 6 months later than we anticipated a year ago: about a 50\%
underestimate. There are three main contributing factors of roughly
equal importance: 
\begin{itemize}
\item Unanticipated personnel changes contributed: the knock-on 
effects of Brian Glendenning leaving, the lack of a dedicated system
support person, the lack of a Chief Tester, and the
reduced role of Wim Brouw from the project (due to his temporary
secondment to the ATNF ATOMS project). All of these were
substantially damaging.  {\bf Required action: Find a system
person. Build in more realistic contingency for personnel factors}
\item The key contributing technical decision was the late integration 
of the Display Library into the package. We failed to anticipate the
difficulty of this step, which contributed about 2 months to the
overall delay of 6 months.
{\bf Required action: Avoid late integration in the future}
\item Stabilizing the core package required testing. In the end,
this testing was most effectively performed by AIPS++ staff,
thus cutting into their time for fixing the same defects.
{\bf Required action: Schedule testing by project staff. An existing user
base inside the consortium will also help.}
\end{itemize}
\item[Outreach] Now that we have a working package, outreach will
get easier and easier. For the moment, it is clear that we are
beginning to erase {\em some} negative perceptions of AIPS++.  The
case for AIPS++ has yet to be made convincingly in a number of
contexts, but we are confident that with the core package in hand,
such a case can be made. Our existing plan for outreach will be
followed.
\item[Advice] Due to the vagaries of the release process, the
STAG has not met since February 1998. To ensure that the next advisory
committee operates more from direct knowledge of the package, we plan
to reconstitute this group drawing from early adopters of the
package. We also plan to split off the technical advice into a
separate committee. We plan that both groups will meet around the time
of the next release.
\end{description}

\section{Developments planned for 1999 Q4}

The prime goals for the next quarter are to support the release, and
to work towards the new developments needed for the April 2000 release
(1.3). Details of the latter are covered in the the
\htmladdnormallink{development plan}{\devplanURL}. We anticipate
that the work allocated to support and to development will be roughly
equal.

\section{Appendix: ATNF contribution {\em  Neil Killeen}}

\subsection{General}

This quarter the ATNF had 3 people working in AIPS++.  These are Neil
Killeen (75\%, also local manager), Mark Wieringa (30\%) and David Barnes
(75\%).  Wim Brouw's time has ramped back up and he spent 60\% of his
time in aips++ this quarter.   He will spend about 20\% of his time
next quarter to conclude his work in ATOMS. In addition, Frederic Badia
at Narrabri started to use the system (see below) for some prototyping
of ATCA pipeline software.

This quarter was heavily focussed on testing, and preparing code and
documentation for the release in early October. 

\subsection{System}

This quarter we jettisoned the Solaris/Gnu system in accordance with
project policy and moved to purely egcs systems.  The Linux file locking
problems appear to have gone away after we finally got our Linux systems
upgraded to Redhat 6.0

At Narrabri, Digital Unix is the main platform.  Since there is no
binary release for that, the release will need to be built.  There may
also be a Solaris or Linux platform provided at Narrabri too.  Parkes
operates Solaris only.  

In addition, we hope that use of the egcs 1.0.3a system at Socorro (for
use with Purify, PureCoverage and Quantify) will be satifactory for our
needs so that we don't have to generate that extra system here (depends
upon link speed). 


\subsection{Talks}

Neil gave a 1-hour talk at Mt.  Stromlo on Glish.  He also gave two
short aips++ talks, one to general ATNF and CTIP staff, and another to
the visiting Gemini technical committee. 


\subsection{Visits}

David spent a few days at Dwingeloo discussing the Display Library
with Tom Oosterloo (one of its designers).  Neil spent a few days
in Dwingeloo talking mainly with Ger van Diepen and Jan Noordam.

\subsection{Individuals}

\subsubsection{David Barnes}

David's time (75\%) was spent on

\begin{itemize}
\item defect correction
\item testing viewer, image, simpleimage and calibrater
\item code tidying
\item documentation of viewer tools
\item development plan for next development cycle
\item Newsletter article
\end{itemize}

\subsubsection{Wim Brouw}

Wim's time (60\%) was spent on
\begin{itemize}
\item defect fixing and reporting
\item QAG related efforts, mostly related to the aips++ rules and the overall
  integrity of the system. Some of it resulting in extensive email
  discussions. 
\item small changes only in the Measures and Fitting classes, at request of users
\item discussion with Ger and Jan at NFRA
\item Development plan for next development cycle
\end{itemize}

\subsubsection{Neil Killeen}

Neil's time (75\%) went on:

\begin{itemize}
\item Testing of the image analysis components of the system
\item Testing of other components of the system
\item Many defect reports
\item Defect fixing
\item General preparation of image-related code and documentation for the release
\item Cookbook chapter on image analysis
\item Many emails
\item Newsletter article
\item Development plan for next development cycle
\item Talks to ATNF/CTIP staff, Mt Stromlo computer staff and 
  Gemini visiting technical committee
\item Started development again
\begin{itemize}
\item Classes to concatenate Lattices and Images
\item Separated ImageStatistics into Lattice and Image 
     dependent parts
\item reworked view function of Image tool
\item started ImageFFT class 
\end{itemize}
\end{itemize}

\subsubsection{Mark Wieringa}

Mark's time (40\%) went on
\begin{itemize}
\item completed work on MS V2 code, both aips and trial code dependent on
  these changes is now complete. Only the various fillers remain to be
  done by each consortium site.
\item Some work on rpfitsfiller (now taken over by Frederic Badia).
\item A lot of time was spent on defect fixes.
\item Some documentation work and some work on the installation of a release
  version.
\item Started development again
\item Started work on automated flagging towards the end of the quarter, got
     the existing code in MSFlagger working and implemented median window
     differencing scheme.
\end{itemize}

\subsubsection{Frederic Badia}

Frederic works in the Narrabri computing group.  

\begin{itemize}
\item He started work on prototyping an aips++ version of an imaging
  pipeline for the ATCA. After some time in Epping to get indoctrinated
  into the Display library and Glish, he has an initial version
  running. The development time was quite fast (apart from Digital Unix
  specific bugs which are frustrating but valuable, as they invariably
  show real coding errors).   
\item In parallel we have started a more formal process of drawing
  up a requirements document (workshop held in Sydney) and getting a
  proper design done in collaboration with Tom Oosterloo at NFRA.  There will 
  also be coordination   with overall aips++ project goals so that we
  maximize possible reuse of componentry.
\item Fred will also work on additions and MS V2 conversion of the ATCA filler until 
  the above process produces some guidelines for him.
\end{itemize}

\section{Appendix: BIMA/NCSA contribution {\em  Doug Roberts}}

Harold Ravlin worked to fix some bugs in the PostScript driver for the
Display Library.  Additionally, Ravlin has worked a bit on Aipsview;
this includes porting Aipsview to compile and work correctly under the
64-bit IRIX compiler.  Ravlin added True Color support to Aipsview,
which now displays on either pseudocolor or True Color visuals.  When
displaying in True Color manipulations of the color map are applied to
all windows after the mouse is released and provides good performance
on the target consortium machines.  Internally, Ravlin changes
Aipsview to use a modified version of the PostScript Driver from the
Display Library.  The integration of the PostScript Driver into
Aipsview resulted in identifying to bugs which were fixed in the DL.
Ravlin finished up some system administration duties including
recovering the NCSA IRIX server after a disk crash.  Doug Roberts was
involved in testing Aipsview, checking the affect of the Aipsview
changes.   Roberts also updated the Aipsview User's Guide to reflect
the changes to the program that we introduced this quarter.

Percent of time on AIPS++: Doug Roberts 70\%, Harold Ravlin 70\%

\section{Appendix: NFRA contribution {\em  Ger van Diepen}}


Local project members: Ger van Diepen (GVD), Jan Noordam (JEN, local
manager).

\subsection{General}

After a prolonged period of enforced introspection due to the
pressures of the WSRT upgrade, NFRA is now ramping up again its
contribution to the global AIPS++ package. Oleg Smirnov will arrive
any day now and will work in the area of uv-data calibration, with the
emphasis on the ionosphere. In addition, NFRA is taking part in a
collaboration with ATCA to create a modular 'calibration pipeline'.
Its modules should be directly usable in the global package.

In this quarter, the only substantial contribution to the global
package was made by Ger van Diepen in his role as Technical Leader,
and as the man responsible for a number of essential infrastructure
modules (tables, lattices, ms2uvfits, etc).  

In the meantime, the upgraded WSRT continues to rely heavily on
AIPS++. The online system (TMS) uses AIPS++ Tables, Measures and
Glish, and MS2uvfits for data export. The local WSRT 'mini-package',
written in Glish by Jan Noordam, is operational and widely used for
inspection of uv-data and determination of instrumental parameters
prior to observations. It is available in the global package too.

This quarter, the NFRA contribution in FTE's was as follows:

\begin{tabular}{|l|c|c|c|}
\hline
&                     Global AIPS++	& Local AIPS++	& Other \\
\hline
Ger van Diepen &	25\%		&     45\%	& 30\%	\\
Jan Noordam:   &        10\%		&     60\%	& 30\%  \\
\hline
\end{tabular}

(NB: The 'other' category includes vacations)

The program ms2scn has been upgraded to deal with the new values
of the WSRT in the ARRAY table.

The program j2convert has been extended considerably to make it
possible to fix errors in older WSRT measurement sets. Also the
way UVW and directions are calculated has been improved.

The program ms2uvfits has been improved a lot to be able to
export AIPS++ data to other packages. It took quite some effort
to get a correct file in the poorly documented UVFITS format.

A framework (MS2IWOS) has been developed to make AIPS++ measurement
sets accessible for the existing IWOS package in Westerbork.
Using this framework a program like MAKCAL could rather easily be
converted.

\section{Appendix: NRAO contribution {\em  Tim Cornwell}}

The core NRAO AIPS++ group is now Barnes (100\%), Braatz (100\%),
Cornwell (100\%), Garwood (100\%), Golap (100\%), Holdaway(100\%),
Kemball(100\%), Marson (100\%), McMullin (100\%), Schiebel (100\%),
Weatherall (50\%), and Young (100\%, Golap and Young are funded by
NCSA-NRAO collaboration). We have two open positions funded by
a three year NSF grant for visualization.

\subsection{Peter Barnes}

This quarter was marked by the conclusion of my trip to Sydney, extensive
testing and system work, and completion of targets (about 2 weeks spent
on science):

\begin{itemize}

\item Extensive amount of time spent on testing: MS DO, daily assays, pre-
release testing, release testing, etc etc etc.

\item Generated and verified (once they were resolved) several bug reports.

\item Very useful tutorials with David Barnes while visiting Sydney on funda-
mental parts of the Display Library: PixelCanvas, WorldCanvas, and
DisplayDatas.

\item Attended numerous meetings and discussions in the aftermath of the
Socorro synthesis meeting (which I missed due to my trip to Sydney).
Gained several useful insights about pixons and other image restoration
methods.

\item Monitoring of daily and weekly solaris builds, notifying various authors
of bugs as they occurred.  Various other system maintenance bookkeepings.

\item Completed top-to-bottom review of assay.g and dependent glish test
scripts.  This compilation was put into a text document listing all
methods of each major script that were tested, and also those not tested.
Altogether about 60,000 lines of glish code was looked over in a fair
amount of detail.

\item Some in-building (AOC) user support.

\item Attended first few "Getting Started" talks presented by Tim and Athol.

\item Mostly completed review of Lattice module, concentrating on documentation,
but also looking over code for usual compliance to guidelines.  This
consisted of generating extensive comments for 15 classes and the module
boilerplate.

\item Contributed to QAG part of development plan for next 6 mos.
\end{itemize}

\subsection{Jim Braatz}

Jim Braatz was hired on to the AIPS++ project in April 1999.  His primary
responsibilities are to manage the Green Bank installation of AIPS++, to
support GBT operations, and to work on single-dish applications.
Specifically, Jim worked on the following tasks during this quarter:

\begin{itemize}
\item Rewrote the standard plugin used with the pgplotter tool to accomodate
   requests from the GB engineering staff.  Specifically, the zoom
   feature needed to be redone to work properly when multiple Y-axes
   were included on a plot.

\item Worked with Ron Maddalena to prepare AIPS++ routines for the holography
   tests which took place in July.  Converted a number of routines
   used in the analysis of holography data from UNIPOPS to Glish and applied
   these routines to test data from the 140-ft telescope.

\item Worked to get several new users started with Glish and AIPS++ in 
   Green Bank.  Worked through some AIPS++ analysis of 140-ft HI data
   with Jay Lockman.

\item Ran tests and continued my own learning process of using AIPS++, with
   particular efforts to check out the pre-release of AIPS++ on the Linux
   operating system.

\item Updated some scripts in use at Green Bank which use features related
   to the old gplot1d routines.  These now use pgplotter routines.  These
   scripts were preventing us from removing the "old stable" installation
   in Green Bank.

\item Prepared a poster for the ADASS meeting - "Software for Testing and
   Commissioning the Green Bank Telescope"

\item Observatory-related duties this quarter included:
   Coordinator of the weekly Science Lunch Talk in Green Bank
   Mentor for the RARE CATS week-long teacher research program
   Lecture on "Masers and Cosmology" to the REU program
   Mentor to an NRAO REU summer student, who is working on VLA data
   
\item Science work included: Reduction and analysis of VLA data on 
several H2O Maser systems, with REU student Dave Sand.
\end{itemize}

\subsection{Tim Cornwell}

Tim worked on:

\begin{itemize}
\item Preparing the release.
\item Testing the release.
\item Gave tutorial lectures at AOC on AIPS++.
\item Visited Imke de Pater to introduce her and David Dunn to
AIPS++.
\item Prepared three posters and a demonstration for ADASS.
\item Implemented OTF imaging in {\tt imager}
\item Added basic editing to {\tt visplot}
\item Contributed three chapters to the cookbook
\item Arranged update of development plan
\item Hosted AIPS++/MMA Mosaicing meeting
\item Collaborated with Holdaway on development of Multi-scale clean
algorithm, and on mosaicing.
\item Collaborated with Golap on development of wide-field imaging.
\end{itemize}


\subsection{Bob Garwood}

Bob Garwood's primary responsibility is to oversee and contribute
towards the single dish work in AIPS++.  This work remains focussed on
the DISH environment and the support of the GBT.  His contribution in
support of the GBT is primarily through the GBT fillers which convert
the GBT FITS data files to an AIPS++ MeasurementSet.  In addition to
these duties, he is responsible for the maintenance and enhancement of
the FITS classes and, as author and maintainer of the GUI guidelines,
he spends some amount of time reminding other AIPS++ workers on the
importance of following those guidelines.

Over the past 3 months (July through September) he has done the
following:

\begin{itemize}
\item	Worked to resolve several defects and address enhancement
	requests in preparation for the first AIPS++ release.
	Most of this work was related to DISH and the FITS classes.

\item	Worked on converting 12-m On-the-fly data from UniPOPS to
	an AIPS++ MeasurementSet.  After a few iterations this was
	successfully done.  Tim Cornwell used this data while
	developing the initial single dish OTF gridding class.
	This work was part of an ADASS poster prepared this quarter
	and present in October (this poster also highlighted work done
	by Joe McMullin this quarter).

\item	Worked on two additional posters for ADASS.  One describing
	the state of DISH - highlighting changes over the past year
	and other describing the Single Dish FITS (SDFITS) binary
	table convention which is the primary archival format for
	single dish data in AIPS++.

\item	Fixed several bugs in the GBT filler which were noted as a
	result of the GBT Holography test last quarter in preparation
	for GBT Holography test this quarter.  The GBT filler was also
	turned into an AIPS++ distributed object in preparation for 
	its use as an on-line data filler.

\item	Participated in discussions with GBT staff that resulted in the
	final agreed format for the GBT spectrometer backend FITS files.

\item	Visited Green Bank several times.

\item	Updated the single dish development plan.
\end{itemize}


\subsection{Mark Holdaway}

Mark Holdaway organized the AIPS++ sponcored Mosaicing techniques
Meeting in Socorro, NM on June 14-17.  The meeting was attended by
about 35 people.  The presentations were informal and stimulating, and
the discussions, led by Tim Cornwell and Ron Ekers, were insightful.

Mark has revived Mark Wieringa's simulation software.  He ran several
suites of simulations comparing Multi-Scale Clean with MEM and the
Clark Clean.  These simulations document MEM's positivity bias due to
noisy data, and demonstrate that the Multi-Scale Clean suffers no
bias.  Yet, the Multi-Scale Clean is much better than MEM and Clark
Clean at accurately reproducing extended structure.  This work was
included in posters on Multi-Scale Clean at the URSI and ADASS
meetings.

Mark has also worked on getting MEM and Multi-Scale Clean into imager.
Many bugs were fixed.

\subsection{Athol Kemball}

I have continued to coordinate synthesis development during this
quarter, in addition to code contributions in the area of calibration,
and related defect identification and correction. This is described in
the synthesis report in more detail. Other direct involvement in
coding or infrastructure includes participation in design of uv-data
visualization code within the Display Library, the production of the
first edition of the cookbook, automated imaging design, wide-field
parallelization, intercomparisons of the AIPS and AIPS++ VLA data
fillers and preparations for MS v2.0. System involvement included
monitoring of overall memory use and development of a megaserver
solution to reduce this use, work in support of the adoption of egcs
1.1.2 as the project compiler, and evaluation of different memory
checking tools and related compilers.  I have continued also to
coordinate the parallelization and high-performance computing
initiative, which is a collaborative venture with NCSA. Work in this
area has included research in parallel I/O, contributions to build
stability on the SGI, wide-field parallelization as described above
and the current NT port and cluster research with P. Cortes. The
parallelization effort is described separately in this report. Release
preparations have required a considerable fraction of time this
quarter, both in testing and defect correction, and in verification of
installation procedures under various architectures. I have also
participated in the public outreach program at the AOC to provide a
tutorial for new scientific users of AIPS++, in addition to general
administration.

\subsection{Joe McMullin}

Joseph McMullin's main responsibilities are work on single dish 
applications within AIPS++; he was away on US Army Reserve training from
20 July - 30 Sept 1999.

\begin{itemize}
\item Added an imaging tool which allows lookmaps (maps of scan locations), 
spectral maps (plots spectra at each scan location) and simple contour
maps of selected velocity intervals.

\item Packaged the GBT commissioning tasks into a single tool accessible through
the AIPS++ toolmanager.

\item Worked on abstracts and posters for the ADASS meeting.
\end{itemize}

\subsection{Kate Weatherall}

\begin{itemize}
\item	Edited "Getting Started in AIPS++"
	    Manual, and replaced obsolete images.

\item  Edited and produced September Newsletter,
	    revised time mast.

\item  Started work with a demonstration copy of Framemaker + SGML

\item  Started a database of CD requests.  

\item  Continue to fix broken links reported by
	    htdig and other bug reports when I received
	    them.
\end{itemize}

\subsection{Wes Young}

Wes and Lisa had a baby.

\section{Appendix: Summary of AIPS++ Personnel Changes}

In this section, we give the names of people in the various AIPS++
groups and the nominal fraction of time allocated to AIPS++.

The ATNF group is: Neil Killeen (75\%, also local manager), Wim
Brouw (60\%), Mark Wieringa (30\%) and David Barnes (75\%).

The BIMA/NCSA group is: Harold Ravlin (70\%), Doug Roberts (70\%).

The NFRA group is: 

\begin{tabular}{|l|c|c|c|}
\hline
&                     Global AIPS++	& Local AIPS++	& Other \\
Ger van Diepen &	25\%		&     45\%	& 30\%	\\
Jan Noordam:   &        10\%		&     60\%	& 30\%  \\
\hline
\end{tabular}

The core NRAO AIPS++ group is now Barnes (100\%), Braatz (100\%),
Cornwell (100\%), Garwood (100\%), Golap (100\%), Holdaway(100\%),
Kemball(100\%), Marson (100\%), McMullin (100\%), Schiebel (100\%),
Weatherall (50\%), and Young (100\%), Golap and Young are funded by
NCSA-NRAO collaboration). We have two open positions funded by
a three year NSF grant for visualization.

Thus, in aggregate, we have 20 people contributing about 16.65 FTEs to
the AIPS++ Project. Of these, 12 are employed by NRAO (one funded by
NCSA), and contribute 11.5 FTEs. The numbers for the other partners
are: ATNF 4 and 2.4, BIMA/NCSA 2 and 1.4, NFRA 2 and 0.35. We have
two open positions in NRAO.


\end{document}

