\documentclass{article}
\usepackage{html,makeidx,epsf}
\pagestyle{headings}

\textheight  8.5truein
\textwidth   6truein
\topmargin   0.25truein
\oddsidemargin 0.25truein
\evensidemargin 0.25truein
%
% define hyperlink URLs:
%
\newcommand{\thisdocURL}
{http://aips2.nrao.edu/docs/project/decisionlog/decisionlog.html}
\newcommand{\searchURL}{../../html/aips2search.html}
\newcommand{\GlishmanualURL}{../../reference/Glish/Glish.html}
\newcommand{\GlishtutorialURL}{../../notes/195/195.html}
\newcommand{\gsURL}{../../user/gettingstarted/gettingstarted.html}
\newcommand{\grURL}{../../gettingresults/gettingresults/gettingresults.html}
\newcommand{\userrefmanualURL}{../../user/Refman/Refman.html}
\newcommand{\specsURL}{../../specs/specs.html}
\newcommand{\consortiumURL}{../consortium.html}
\newcommand{\defectlink}[2]{
\par
\htmladdnormallink{{#1: #2}}{http://aips2.nrao.edu/ddts/ddts_main?bug\_id=#1&REMOTE_USER=nobody&LastForm=DumpBug}
}
%
% URL to Help system
%
\externallabels{../../user/Refman}{../../user/Refman/labels.pl}
\externallabels{../../reference/Glish}{../../reference/Glish/labels.pl}
\externallabels{../../user/Utility}{../../user/Utility/labels.pl}
\externallabels{../../user/General}{../../user/General/labels.pl}
\externallabels{../../user/Display}{../../user/Display/labels.pl}
\externallabels{../../user/Dish}{../../user/Dish/labels.pl}
\externallabels{../../user/SynthesisRef}{../../user/SynthesisRef/labels.pl}
\externallabels{../../user/ATNF}{../../user/ATNF/labels.pl}
\externallabels{../../user/NFRA}{../../user/NFRA/labels.pl}
\externallabels{../../user/NRAO}{../../user/NRAO/labels.pl}


\begin{document}

\title{{\bf AIPS++ decision log}}

\author
{
AIPS++ Project Management\\
National Radio Astronomy Observatory\\
P.O. Box O, Socorro, NM, 87801, USA
}

\date{{Text last updated: 2001 July 19}\\
{HTML version available at:}\\
\texttt{{\thisdocURL}} 
}

\markright{AIPS++ decision log}
\maketitle

\tableofcontents

\section{Purpose}

This document records the date, background and results of major
decisions taken during project development. It is edited and
maintained by the project management.

%---------------------------------------------------------------------

\section{\underbar{Date: 2001-06-11: Open formal decision log}}

\subsection{Background}

While decisions have been recorded in the project documentation up to
now, including project status reports, weekly meeting minutes and
e-mail exploders, a central reference is important for easy access by
project members. This was agreed in the Software Engineering practices
described in Note 237.

\subsection{Result}

Created file doc/project/decsionlog.latex. Will make visible on the
AIPS++ web page and maintain as a central repository of major AIPS++
project decisions.

%---------------------------------------------------------------------

\section{\underbar{Date: 2001-06-15: Drop support for IRIX 32-bit}}

\subsection{Background}

The parallelization group have been using a 64-bit IRIX build with
the SGI native C++ compiler for all production runs and regular
devlopment. A 32-bit version was maintained however, as Rational did
not support 64-bit development in the IRIX version of Purify. The
latest version of Purify IRIX now provides this.

\subsection{Result}

Dropped support for IRIX 32-bit, and terminated all AOC builds of this
type.

%---------------------------------------------------------------------

\section{\underbar{Date: 2001-07-02: Check-in policies and build stability}}

\subsection{Background}

Build stability is vital in the current operational phase, both for
developers and scientific end-users. Strategies for achieving this
are defined in Note 237 (Software engineering practices in AIPS++),
and in prior e-mail discussions (TC aips2-workers \#2240).

\subsection{Result}

Refresh and clarify the current check-in policy (revised aips2-workers \#2240).

\begin{itemize}

\item{Maintaining the stability of daily is a high priority for
everyone in the project. Breaking daily affects the work of a large
number of people.}

\item{Maintaining the stability of weekly is essential. Breaking
weekly is disastrous for the work of a large number of people. Our
goal should be that every weekly can be marked as stable.}

\item{Checkins of changes to existing code should be made with
the above two goals in mind.}

\item{Changes to code that is used by others must be well-tested on
the project compiler before checkin, including running the associated
assay() tests. Checkin of purely developmental code is less crucial
but still good habits should be followed.}

\item{All checkins except simple ones, and bug fixes should be
done early in the week. Early means with two more of your inhales 
to get it right before the final weekly.}

\item{If you have a substantial change that will require more than a
week to test, this should be explictly scheduled in consultation with
the project management so that it can be coordinated across the
consortium.}

\item{All development sites should be running a daily using the
project compiler.}
 
\end{itemize}
%---------------------------------------------------------------------

\end{document}
