% AIPS++ Implementation Memo Series
%
% This file lists the available memos. It has some texinfo
% formatting embedded within it so it may be automatically
% included in texinfo documentation (in particular, the 
% Project Book).
\begin{enumerate}

% 112
\item[112] \htmladdnormallink{\textit{\raggedright AIPS++ Review: Report of the Review Panel}}{../112/112.html}
\linebreak  1994/12/15 Offen et al.
\linebreak  The findings and recommendations of the review panel.

% 111
\item[111] \htmladdnormallink{\textit{\raggedright The Design of AIPS++}}{../111/111.html}
\linebreak  1995/02/17 Glendenning.
\linebreak  Written (and only very slightly modified) for the AIPS++ review of
   December 1994.

% 110
\item[110] \htmladdnormallink{\textit{\raggedright AIPS++ Mission Statement}}{../110.text}
\linebreak  1993/09/03 Hunt (Ed)
\linebreak  AIPS++ Mission Statement.  Adopted by the AIPS++ steering committee
   16th October, 1993.

% 109
\item[109] \htmladdnormallink{\textit{\raggedright An Overview of AIPS++ Design}}{../109.ps}
\linebreak  1993/05/08 Hjellming, Glendenning.
\linebreak  Overview document on AIPS++ covering the status of class design as of
   April, 1993.

% 108
\item[108] \htmladdnormallink{\textit{\raggedright AIPS++ Software Design: Telescope Data Handling}}{../108.ps}
\linebreak  1993/04/26 Bottomly.
\linebreak  Software design document for handling instrument data input using
   Rumbaugh methodology.

% 107
\item[107] \htmladdnormallink{\textit{\raggedright Spherical Map Projections}}{../107/107.html}
\linebreak  1992/10/27 Calabretta.
\linebreak  Concise mathematical description of spherical map projections for use in
   astronomy.

% 106
\item[106] \htmladdnormallink{\textit{\raggedright AIPS++ Familiarization}}{../106.text}
\linebreak  1992/09/02, 1992/10/06 Hjellming, Glendenning.
\linebreak  Recommendations for learning about, and programming in, AIPS++.

% 105
\item[105] \htmladdnormallink{\textit{\raggedright Tentative Proposal for an \textsc{aips++} Project Book}}{../105/105.html}
\linebreak  1992/04/09 Hjellming, Glendenning.

% 104
\item[104] \htmladdnormallink{\textit{\raggedright \textsc{aips++} prototype}}{../104/104.html}
\linebreak  1992/03/20 Olnon (Ed).
\linebreak  Report on a small prototype the development group undertook.
   Two figures are only available in paper.

% 103
\item[103] \htmladdnormallink{\textit{\raggedright Calibration, Imaging, and Datasystems for \textsc{aips++} - Report of the
   meeting held at Green Bank, West Virginia, 3rd-14th February, 1992}}{../103/103.html}
\linebreak  1992/02/18 Shone, Cornwell (Eds).
\linebreak  Has several figures and many pages of text that are only available in
   paper, but the important expository text is available electronically.

% 102
\item[102] \htmladdnormallink{\textit{\raggedright An Initial Design for Major \textsc{aips++} Objects}}{../102.ps.gz}
\linebreak  1992/01/31 Hjellming, Glendenning. 
\linebreak  An incomplete note that set context for the Green Bank meeting. 
   Has two figures unavailable in the electronic version.

\end{enumerate}
