%                      AIPS++ User Specifications Memo 101
%
%                   The AIPS++ User Specification Memo Series
%
%                               R. M. Hjellming
%
%          National Radio Astronomy Observatory, Socorro, NM  87801-0387
%
%         LaTex Document file:   /zia/u/rhjellmi/memo101.tex
%
\documentstyle[11pt]{article}  % YOUR INPUT FILE MUST CONTAIN THESE TWO LINES 
\renewcommand{\baselinestretch}{1.5}
\textwidth 6.25in
\textheight 9.0in
\topmargin -1cm 
\oddsidemargin 0.15in
\evensidemargin 0.15in
\marginparwidth 0.5in
\begin{document}          % Command prefacing all document text
\hfill{August 28, 1991}
\centerline{ AIPS++ User Specifications Memo 101}
\vskip 0.3cm
\centerline{\bf The AIPS++ User Specifications Memo Series}
\vskip 0.3cm
\centerline{\it R. M. Hjellming}
\vskip 0.3cm
\centerline{National Radio Astronomy Observatory\footnote{ The
National Radio Astronomy Observatory is operated by Associated
Universities, Inc. under a cooperative agreement with the National
Science Foundation}, Socorro, NM  87801-0387}
\section{Purpose of this Memo Series}
    AIPS++ is a software system that will be developed and implemented
by a consortium consisting of the NRAO and a number of other
observatories and universities in the United States and around the
world.  It will be implemented in C++ using techniques of
object-oriented programming.  This memo series is intended to provide
a means for orderly, referenceable documentation of concepts about
what AIPS++ should be from the point of view of the user.

G. Croes has asked me to form a small group to organize the
writing of user
specifications for AIPS++.  Out of the E-mail discussions, and based
upon obvious ideas from AIPS and other systems, we will invite
contributions to this memo series.  However, anyone is invited to
contribute.   The elements of of this memo series will be mostly
individual contributions, but documents that
synthesize specifications, for AIPS++ in general or particular
instrument/applications, will also be included.
\section{Implementation of AIPS++}
   The initial synthesis of concepts defining AIPS++ will be carried
out under the direction of the AIPS++ project manager, G. Croes, by
a group of astronomers and programmers gathered at the NRAO in
Charlottesville, VA, USA, in the first half of 1992.  It is hoped that
this memo series will, before and during that period, grow into an
extensive set of useful astronomical, user, etc specifications. 
During the design
and subsequent implementation phase, programmers associated
with the AIPS++ development groups
will evolve selected specifications into 
consistent C++/OOP code for AIPS++.   This will begin during the
design phase, then continue at multiple sites around the
world doing code development 
coordinated, maintained, and re-distributed by the group
in Charlottesville.
\section{Other Documentation}
   There may be other memo series 
that will deal specifically with AIPS++ components as
they are implemented, and as they require documentation as completed
``tasks''.   Eventually programmer manuals, user guides, etc will
provide summary documentation, but memo series are an excellent way to
maintain, and make widely available, a written record of an evolving system.
\section {Submission and Distribution of This Memo Series}
   The AIPS++ User Specification Memo series will be maintained and distributed
with a serial
numbering system by Meri Stanley at the National Radio Astronomy
Observatory, P.O. Box O, Socorro, NM 87801-0379, USA.  Submissions to
the series should be sent to Robert M. Hjellming at the same address.
Because time is critical in late 1991 and early 1992, documents may be
submitted
in \TeX\ or \LaTeX\ form, in a single ASCII text file sent to
rhjellmi@nrao.edu via E-mail; please do not use any macros except
those defined in a single \TeX\ or \LaTeX\ input file.
Submissions with figures should be sent by mail.  Anyone may get
on the memo distribution list by: writing to the above; sending an E-mail
message to mstanley@nrao.edu or rhjellmi@nrao.edu; or calling 505-835-7310.

\end{document}            % Last command in LATEX document file









