\documentstyle[11pt,a4]{article}

\newcommand{\tbt}[4]{\left(\begin{array}{cc}#1 & #2\\ #3 & #4 \end{array}\right)}
\newcommand{\obt}[2]{\left(\begin{array}{cc}#1 \\ #2 \end{array}\right)}
\setlength{\parindent}{0mm}
\setlength{\parskip}{\medskipamount}


\begin{document}
\title {ATCA Calibration requirements for AIPS++ \\
        AIPS++ User Specification Memo 117}
\author{M.H.\ Wieringa}
\date{1995/07/18}
\maketitle
\section{Introduction}
Currently, most calibration needs for the Compact Array are covered by the
calibration routines in the Miriad package, written by Bob Sault.
In this document I will summarize the available algorithms and some
obvious extensions that AIPS++ would be expected to provide.
I will not venture into any `exotic' calibration schemes or image
plane corrections at this stage, but restrict this to what we would need
to allow users to switch from Miriad to AIPS++ for their ATCA processing.
This document uses the terminology of Jan Noordam's AIPS++ note 182.

\section{The Miriad model}
Bob Sault has given a one-page outline of Miriad visibility corrections
as part of the Dwingeloo UVCI effort, I summarize that here.

Miriad models the transformation of true sky visibility to observed 
visibility using the following, `antenna-based' gain function:
\begin{equation}
 g_p(t) g_p(\nu) \left({\nu\over\nu_0}\right)^{\alpha(t)}
\exp(2\pi i\tau(t)(\nu-\nu_0)).
\end{equation}
Here $g_p(t)$ is the frequency independent, time variable gain for 
IF-channel p; $g_p(\nu)$ is the time independent bandpass
function for IF-channel p; the $\alpha(t)$ term models time dependent
attenuation with a power law dependence on frequency; 
and $\tau(t)$ is a frequency and polarization
independent, time variable delay. The latter two terms are seldom used.
The terminology here is still a problem: antenna-based gains are really
the gains associated with a single IF-channel in a single antenna, possibly
but not necessarily associated with a single receptor (e.g., X-dipole).
The AIPS++ glossary refers to them as antenna/i.f.-gains.

In addition to these `single IF-channel' effects there are two complex
leakage terms, $D_p, D_q$, which describe the polarization mixing that occurs.
Given two ideal, orthogonal polarization signals, $E_p, E_q$, the measured
signals are given by:
\begin{equation}
\obt{E'_p}{E'_q}=\tbt{1}{D_p}{-D_q}{1}\obt{E_p}{E_q}.
\end{equation}
These leakage terms are assumed to be time independent and constant over the
bandpass. Each pair of IF-channels for a particular frequency will have its
own leakages however.

Two additional corrections Miriad can perform:
\begin{itemize}
\item for planet observations Miriad can remove variations in the planet flux and 
orientation using ephemeris values for apparent size and orientation,
\item for spectral line observations Miriad can perform Doppler tracking to 
 correct for observatory velocity.
\end{itemize}

\section{Calibration Schemes}

A `standard' ATCA observation consists of a short observation of a primary
flux calibrator and a sequence of secondary calibrator and program source
observations. In Miriad calibration proceeds as follows:
\begin{enumerate}
\item use primary calibrator to solve for bandpass and complex gains in
coupled fashion (assuming zero leakage and unpolarized calibrator),
\item use primary calibrator (with known polarization) to solve for 
 leakage and complex gains in coupled fashion (applying bandpass),
\item transfer bandpass and leakage corrections to secondary and use it
to derive complex gains (and the Stokes Q, U parameters of the secondary);
\item alternatively, for a strong secondary with good parallactic angle 
coverage, determine the bandpass from the secondary followed by
the leakage and complex gains (and Stokes Q, U parameters of the secondary),
\item rescale the gains of the secondary using the known flux of the 
primary,
\item transfer all corrections to the program source and image it,
\item optionally perform selfcal of the complex gains.
\end{enumerate}

The main difference with classic-AIPS calibration is the determination of
the Stokes parameters of the secondary in the solution for the complex gains.
This is necessary because generally XX and YY are not good estimates of
Stokes I (unlike RR, LL) due to linear polarization of the secondary.

The leakage solution cannot solve for all leakage terms using just
a primary with known and secondary with unknown polarization. 
The absolute leakage, or e.g., the X-leakage on the reference antenna,
has to be assumed to be zero. Using an observation of a strongly polarized
calibrator with known polarization and good parallactic angle coverage these
terms can be solved for. The Stokes V parameter also cannot be solved for.

Presently the Miriad bandpass and leakage solvers allow only 
a point source in the phase center as the model. In AIPS++ we should 
generalize this to arbitrary models. The bandpass solver allows a 
frequency dependent model for the source to be specified (to avoid 
contaminating the bandpass with the spectral index of the calibrator).
Similarly, the polarization leakage solver allows a full Stokes model to
be specified. Solving for and applying a time dependent bandpass is 
difficult in Miriad, in AIPS++ we should provide this capability.

There are two selfcal tasks in Miriad, one assumes there is only a single
time variable complex gain per antenna, the other allows two gains (X and Y) 
to be solved for, either independently or phase coupled.
There are provisions to specify selfcal models that vary with frequency, 
polarization or pointing (mosaicing). Not all these options can be combined 
at present. 

We need to be able to cope with several special cases: the data may consist
of only a single polarization (XX), two polarizations (XX,YY) or the full
set (XX,XY,YX,YY). Only in the last case can we do polarization leakage
correction. 

One area where Miriad has a rather simplistic approach compared to classic-AIPS
is the association of calibration with data. Only a single version of each
table (bandpass, leakage, gain) is supported. Recalibration overwrites the old
table. There are no provisions for applying the gains from calibrators
selectively to some program sources but not others. The solution for these 
problems in Miriad is to split the data into different sources or groups and
apply calibration to these separately. AIPS has a task, CLCAL, which allows 
one to specify in detail which calibrator to associate with which program
sources. Having used both methods, I tend to favor the Miriad approach as 
the conceptually easier one. In the case with multiple frequencies and
multiple 'calibration groups' selectively redoing the calibration can get 
rather difficult in AIPS. 
In AIPS++ we might avoid the copy of the actual data when 
splitting up the observation and provide 'referencing' datasets to which 
calibration can be applied. Comments on the merits of these schemes are
appreciated.

\section{Some effects we might want to consider}
\begin{itemize}
\item Faraday rotation - could be significant at 20cm, may be non-isoplanatic
\item Bandpass - proper treatment of S/N at band edges, 'effective frequency'
for continuum images
\item Tied array mode, Adding array mode (incoherent)
\item Interferometer based errors - correlator, time \& bandwidth smearing
\item Off-axis polarization, Asymmetry of primary beam
\item Redo/undo on-line Tsys calibration
\item Doppler tracking
\end{itemize}

\section{Summary of basic requirements}

\begin{itemize}
\item A Solver for bandpass and gain using the parallel hand correlations.
\item A Solver for polarization leakage, gain and optionally 
source polarization using all four correlations.
\item Selfcal Solvers using either single Stokes (I) correlation data or 
multiple correlations.
\item A Corrector for each of bandpass, leakage and antenna/i.f.-gain and
a versioning scheme for either the Correctors or their underlying tables.
\end{itemize}

\section{Acknowledgements}
Thanks go to Bob Sault and Wim Brouw for their suggestions.

\end{document}
