\section{The sdrecord}
\label{app:sdrecord}

The SDRecord is the fundamental data atom of dish.

The current sdrecord structure was arrived at early in the design
process.  The next version of the sdrecord will more closely resemble
individual rows of the \texttt{MeasurementSet}. 

Unless otherwise stated, the units of all values in an SDRecord are
SI units.  Angles are expressed in degrees.  This structure was
arrived at before Measures and coordinates were fully supported in 
\glish.  As a result, temporary fields in
an SDRecord are used to hold sufficient information to indicate
the coordinate system and reference frame where appropriate.
All of the fields described below are required unless stated otherwise.

The value of a field is undefined in the following circumstances:
\begin{itemize}
\item The \glish\ function \textbf{is\_nan(x)} returns \textbf{T} for a floating point
value.
\item The value is $< 0$ for an integer field.
\item The string is empty.
\end{itemize}
For boolean fields, all values are considered valid.  In addition,
all values in the data array should be considered valid.  The
data array flag field should be used to indicate invalid values
in the data array.

An  SDRecord is a \glish\ record having the following structure:
\begin{description}
%		SDRecord structure
\item[data] A \glish\ record which contains the data array and a
description of the data array (axis type, values, increments, and
units).
\begin{description}
%		data structure
\item[arr] The data array.  The dimensionality of the data array is
\texttt{(nchan, nstokes)} where \texttt{nchan} is the number of channels and 
\texttt{nstokes} is the number of different stokes types found in this data.
\texttt{nstokes} is either 1, 2 or 4.
\item[desc] A \glish\ record which describes the data array.
\begin{description}
%		data.desc structure
\item[ctype, crval, crpix, and cdelt] Follow standard FITS usage.
They describe the first axis of the data array.  An SDRecord can
therefore only contain data where the first axis is linear in some
coordinate system.  This limitation will be removed in the next version
of this structure.
\item[units] A string describing the units of the data array.
\item[stokes] A vector of strings describing the Stokes type of each
element of the second axis of the data array.  The possible Stokes
types are:
\begin{quote}
I, Q, U, V, RR, RL, LR, LL, XX, XY, YX, YY, RX, RY, LX, LY, XR, XL
YR, YL, PP, PQ, QP, QQ
\end{quote}
For a typical single dish telescope, raw data will have 
\texttt{stokes $=$ ("XX", "YY") or ("RR", "LL")}.
%		end data.desc structure
\end{description}
\item[flag] A boolean matrix with the shape as \textbf{arr}.  When true, the
corresponding value in \textbf{arr} has been flagged as bad and should not be
used.  When false, the corresponding value in \textbf{arr} is good.
\item[weight] A floating point matrix having the same shape as \textbf{arr}.
%		end data structure
\end{description}
\item[header] A \glish\ record with the following fixed structure.
The bulleted items are mearly separators to make this list easier to read,
they are not members of \textbf{header}.  
The type of each
field is indicated at the end of each description.  Vector fields
are indicated by giving the length of the vector in parenthesis
after the field type.
\begin{itemize}
%		header structure - grouped by type
\item Identity
\begin{description}
%		header structure - identity
\item[time] The mid-point of the observation (UT).  Double.
\item[scan\_number] The scan ID number.  Typically this is an identification
number given to a chunk of data when the data is taken.  Not all
telescopes provide a scan ID number.  Integer.
\item[object] The name of the thing being observed.  String.
\item[direction] Where the telescope was pointed.  Double(2).
\item[direction\_rate] Any slew rate during the observation.  Double(2).
%		end of header header structure - identity
\end{description}
\item Times
\begin{description}
%		header structure - time
\item[ut\_date] The UT date corresponding to time.  String.
\item[ut] The UT on ut\_date corresponding to time.  Double.
\item[lst] The LST corresponding to time.  Double.
\item[exposure] The total amount of time actually spent collecting photons
(e.g excludes blanking time, etc).
\item[duration] The elapsed (clock) time.  Float.
%		end of header structure - time
\end{description}
\item Observer
\begin{description}
%		header structure - observer
\item[observer] String
\item[project] String
%		end of header structure - observer
\end{description}
\item Backend
\begin{description}
%		header structure - backend
\item[rest\_frequency] Double.
\item[resolution] The frequency resolution.  This need not be the same
as the channel spacing.
\item[bandwidth] The total bandwidth. Double.
\item[obs\_frequency] The sky frequency at the reference channel.  Double.
\item[tcal] The cal temperature.  Float(nstokes).
\item[trx] The receiver temperature.  Float(nstokes).
\item[tsys] The system temperature.  Float(nstokes).
\item[sigma] The theoretical noise per channel.  Float(nstokes).
%		end of header structure - backend
\end{description}
\item References
\begin{description}
%		header structure - References
\item[reference\_position] A reference position on the sky for position
switched data.  Double(2).
\item[reference\_frequency] The frequency reference for frequency
switched data.  Double.
%		end of header structure - References
\end{description}
\item Telescope
\begin{description}
%		header structure - Telescope
\item[telescope] A string identifying the telescope.  String.
\item[telescope\_position] The (x, y, z) position of the telescope in
the ITRF (VLBI) coordinate system.  Double(3).
\item[telescope\_diameter] The diameter of the dish.  More detailed 
telescope models will no doubt be useful but the parameters to describe
these models are beyond the scope of the header record.  If
standard components could be identified, they would eventually be
placed in the header.  Until then, additional parameters to describe
the shape of the telescope surface should be placed in ns\_header. Float.
\item[telescope\_mount] The mount type.  Chosen from the following
values: ``alt-az'', ``equatorial'', ``X-Y'', ``orbiting'', ``bizarre''.
\item[azel] The azimuth and elevation at time.  Double(2).
\item[vcorr] The velocity of the telescope with respect to the
velocity reference frame at time using the given velocity definition.
%		end of header structure - Telescope
\end{description}
\item Weather
\begin{description}
%		header structure - Weather
\item[pressure] The atmopheric pressure.  Float.
\item[dewpoint] The dew point.  Float.
\item[tambient] The ambient temperature.  Float.
\item[wind\_dir] The wind direction.  Float.
\item[wind\_speed] The wind speed.  Float.
%		end of header structure - Weather
\end{description}
\item Coordinate system kludges
\begin{description}
%		header structure - Kludges
\item[direction\_coord] A \glish\ Record describing the coordinate system
for the \textbf{direction} field.  It has the following structure.
\begin{description}
\item[name] The coordinate name.  String.
\item[equinox] The equinox for the coordinate.  Double.
\end{description}
\item[reference\_coord] A \glish\ Record describing the coordinate system
for the \textbf{reference\_position} field.
It has the following structure.
\begin{description}
\item[name] The coordinate name.  String.
\item[equinox] The equinox for the coordinate.  Double.
\item[is\_offset] True if this is an offset position, else this is an
absolute position.  If this is an offset position then it is
offset from the \textbf{reference\_position}.  Bool.
\end{description}
\item[veldef] The velocity definition (e.g. ``radio'', ``optical'').  String.
\item[velref] The velocity reference frame (e.g. ``LSR'', ``HEL'', ,etc.).
String.
%		end of header structure - Kludges
\end{description}
%		end of header structure - grouped by type
\end{itemize}
\item[hist] This vector of strings contains a history of the operations
which resulted in this SDRecord.  Each element is a valid \glish\
command.  Taken as a whole, the hist field should be sufficient to
regenerate the SDRecord from the original data source.
%		end of SDRecord structure
\end{description}


In addition to the fixed structure described above.  An SDRecord
may contain an optional {\tt ns\_header} field.  This is a \glish\
record which can contain any additional information not found
in the fixed portion of an SDRecord.  This field will typically be
used for telescope-dependent information.

The x-axis can be easily constructed from the {\tt data} field
by the following simple \glish\ function:
\begin{tt}
\begin{verbatim}
	make_xaxis := function(ref data)
	{
	   x := data.crval;
	   x +:= ([1:data::Shape[1]] - data.desc.crpix)*data.desc.cdelt;
	   return x;
	}
\end{verbatim}
\end{tt}

To summarize, in pseudo-\glish\ an SDRecord looks like this:
\begin{verbatim}
	sdrecord := [data=[=], header=[=], hist=[=], ns_header=[=]];
	sdrecord.data := [arr=[...], flag=[...], desc=[=], weight=[...]];
	sdrecord.data.desc := [units=,stokes=,ctype=,crpix=,crval=,cdelt=];
        sdrecord.header := [time=,scan_number=,object=,etc.];
	sdrecord.hist := ["# useful history","# as a vector","# etc. etc."];
        sdrecord.ns_header := [non-standard fields]

ns_header is optional
\end{verbatim}

Additional information can be found in the 
Reference Manual section on
the \htmlref{\textbf{get}}{dish:sditerator.get.function} function of the 
\htmlref{Sditerator \texttt{tool}}{dish:sditerator}.  See also
the \htmlref{\textbf{is\_sdrecord}}{dish:is_sdrecord} function.

