\documentclass[11pt]{article}
\usepackage{epsfig}
\include{ahrecipe}
\begin{document}
\begin{ahrecipe}

\argoals{To reduce an on/off total power scan pair by extracting the ``on'' and the ``off''
source scans from an opened data set, constructing a difference scan, and inserting the
result into the DISH Results Manager.}

\arusing{dish}

\arassume{You have a data set named {\tt rawdata} opened and available in the DISH Results 
Manager.  An ``on'' scan is located at the first record and an ``off'' scan is located
at the third record.}
\arscript{}
\begin{arexample}
\arline{rawdata.setlocation(1)}{Move the {\tt rawdata} pointer so that it points at the first
record, where the ``on'' scan is located.}
\arline{on := rawdata.get()}{Get that scan and assign it to a variable named {\tt on}.}
\arline{rawdata.setlocation(3)}{Move the pointer to point at the ``off'' scan location.}
\arline{off := rawdata.get()}{Get it and assign it to {\tt off}.}
\arline{result := on}{Initialize the result by making it a copy of {\tt on}.  This sets
the structure and header values appropraite for this type of data.}
\arline{result.data.arr := (on.data.arr - off.data.arr) / off.data.arr}{Subtract the ``on''
data array from the ``off'' data array and divide the result by the ``off'' data array.}
\arline{dish.rm().add('result', 'Diff. of rows 1 and 3', result, 'SDRECORD')}{Add this
result to the DISH Results Manager.  The final argument tells the Results Manager that this
is an SDRECORD, a type of result which the Results Manager can display and interact
with.}

\arconclusion{This was a simple example.  Addition operations could be done
to check that this operation was appropriate for this data, to scale the
data, etc.  The whole range of AIPS++ utilities are available at the glish command
line for use on results known to DISH.  Any glish value can be stored in the Results
Manager.  This will preserve it, along with the results of the contents of the
Results Manager, between invocations of DISH.}
\end{arexample}
\end{ahrecipe}
\end{document}

