\documentclass{book}
\usepackage{html}

\externallabels{../../reference/Glish}{../../reference/Glish/labels.pl}
\externallabels{../../user/Utility}{../../user/Utility/labels.pl}
\externallabels{../../user/General}{../../user/General/labels.pl}
\externallabels{../../user/Display}{../../user/Display/labels.pl}
\externallabels{../../user/Dish}{../../user/Dish/labels.pl}
\externallabels{../../user/SynthesisRef}{../../user/SynthesisRef/labels.pl}
\externallabels{../../user/ATNF}{../../user/ATNF/labels.pl}
\externallabels{../../user/BIMA}{../../user/BIMA/labels.pl}
\externallabels{../../user/NFRA}{../../user/NFRA/labels.pl}
\externallabels{../../user/NRAO}{../../user/NRAO/labels.pl}

\begin{document}

\input{aips2defs}

\title{AIPS++ Recipes Repository}
\author{George Moellenbrock, editor}
\maketitle

This page is the entry point into the \aipspp\ Recipes Repository.
Recipes are scripts of limited scope which perform or demonstrate a
useful operation. There are currently three categories of recipes.
Recipes in the General category are scripts which use tools and
functions provided in the \htmlref{General}{general.package} and
\htmlref{Utility}{utility.package} packages in aips++, including
fitting, plotting, FFTs, etc.  Recipes in the Synthesis category
illustrate use of tools and functions in the
\htmlref{Synthesis}{synthesis.package} package, including calibration
and imaging of data taken with consortium synthesis instruments.
Recipes in the Single Dish category use tool and functions found in
the \htmlref{Dish}{dish.package} for single dish reductions.  Each
recipe page contains information on how to use the recipe, as well as
links to relevant documentation and the script itself.

A large and growing number of typical data reduction sequences exist
as user scripts, and the \aipspp\ project welcomes submissions.  A
template script file is available here:
\htmladdnormallink{recipe.g}{../../../code/doc/notes/249.dir/recipe.g}
Submissions should be emailed to aips2-recipes@nrao.edu.  More
information about recipe submissions can be found in \aipspp\ 
\htmladdnormallink{Note 249: The AIPS++ Recipes Repository}{../../notes/249/249.html}.

\begin{tabular}{ll} 

\htmladdnormallink{General}{../General/index.html} & generic data reduction, including fitting, plotting, FFTs, etc. \\
\htmladdnormallink{Synthesis}{../Synthesis/index.html} & synthesis data reduction and analysis \\
\htmladdnormallink{Single Dish}{../SingleDish/index.html} & single dish data reduction and analysis \\

\end{tabular}

\end{document}

