\section{The sdrecord}
\label{app:sdrecord}

The SDRecord is the fundamental data atom of dish. It has recently
been updated (V1.4) to more closely resemble individual rows of the
\texttt{MeasurementSet}.

Unless otherwise stated, the units of all values in an SDRecord are
SI units.  Angles are expressed in degrees.  This structure was
arrived at before Measures and coordinates were fully supported in 
\glish.  As a result, temporary fields in
an SDRecord are used to hold sufficient information to indicate
the coordinate system and reference frame where appropriate.
All of the fields described below are required unless stated otherwise.

The value of a field is undefined in the following circumstances:
\begin{itemize}
\item The \glish\ function \textbf{is\_nan(x)} returns \textbf{T} for a floating point
value.
\item The value is $< 0$ for an integer field.
\item The string is empty.
\end{itemize}
For boolean fields, all values are considered valid.  In addition,
all values in the data array should be considered valid.  The
data array flag field should be used to indicate invalid values
in the data array.

An  SDRecord is a \glish\ record having the following structure:
\begin{description}
%		SDRecord structure
\item[data] A \glish\ record which contains the data array and a
description of the data array (axis type, values, increments, and
units).
\begin{description}
%		data structure
\item[arr - float | complex] The data array.  The dimensionality of the data 
array is
\texttt{(nstokes,nchan)} where \texttt{nchan} is the number of channels and 
\texttt{nstokes} is the number of different stokes types found in this data.
\texttt{nstokes} is either 1, 2 or 4.
\item[desc] A \glish\ record which describes the data array.
\begin{description}
%		data.desc structure
\item[chan\_freq - Quantity[nchan]] 'unit' defines the units of the x-axis and
'value' defines the values of the x-axis.
\item[refframe - string] Defines the reference frame.
\item[reffrequency - double[Hz]] The frequency of the center pixel.
\item[chan\_width - double] The width of the channels in units of x.
\item[restfrequency - double[Hz]] The rest frequency of the observed line.
\item[corr\_type - string[nstokes]] A vector of strings describing the Stokes 
type of each element of the second axis of the data array.  The possible Stokes
types are:
\begin{quote}
I, Q, U, V, RR, RL, LR, LL, XX, XY, YX, YY, RX, RY, LX, LY, XR, XL
YR, YL, PP, PQ, QP, QQ
\end{quote}
For a typical single dish telescope, raw data will have 
\texttt{stokes $=$ ("XX", "YY") or ("RR", "LL")}.
%		end data.desc structure
\item[unit - string] A string describing the units of the data array.
\end{description}
\item[flag -bool[nchan,nstokes]] A boolean matrix with the shape as 
\textbf{arr}.  When true, the corresponding value in \textbf{arr} has been 
flagged as bad and should not be used.  When false, the corresponding value 
in \textbf{arr} is good.
\item[weight - float[ntokes,nchan]] A floating point matrix having the same shape as \textbf{arr}.
\item[sigma - float[nstokes,nchan]] Theoretical rms.
%		end data structure
\end{description}
\item[header] A \glish\ record with the following fixed structure.
The bulleted items are merely separators to make this list easier to read,
they are not members of \textbf{header}.  
The type of each
field is indicated at the end of each description.  Vector fields
are indicated by giving the length of the vector in parenthesis
after the field type.
\begin{itemize}
%		header structure - grouped by type
\item Identity
\begin{description}
%		header structure - identity
\item[time - MEpoch] The mid-point of the observation (UT). A measure. 
\item[scan\_number - int] The scan ID number.  Typically this is an 
identification number given to a chunk of data when the data is taken.  Not all
telescopes provide a scan ID number. 
\item[source\_name - string] The name of the thing being observed.  
\item[direction - MDirection] Where the telescope was pointed. A measure
with the reference frame. 
%		end of header header structure - identity
\end{description}
\item Times
\begin{description}
%		header structure - time
\item[exposure - double[s]] The total amount of time actually spent 
collecting photons (e.g excludes blanking time, etc).
\item[duration - double[s]] The elapsed (clock) time.  
%		end of header structure - time
\end{description}
\item Observer
\begin{description}
%		header structure - observer
\item[observer - string]
\item[project - string] 
%		end of header structure - observer
\end{description}
\item Backend
\begin{description}
%		header structure - backend
\item[resolution - double] The resolution.  The units are whatever chan\_freq 
has.
\item[bandwidth - double] The total bandwidth.
\item[tcal - float(nstokes)[K]] The cal temperature.  
\item[trx] - float(nstokes)[K]] The receiver temperature. 
\item[tsys - float(nstokes)[K]] The system temperature. 
%		end of header structure - backend
\end{description}
\item References
\begin{description}
%		header structure - References
\item[reference\_direction - MDirection] 
%		end of header structure - References
\end{description}
\item Telescope
\begin{description}
%		header structure - Telescope
\item[telescope - string] A string identifying the telescope.
\item[telescope\_position - MPosition] The (x, y, z) position of the telescope 
in the ITRF (VLBI) coordinate system. 
\item[azel - MDirection] The azimuth and elevation at time.  
%		end of header structure - Telescope
\end{description}
\item Weather
\begin{description}
%		header structure - Weather
\item[pressure - double[hPa]] The atmopheric pressure. 
\item[dewpoint - double[K]] The dew point.  
\item[tambient - float] The ambient temperature.
\item[wind\_dir - float] The wind direction. 
\item[wind\_speed - float] The wind speed.  
%		end of header structure - Weather
\end{description}
\item Coordinate system kludges
\begin{description}
%		header structure - Kludges
\item[veldef - string] The velocity definition (e.g. ``radio'', ``optical''). 
\item[molecule - string]
\item[transition - string]
%		end of header structure - Kludges
\end{description}
%		end of header structure - grouped by type
\end{itemize}
\item[hist - string] This vector of strings contains a history of the operations
which resulted in this SDRecord.  Each element is a valid \glish\
command.  Taken as a whole, the hist field should be sufficient to
regenerate the SDRecord from the original data source.
%		end of SDRecord structure
\end{description}


In addition to the fixed structure described above.  An SDRecord
may contain an optional {\tt ns\_header} field.  This is a \glish\
record which can contain any additional information not found
in the fixed portion of an SDRecord.  This field will typically be
used for telescope-dependent information.

To summarize, in pseudo-\glish\ an SDRecord looks like this:
\begin{verbatim}
	sdrecord := [data=[=], header=[=], hist=[=], ns_header=[=]];
	sdrecord.data := [arr=[...], flag=[...], desc=[=], weight=[...]];
	sdrecord.data.desc := [units=,stokes=,ctype=,crpix=,crval=,cdelt=];
        sdrecord.header := [time=,scan_number=,object=,etc.];
	sdrecord.hist := ["# useful history","# as a vector","# etc. etc."];
        sdrecord.ns_header := [non-standard fields]

ns_header is optional
\end{verbatim}

Additional information can be found in the 
Reference Manual section on
the \htmlref{\textbf{get}}{dish:sditerator.get.function} function of the 
\htmlref{Sditerator \texttt{tool}}{dish:sditerator}.  See also
the \htmlref{\textbf{is\_sdrecord}}{dish:is_sdrecord} function.

