\section{Data editing}
\label{GRediting}

% TBD: Something should be added to describe Oleg's autoflag.

Command-based, interactive and automated editing capabilities are
supported in the \flagger\ tool. 

To start the \flagger\ tool, type the following:

\small
\begin{verbatim}  

fg:= flagger(msfile='mydata.ms')        # start the flagger tool

\end{verbatim}
\normalsize

\subsubsection{Direct (command-based) editing}

The \flagger\ tool offers several directed editing (flagging)
operations.  In general, this involves selecting the data to be
flagged (based in indices, timeranges, etc.), then running a flagging
operation.  The flagging commands all have the option to run them on a
trial basis.  To flag data from antenna=5, spw=1, field=2, within a
timerange:

\small
\begin{verbatim}
fg.setantennas(ants=5)                       # Select data including antenna 5
fg.setids(spectralwindowid=1,fieldid=2)      # Select spectral id and field id
fg.settimerange(starttime="24-FEB-2000/10:00:05",  # Select timerange
                  endtime="24-FEB-2000/10:09:37");   
fg.state();                                  # Review data selection   
fg.flag(trial=T);                            # Trial run to see how many rows
                                             #   will be flagged
fg.flag();                                   # Flag it
\end{verbatim}
\normalsize

\noindent
To reset the data selection, type:

\small
\begin{verbatim}
fg.reset();
\end{verbatim}
\normalsize

\noindent
To flag data from baseline 3-4, spw=2, field=1, within $\pm$10s of a
timestamp:

\small
\begin{verbatim}
fg.setbaselines(ants=[3,4])                  # Select data for baseline 3-4
fg.setids(spectralwindowid=2,fieldid=1)      # Select spectral id and field id
fg.settime(centertime="24-FEB-2000/10:00:05",  # Select time and range
           delta="10s");
fg.state();                                  # Review data selection   
fg.flag();                                   # Flag it
\end{verbatim}
\normalsize

The logic in the data selection options is as follows.  The two
time-related data selection options (set by \flagsettimerange\ and
\flagsettime) are logically OR'd together, and the two
antenna-related data selection options (set by \flagsetantennas\ and
\flagsetbaselines\ are also logically OR'd.  The net time selections,
net antenna selections, and all other selections are then AND'd
together.  Note that the data selection persists after the flagging
operation \flagflag\ is executed.  To clear the selection, execute
the \flagreset\ function, or execute individual selection commands
without arguments to clear only that particular selection.  For
example, to flag all data to antenna 11 for the same spectral window,
field, and timestamp as for baseline 3-4 above:

\small
\begin{verbatim}
fg.setbaselines();                # Clears baseline selection
fg.setantennas(ants=11);          # Select antenna 11
fg.state();                       # Review data selection
fg.flag();                        # Flag it
\end{verbatim}
\normalsize

To unflag data based on a selection, use the \flagunflag\ function.
For example, to restore data for baseline 11-15 (that was flagged
above):

\small
\begin{verbatim}
fg.setantennas();                 # Clears antenna selection
fg.setbaselines(ants=[11,15]);    # Select baseline 11-15
fg.state();                       # Review data selection (note 
                                  #  other selections still in
                                  #  effect)
fg.unflag();                      # undo previous flags
\end{verbatim}
\normalsize

Note that use of \flagunflag\ may restore data that had been flagged
by other means (e.g., by the on-line system, or by other criteria in
previous use of \flagger\ or \msplot).  In the near future, a
flexible means of manipulating flagging versions will be made
available in \aipspp.

Consult the \flagger\ tool documentation for more information on the
details of data selection.

\subsubsection{Heuristic editing}

Several simple heuristic editing methods are available in the \flagger\
tool.  These methods act on the data selected as described above, but
add an additional heuristic constraint.

For example, it is usually desirable to flag the auto-correlation data
from the VLA (otherwise, the \imager\ tool may attempt to use it
in image formation as if it were single dish data).  To do this:

\small
\begin{verbatim}
fg.reset();                   # Reset the data selection
fg.flagac();                  # Flag the auto-correlations
\end{verbatim}
\normalsize

It is often the case that the VLA data near the beginning of scans
(e.g.  after field or frequency band changes is bad.  To flag such
data, use the \flagquack\ function.  For example, to remove 20
seconds of data at the beginning of scans (defined by gaps larger than
30 seconds) on field 3:

\small
\begin{verbatim}
fg.reset();                   # Reset the data selection
fg.setids(fieldid=3);         # Limit quack to fieldid=3 only

fg.quack(scaninterval='30s',  # Trial run of quack
         delta='20s',
         trial=T);

fg.quack(scaninterval='30s',  # Quack it
         delta='20s');
\end{verbatim}
\normalsize


Automated editing methods are available in the \flagger\ and \autoflag\
tools.  These include clipping, median flagging and related methods.

For example, to clip data outside of a specified amplitude range with
the \flagger\ tool:

\small
\begin{verbatim}
fg.setids(fieldid=2, spwid=[1,2]);    # Select field, spw
fg.setpol(pol=[1,4]);                 # Select polarizations 1,4
                                      #  (e.g., RR and LL)
fg.filter(column='DATA',              # Flag data if, DATA column
          operation='range',          #   Amplitudes are outside
          comparison='Amplitude',     #   1.12-1.34
          range=['1.12Jy','1.34Jy'],
          fullpol=T,                  # Extend flags to pols and
          fullchan=T);                #  and chans that weren't 
                                      #  tested
\end{verbatim}
\normalsize

The \autoflag\ tool has more sophisticated algorithms. Examples of
its use will be included here in the near future.

\subsubsection{Finishing flagger}

To finish the \flagger\ tool:

\small
\begin{verbatim}
fg.done();     # Done with flagging
\end{verbatim}
\normalsize



\subsection{Interactive editing}

Interactive editing is possible on almost all graphical views of the
data which are possible in \msplot.  To enable the editing option, it
is necessary to set parameter {\tt edit=T} when the \msplot\ tool is
created.  The current set of flags will be saved (for possible
restoration later).  Select and plot data as in non-editing mode, then
draw boxes with the left mouse button to select regions to flag and
press the {\tt Flag} button to flag it.  As many flag boxes as desired
may be drawn before executing the flags.  Use the {\tt Locate} button
to obtain a listing (in the {\tt logger}) of the data points within
the box(es).  Note that the listing may be very long if the region
selected covers a large number of visibility points.

At this writing the channel and polarization flagging scope selections
in {\tt msplot} will only work if you have elected to plot all channels
and/or polarization.  This constraint will be relaxed in the near future.

Raster displays of visibility data are possible also, and maybe useful
for spectral line data editing.

