\label{GRmsinspection}

\section{Data examination and inspection}

Once the data have been filled to a MeasurementSet, they can be examined or
displayed using various tools. The
\htmlref{Catalog}{catalog:catalog} tool provides an overview of
\aipspp\ data files at the highest level, as individual \aipspp\
tables of different types (e.g. images, MS, etc) in a given
directory. The \htmlref{Catalog}{catalog:catalog} tool can be started
from Glish as:

\small
\begin{verbatim}
include 'catalog.g'
dc.gui()
\end{verbatim}
\normalsize

Here, {\tt dc} is the default \catalog\ tool. It can also be accessed
via the Toolmanager interface from the {\tt Tools in Use} button. A
mouse click on an individual MS listed in the catalog GUI followed by
a click on the {\gif Summary} button, will write a summary of the data
in the MS to the logger (including frequency and polarization settings
and fieldnames).  This information is useful during calibration.

When you click on the {\gif View} button (to view the presently selected
catalog item) you will launch the \tablebrowser\ which provides facilities to
browse and examine all data fields in the MS main and sub-tables. 

Note that indices in the MS, such as FIELD\_ID, are stored as integers
beginning with zero when viewed at the low level provided by the
\tablebrowser.  All \aipspp\ synthesis tools correct for this artifact
of c$++$ and present selection of indices beginning at one.

The primary graphical visibility data visualization tool in \aipspp\
at present is \msplot.  \msplot\ allows line and raster plots of the
observed, corrected, and model data against a range of MS indices and
parameters, as well as associated plots such as uv-coverage.
Facilities are provided to support versatile data selection, range
setting and plot iteration. Interactive editing is also supported, as
described in the next section.  It can be started from \glish\ via:

\small
\begin{verbatim}
include 'msplot.g'
p := msplot(msfile='data.ms',
            edit=T);                # Allow editing as well as inspection
\end{verbatim}
\normalsize

This will launch a gui for plot type and data selection.  It can also
be launched from the \toolmanagerref\ via the {\mf Synthesis} package, and
\ms\ module.

Consult the \msplot\ documentation for more information on using this
tool.

