\chapter{Arecibo Observatory Spectral Line Reduction with DISH}
\label{aohi}

{\em K. O'Neil and J. McMullin}

\section{AIPS++ Versions Available at Arecibo}

The are a number of different version of AIPS++ available at Arecibo: 

\begin{itemize}

\item{}
    {\bf Stable:} This is the version which should be the most reliable aips++ 
	version available at Arecibo. On the other hand, this is often the least
	recent version available, as it is only updated every month. 

\item{}
    {\bf Weekly:} This version exists for those people who wish to run an 
up-to-date version of aips++, and who are willing to accept a few bugs/problems
with the version. As its name indicates, this version is updated once a week 
and is used mainly be folks working on developing aips++ software. 

\item{}
    {\bf Daily:} This version should only be used by aips++ software 
	developers. Because it is updated daily, it can be extremely unreliable. 
\end{itemize}

\section{Setting Your Environment}\label{sect:2}

\subsection{Starting AIPS++}

To start AIPS++, issue the following commands (from the c shell):

\begin{verbatim}
> source /pkg/aips++/stable/aipsinit.csh
> aips++ -l naic.g
\end{verbatim}

\subsection{Making AIPS++ run faster}

Here are several hints for speeding up your performance,
if the operation/startup of AIPS++ seems sluggish.

\begin{itemize}
\item{} Store cache files on the computer you run from by adding the
        following line to your .aipsrc file:
        \begin{verbatim}
        user.aipsdir: /tmp/username_aips++

        where username is your username, e.g.,
        user.aipsdir: /tmp/rphillips_aips++
        \end{verbatim}


\item{} If you do not want to have the toolmanager GUI available on starting
        AIPS++, you can suppress it and speed up the startup time by including
        the following lines in your .aipsrc file (located in your home directory):
        \begin{verbatim}
        toolmanager.fullgui: F
        toolmanager.gui.auto: F
        \end{verbatim}

        The toolmanager isn't directly needed for any DISH operations. Should
        you want it, it can be used at a later time by typing:
        \begin{verbatim}
        tm.gui();
        \end{verbatim}

\item{} Run AIPS++ from the computer which hosts the disks with your data.
        Typically the data I/O over NFS is slow and the above bypasses this
        as much as possible

\end{itemize}

AIPS++ builds its cache file structure the first time it is started with a
new user.aipsdir, and each time you start a new version. Judge its startup
speed on the {\it second} time you start it up, not the first.

\subsection{Paging}

If you find that error messages, code, and data listings scroll off of your
terminal window, you might need to assign the {\bf PAGER} environment
variable. You can remedy this by setting {\bf PAGER} to 'less' in your
UNIX startup scripts.

\begin{verbatim}
export PAGER='less'
\end{verbatim}

\section{Converting AO Data into SDFITS}

Arecibo data must first be converted from its native format
into single-dish FITS format, through use of the "sdtest" program
(/usr/S2local/bin/sdtest). Single-dish fits data are stored as FITS binary 
tables with a single row for each spectrum: A single dump usually generates 
4 or 8 spectra (four subcorrelators with either one or two polarizations per 
board). If data are taken with different numbers of channels per spectrum, 
the file size will be as if they all had the largest number of channels. 
If the number of channels increases during the conversion, then sdtest will 
resize the entire file, which may take a while. 

Before running the sdtest program you need to know the following:

\begin{itemize}
\item{} Where is your data located (i.e. at /share/olcor/... or /proj/a9999/...) 
\item{} What was the system temperature for both channel A \& B ? (This is
only used for the initial data headers, and will be replaced once the data is
calibrated.) 

\item{} Whether the receiver was set to linear or circular polarization

\item{} The total number of scans/cycle (i.e. its four if you took: ON, OFF, 
CAL ON, CAL OFF) 
\end{itemize}

Once you feel confident you know these things, login to an Arecibo computer 
and then go to whatever directory you wish your files to reside in. 
Type sdtest to run the SDFITS conversion program. The program will will ask a 
few questions and then start converting your data. 

{\it Note that you can also run this program "on-the-fly" to continuously update
a datafile with the online data via: sdtest -c}.

In that case, when you are finished you must quit out of the conversion 
program by typing \begin{verbatim}^C\end{verbatim} (possibly a few times). 

See http://www.naic.edu/\~astro/data/aips++/ for more complete details.

\subsection{SDFITS and AIPS++}

\subsection{SDFITS Example 1}

A data file with only two observations, each of a different galaxy (UGC 3489 
and UGC 4431).  Each cycle consisted of ON, OFF, CAL ON, CAL OFF 

\begin{verbatim}
sdtest
Enter Input (AO) filename: /proj/a9999/corfile.01jan00.a9999.1
Enter Output File: 01Jan00a.fits

Channel A system temp? :35

Channel B system temp? :35

Linear or Circular inputs? :C
Total number of scans in a cycle:4
Scan 930900005, UTC=2000-01-05T12:06:42, AST=2000-01-05T08:06:42
Object: U3489
nter Command (UPPER CASE) or Object Name: (cr will keep U3489
  QUIT to stop converting,
  DRIVE to finish entire file with same object name,
  UNTIL <num|scan>  to continue for num or until reaching scan,
  NEWCYCLE <num> to change the number of scans in a cycle to num,
  SKIP  [num | scan] to skip this scan, or num scans, or until scan
  ASKFREQ to prompt for frequencies each time,
  ASK to ask basic questions again.): U3489
Scan 930900009: U3489-1.  RA 11:12:24.40 Dec +31:24:41.0
Scan 930900010: U3489-2.  Offset 5400.000000 (5400.165333)
Scan 930900011: U3489-3.  Offset 5400.000000 (5399.659220)
Scan 930900012: U3489-4.  Offset 5400.000000 (5399.719945)
Object:U3489
Enter Command (UPPER CASE) or Object Name: (cr will keep U3489
  QUIT to stop converting,
  DRIVE to finish entire file with same object name,
  UNTIL <num|scan>  to continue for num or until reaching scan,
  NEWCYCLE <num> to change the number of scans in a cycle to num,
  SKIP  [num | scan] to skip this scan, or num scans, or until scan 
  ASKFREQ to prompt for frequencies each time,
  ASK to ask basic questions again.): U4431
Scan 930900009: U4131-1.  RA 07:59:11.43 Dec +31:48:23.6
Scan 930900010: U4131-2.  Offset 5400.000000 (5400.165333)
Scan 930900011: U4131-3.  Offset 5400.000000 (5399.659220)
Scan 930900012: U4131-4.  Offset 5400.000000 (5399.719945)
9632 records translated

\end{verbatim}

\subsection{SDFITS Example 2}

A data file with three observations, two of one galaxy (U5361) and one of 
another (U5592).  Each cycle consisted of ON, OFF, CAL ON, CAL OFF 

\begin{verbatim}
sdtest
Enter Input (AO) filename: /proj/a9999/corfile.01jan00.a9999.4 
Enter Output File: 01Jan00d.fits   

Channel A system temp? :9.2

Channel B system temp? :9.2

Linear or Circular inputs? :L
Total number of scans in a cycle:4
Scan 930900061, UTC=2000-01-05T15:30:13, AST=1999-11-05T11:30:13
Object: U5361
Enter Command (UPPER CASE) or Object Name: U5361
Enter Command (UPPER CASE) or Object Name: (cr will keep U5361
  QUIT to stop converting,
  DRIVE to finish entire file with same object name,
  UNTIL <num|scan>  to continue for num or until reaching scan,
  NEWCYCLE <num> to change the number of scans in a cycle to num,
  SKIP  [num | scan] to skip this scan, or num scans, or until scan 
  ASKFREQ to prompt for frequencies each time,
  ASK to ask basic questions again.): U5361
Scan 930900061: U5361-1.  RA 09:55:03.25 Dec +25:12:10.6
Scan 930900062: U5361-2.  Offset 8955.000000 (8948.313686)
Scan 930900063: U5361-3.  Offset 8955.000000 (8948.299047)
Scan 930900064: U5361-4.  Offset 8955.000000 (8948.302798)
Scan 930900065, UTC=2000-01-05T15:42:43, AST=1999-11-05T11:42:43
Object: U5361
Enter Command (UPPER CASE) or Object Name: (cr will keep U5361
  QUIT to stop converting,
  DRIVE to finish entire file with same object name,
  UNTIL <num|scan>  to continue for num or until reaching scan,
  NEWCYCLE <num> to change the number of scans in a cycle to num,
  SKIP  [num | scan] to skip this scan, or num scans, or until scan 
  ASKFREQ to prompt for frequencies each time,
  ASK to ask basic questions again.): 
Scan 930900065: U5361-1.  RA 09:55:03.25 Dec +25:12:10.2
Scan 930900066: U5361-2.  Offset 8955.000000 (8948.189632)
Scan 930900067: U5361-3.  Offset 5400.000000 (5399.836752)
Scan 930900068: U5361-4.  Offset 8955.000000 (8948.473693)
Scan 930900069, UTC=2000-01-05T15:57:01, AST=1999-11-05T11:57:01
Object: U5361
Enter Command (UPPER CASE) or Object Name: (cr will keep U5361
  QUIT to stop converting,
  DRIVE to finish entire file with same object name,
  UNTIL <num|scan>  to continue for num or until reaching scan,
  NEWCYCLE <num> to change the number of scans in a cycle to num,
  SKIP  [num | scan] to skip this scan, or num scans, or until scan 
  ASKFREQ to prompt for frequencies each time,
  ASK to ask basic questions again.): U5592
Scan 930900069: U5592-1.  RA 10:21:31.63 Dec +22:32:51.0
Scan 930900070: U5592-2.  Offset 5400.000000 (5401.432019)
Scan 930900071: U5592-3.  Offset 5400.000000 (5400.222629)
Scan 930900072: U5592-4.  Offset 5400.000000 (5400.127029)
14448 records translated
\end{verbatim}

\section{Available AIPS++ Routines}

The following provides details on the available routines to apply 
gain/temperature corrections for Arecibo data, as well as details of the
"ao.import" routine:

\input{README.latex}

\section{Sample Data Reduction Session}\label{sect:1}

This example assumes that an AIPS++ installation is available in directory
/pkg/aips++/stable and that the AIPS++ sample data is available in the
data repository at /pkg/aips++/data/demo/dishdemo/U3294.sdfit.
Modify the first few lines of this example appropriately if your installation
has a different path.

\begin{verbatim}
% source /pkg/aips++/stable/aipsinit.csh
% cp /pkg/aips++/data/demo/dishdemo/U3294.sdfit .
% aips++ -l naic.g
...
Time to initialize DISH =  11.978507
DISH tool is --> d
AO   tool is --> ao
Glish version 2.7. 
- ao.import('/pkg/aips++/data/demo/dishdemo/U3294.sdfit');
T
- #this automatically opens up the file
- ao.gaincorr('U3294_ms1');
\end{verbatim}

After the above commands, the calibrated data is placed in the dish results manager.
It can be plotted with the "ao.plot" command.

