\section{SDFITS}
\label{app:sdfits}

SDFITS is a convention for the use FITS binary tables to store 
single dish data.  This appendix will eventually be turned into a full 
fledged note.
At the present time, it is simply a summary of the various rules
regarding the interpretation of keywords and
columns which, taken together, constute the Single Dish
FITS convention.

A FITS binary table is also a SDFITS binary table if it follows
these rules.

\subsection{EXTNAME keyword}
\begin{verbatim}
EXTNAME = 'SINGLE DISH'
\end{verbatim}
The EXTNAME keyword must have a value of ``SINGLE DISH''.
A table having this value for EXTNAME is assumed to be fully compliant
with this convention.

\subsection{Virtual columns}
Most keywords should be thought of as
a column of the same name which has a constant value for all rows in 
the table.

There are a few exceptions to this rule.  FITS keywords which are
required to describe the structure of the table are not virtual
columns.  These include NAXIS, NAXIS1, NAXIS2, BITPIX, XTENSION, \linebreak[0]
EXTNAME, EXTVER, EXTLEVEL, PCOUNT, GOUNT, TFIELDS, TFORM, TTYPE, TUNIT, TNULL,
and TDISP.  The TDIMnnn keyword is the only table description keyword
which may be a column as described elsewhere in this appendix.

DATAMAX and DATAMIN are special keywords.  When used as keywords
it is most likely that the writer of that table meant them to
reflect the range of all of the data found in that table and
not to be used on a row-by-row sense.  If the writer feels a
need to supply row-by-row values for these two keywords, they should
explicitly make them true columns.

When creating a table, a column may be made virtual if it's
value can be represented as a single keyword (if it
is constant over all rows) and if that constant value is NOT
an IEEE special value (it has a standard ASCII 
representation) and the name of the column is 8 character or less.

If a column exists which has the same name as a keyword, the
column values should be assumed to take precedence (writers
should attempt to ensure that this confusing situation
does not happen).
		

\subsection{The DATA column and the DATA axes}
The DATA column containsthe primary data array.  It is required.
Each row of this column contains an n-dimensional array of data.  
The axis descriptions are handled in a similar way to FITS images.  

\begin{description}
\item[The TDIMnnn convention]This is how the shape of the DATA array should
be specified.
\begin{verbatim}
TDIMnnn = '(n1,n2,n3,n4...)'
\end{verbatim}
\begin{itemize}	
\item The product of all of the n1, n2, n3, etc.,  must be $\leq$ the width of
the column as given by the TFORMnnn keyword.
\item TDIMnnn may either be a keyword (constant shaped columns)
or it may itself be a column (variable shaped columns).
\item The data may be stored directly - unused elements in 
a particular row should contain the appropriate value
to indicate an undefined value as described in the
Binary Table Extension paper (i.e. NaN or use TNULL),
Or the DATA column may use the P type of storage 
(storage on ``the heap'').
\end{itemize}
This convention may be used with other columns which
do NOT use the same axis description (these will be
site specific columns and readers are free to ignore
these columns).
\item[Axis description]
This description applies to any column which has \\
\verb+TMATXnnn = T+.
It is assumed that \verb+TMATXnnn = T+ for the DATA column even when not
present.  The use of \verb+TMATXnnn = F+ for the DATA column is an error.

Axes descriptor keywords/columns include (not all need be present for each axis):
\begin{description}
\item[CTYPEn]The type of physical coordinate on axis n.
\item[CRVALn]The value of the physical coordinate on axis n at the 
reference pixel.
\item[CRPIXn]The array location of the reference pixel along axis
n.  CRPIX may be a fractional pixel and/or be outside
of the limits of the array.  This descriptor is optional for degenerate axes.
\item[CDELTn]The increment in physical coordinates along axis
n.  This descriptor is optional for degenerate axes.
\item[CROTAn]The axis rotation
\item[CUNITn]The unit of coordinate values along this axis.
\end{description}
Note that these are the original FITS axes description
keywords.  It is anticipated that when an agreement is
reached on the WCS convention that the appropriate keywords
specified in that convention will be used to describe
the axes here (e.g. CROTA is deprecated in favor of 
a more general mechanism).

Due to the virtual column convention, the following conflict may arrise:
\begin{verbatim}
TTYPE10 = 'CDELT4'
TFORM1  = '1E'
TUNIT10 = 'radian'
CUNIT4  = 'degree'
\end{verbatim}
Writers should avoid this conflict.  Readers encountering
such a conflict should report this as an error in the FITS
table.  However, readers may also choose to proceed by 
arbitrarily resolving the conflict so that some attempt
may be made within the anaysis package to deal with
this error.  It is recommended that the CUNITxxx specification
take precedence over any TUNITxxx specification.  In any
case, this conflict and the readers resolution of the
conflict (if at all) should be reported to the user.

The other standard FITS Image axis keywords may be used here.
This will include all of the WCS convention keywords
when they become accepted as part of the FITS standard.

Defined axis types (CTYPEnnn values) for this convention (others are allowed
but their values may be ignored by readers) are:

\begin{description}
\item[frequency-like]An 8 character string consisting
of the 4 character axis type plus 4 characters describing the reference frame.
An axis having this type is required.
\begin{description}
\item[axis types]
\begin{itemize}
\item 'FREQ' - frequency (Hz)
\item 'VELO' - velocity (m/s) (radio convention, unless 
overridden by use of the VELDEF SHARED keyword)
\item 'FELO' - regularly gridded in frequency but expressed 
as velocity in the optical convention (m/s)
\end{itemize}
\item[reference frames]
\begin{itemize}
\item '-LSR' : Local Standard of Rest
\item '-HEL' : heliocentric (barycentric)
\item '-OBS' : the frame of rest of the observer/telescope (topocentric)
\item 'LSRK' : LSR as a kinematical definition
\item '-GEO' : Geocentric
\item 'REST' : rest frequency
\item '-GAL' : Galactocentric
\end{itemize}
\end{description}
\item[Longitude-like] 'RA','GLON','ELON' plus an optional WCS 
projection code (degrees). This axis is required.
\item[Latitude-like] 'DEC','GLAT,'ELAT' plus an optional WCS 
projection code (degrees). This axis is required.
\item['TIME']The time since DATE-OBS (seconds) If the TIMESYS 
keyword is present, that keyword defines the time system for 
this table, including this column, otherwise UTC
is assumed.  This axis is optional.  If not present,
a value of 0.0 should be assumed.  This axis will
often be absent if DATE-OBS contains a time as well as a date.
\item['STOKES']The Stokes parameter of the data.
\begin{itemize}
\item $1,2,3,4$ $\Rightarrow$ I,Q,U,V
\item $-1,-2,-3,-4$ $\Rightarrow$ RR, LL, RL, LR
\item $-5,-6,-7,-8$ $\Rightarrow$ XX, YY, XY, YX
\end{itemize}
This axis is optional.  If not present, a value of 1 (Stokes I) should 
be assumed.
\item['BEAM']Beam ID.  This axis is optional.  If not present,
a value of 1 should be assumed.
\item['RECEIVER']receiver ID.  This axis is optional.  If not
present, a value of 1 should be assumed.
\end{description}
\end{description}

\subsection{CORE keywords and columns}
These must be provided in all SDFITS tables.  They are essential
and common to all observations and telescopes.  All single dish
FITS readers and writers must acknowledge (write and properly
interpret) all CORE keywords.

\begin{description}
\item[OBJECT]A string value giving an object name.
\item[TELESCOP]A string value giving the telescope name.
\item[BANDWID]The total bandwidth of the backend in units of Hertz.
\item[DATE-OBS]A string giving the observation date and optionally
the time at the start using the new FITS y2k 
convention.  The TIMESYS keyword may be used to
indicate the time system.  UTC is assumed if TIMESYS is absent.
\item[EXPOSURE]The effective integration time in seconds.
\item[TSYS]The system temperature in Kelvin.
\end{description}

\subsection{SHARED keywords and columns}
These have agreeded definitions and interpretions however their
presense is obtional.  These are largely common to all observations
and telescopes but not essential.  These may be ignored by a 
single dish FITS reader.
\begin{description}
\item[OBSERVER]A string giving the observer's name.
\item[OBSID]A string describing the observation.
\item[PROJID]A string describing the project.
\item[SCAN]A scan ID number.  Typically this is an identification number
given to a chunk of data when the data is taken.  Not all telescopes
provide a scan ID number.
\item[OBSMODE]The type of data and observing mode (8 characters 
total).  The type (LINE, CONT, PULS, etc) $+$ the 
mode (PSSW, FQSW, BMSQ, PLSQ, LDSW, TLPW, etc).
These rules do NOT define these observing modes.  Writers are strongly 
encouraged to use the FITS comments to document these modes.
\item[MOLECULE]A string used as a line identifier (with TRANSITI).
\item[TRANSITI]A string used as a line identifier (with MOLECULE.)
\item[TEMPSCAL]A string describing the scaling applied to reach 
the output intensity scale (``TB'',``TA'',``TA*'',``TR'',``TR*'').
\item[FRONTEND]A string giving the name of the front end device.
\item[BACKEND]A string giving the name of the back end device.
\item[TCAL]The calibration temp (K).
\item[THOT]The hot load temp (K).
\item[TCOLD]The cold load temp (K).
\item[TRX]The receiver temp (K).
\item[FREQRES]The frequency resolution in Hz.  This may differ 
from the channel spacing.
\item[TIMESYS]The time system which applies to all time columns
and keywords (see the y2k FITS DATE agreement).
\item[VELDEF]The velocity definition and frame (8 characters).
The first 4 characters describe the velicity definition.  Possible
definitions include:
\begin{description}
\item[RADI]radio
\item[OPTI]optical
\item[RELA]relativistic
\end{description}
The second 4 characters describe the reference frame (e.g. ``-LSR'',
``-HEL'', ``-OBS'').  If the frequency-like axis gives a frame,
then the frame in VELDEF only applies to any velocities given
as columns or keywords (virtual columns).
 \item[VFRAME]The radial velocity of the reference frame wrt the
observer.  $V_frame - V_telescope$.
\item[RVSYS]The radial velocity, $V_source - V_telescope$.
\item[OBSFREQ]The observed frequency (Hz) at the reference pixel 
of the frequency-like axis.
\item[IMAGFREQ]The image sideband freq (Hz) corresp. to OBSFREQ.
\item[LST]The LST (seconds) at the start of scan.
\item[AZIMUTH]The azimuth at TIME (deg) (if the TIME axis is
non-degenerate, then this is the azimuth at the TIME
of the first pixel on the TIME axis.
\item[ELEVATIO]The elevation at TIME (deg) (same caveat as for AZIMUTH)
\item[TAU]The opacity at OBSFREQ.
\item[TAUIMAGE]The opacity at IMAGFREQ.
\item[TAUZENIT]The opacity per unit air mass.
\item[HUMIDITY]The relative humidity (fraction, 0..1).
\item[TAMBIENT]The ambient temp (K).
\item[PRESSURE]The atmospheric pressure (mm Hg).
\item[DEWPOINT]The dew point (K).
\item[WINDSPEE]The wind speed (m/s).
\item[WINDDIRE]The wind direction (deg. west of north).
\item[BEAMEFF]The main-beam efficiency.
\item[APEREFF]The antenna aperture efficiency.
\item[ETAL]The rear spillover and scattering efficiency.
\item[ETAFSS]The forward spillover and scattering efficiency.
\item[ANTGAIN]K per Jy.
\item[BMAJ]The major main-beam FWHM (deg).
\item[BMIN]The minor main-beam FWHM (deg).
\item[BPA]The beam position angle (degrees east of north).
\item[SITELONG]The site longitude (deg).
\item[SITELAT]The site latitude (deg).
\item[SITEELEV]The site elevation (m).
\item[RESTFREQ]The rest frequency (Hz).
\end{description}

\subsection{Other columns}
Any additional columns not explicitly mentioned here
can be added to an SDFITS table, although, obviously,
most readers will not be able to propertly interpret
those columns.

\subsection{Multiple SDFITS tables in a single file}
Any number of FITS binary tables can be attached to
the same FITS table. This convention doesn't have anything
to say about how the tables in such a file might be
related.  This strategy, attaching multiple SDFITS binary
tables to a single FITS file, is one possible strategy
for storing variable length DATA.  DATA with similar
sizes could be stored in separate tables, minimizing the
amount of padding required in any one table without
resorting to using the table heap convention to store
variable length arrays.  SDFITS readers should be
capable of appending the contents of several SDFITS
tables to a single result.

