\section{Basic Calibration Fundamentals}
\label{GRsyncalbasics}

Calibration and imaging responsibilities are divided between the
\calibrater\ and \imager\ tools. The \calibrater\ tool handles
visibility-plane calibration while the \imager\ tool deals with image
formation, deconvolution and image-plane calibration.

In a typical synthesis observation, a number of calibrators of known
structure (usually point sources, and at least one of known flux
density) are observed in addition to the scientific target(s) in order
to obtain a measure of the systematic properties of the instrument
and, if relevant, the corrupting media above it.  The effects of these
systematic properties are then removed from the data for the
scientific target.  This technique is often referred to as {\em
cross-calibration} and its success depends implicitly on the
relationship between: 1) the true brightness distribution of the
observed radio sources (what we are after); 2) the formal response of
the instrument and its variety of imperfections of varying importance
(what we are calibrating); 3) and the data that is actually recorded
(what we have measured).

The \aipspp\ Measurement Equation, originally defined by Hamaker,
Bregman, \& Sault (1996a, 1996b), is an explicit description of this
relationship.  The Measurement Equation (in the visibility plane) for
a spectral line polarization calibration is given by

\begin{equation}
\vec{V}_{ij}~=~B_{ij}~G_{ij}~D_{ij}~P_{ij}~T_{ij}~\vec{V}_{ij}^{\mathrm{~IDEAL}}
\end{equation}

\noindent
where:
\begin{description} 
\item $\vec{V}_{ij}~=~$ cross-correlations between two polarizations
  for each of two feeds ($i,j$) which characterize an individual
  baseline (e.g. the measured visibilities). 
\item $\vec{V}_{ij}^{\mathrm{~IDEAL}}~=~$ visibilities measured by an ideal 
  interferometer (e.g. no instrumental errors or calibration effects). 
  $\vec{V}_{ij}^{\mathrm{~IDEAL}}~=~$ is directly related to the Fourier 
  Transform of the true polarized sky brightness, $\vec{I} = (I,Q,U,V)$.
\item $T_{ij}~=~$ Complex gain effects which are polarization-independent, 
  e.g. tropospheric effects, high-frequency opacity corrections, 
  antenna gain as a function of elevation, baseline corrections.
\item $P_{ij}~=~$ Parallactic angle. 
\item $D_{ij}~=~$ Instrumental polarization response.  "D-terms" 
  describe the polarization leakage between feeds (e.g. how much the 
  R-polarized feed picked up L-polarized emission, and vice versa).  
\item $G_{ij}~=~$ Electronic gain response due to components in the signal  
  path between the feed and the correlator.  This complex gain term 
  $G_{ij}$ includes the scale factor for absolute flux density calibration,
  and may include phase and amplitude corrections due to changes 
  in the atmosphere (in leiu of $T_{ij}$).
\item $B_{ij}~=~$ Bandpass response. 
\end{description}

Note that the Measurement Equation is a matrix equation.
$\vec{V}_{ij}$ and $\vec{V}_{ij}^{\mathrm{~IDEAL}}$ are 4-vectors with
elements for each correlator product, and the $T_{ij}$, etc., are
4$\times$4 matrices that `corrupt' the correlator products according
to the properties of the specific term.  The order of the corrupting
terms from right-to-left, for the most part, is the order in which the
errors affect the incoming wavefront.  \aipspp\ hardwires the
Measurement Equation order and all of the specific algebra associated
with the individual terms.  As such, it is usually possible to ignore
the fact that the Measurement Equation is a matrix equation, and treat
the different terms as labeled `components' (black boxes) along the
signal path.

In practical calibration, it is often best to begin the calibration
process by determining solutions for those terms which affect the data
most.  Thus, the user would normally determine gain solutions ($G$)
first (applying pre-computed parallactic angle corrections $P$ if doing
polarization calibration), then bandpass solutions ($B$) and/or
polarization solutions ($D$).  Errors which are
polarization-independent ($T$) can be determined at any point in the
calibration process.  $T$ has nearly identical form to $G$ but it
located at a different point in the Measurement Equation.  A practical
advantage to this factorization is the ability to store
polarization-independent effects which are characterized by a certain
timescale or a particular parametrization (e.g., tropospheric opacity)
as a separate calibration factor, thus isolating them from electronic
gain effects which may vary on a different timescale and/or with
different systematics.  Although most data reduction sequences will
probably proceed in this order, the user has complete freedom and
flexibility to determine the order in which calibration effects are
solved for.  Self-calibration, the process by which an improved model
of the target source is used to obtain improved solutions for the
calibration, is also not limited to just the electronic gain ($G$)
corrections, as has been traditionally the case.  Improved solutions
for, say, the instrumental polarization ($D$) may be obtained by
self-calibration as well.  In general, decision-making along the
calibration path is a process of determining which effects are 
dominating the errors in the data and deciding how best to converge
upon the optimal set of calibration solutions (for all relevant components)
and source models which best describe the observed data.   

Once the calibration solutions have been determined, the \aipspp\
software applies the solutions according to the order defined by
the Measurement Equation.  The individual calibration components
represented in the Measurement Equation are described in detail in the
\htmlref{Synthesis Calibration}{GRcalibration} chapter of {\tt Getting
Results}.  The generalization of the Measurement Equation to image
plane effects is also described there.

% For standard continuum and spectral line observations (e.g. the VLA or
% ATCA), the unknown properties of the instrument can usually be treated
% entirely in the visibility plane in the \calibrater\ tool.  Knowledge of
% the observed data and brightness distribution models for the
% calibrators enable solutions for the gains, which may be applied to
% the data for the scientific target.  The corrected data for the
% scientific target is then Fourier transformed and deconvolved to
% obtain a brightness distribution model.  The method of {\em
% self-calibration} extends this process to iterative improvement of the
% gains and brightness distribution model of a single source.

The observed data, $\vec{V}_{ij}$, are stored in the DATA column in
the MAIN table of the MS; these are the raw data as loaded by the
filler tool or imported from a UVFITS file.  Associated with the DATA
column are related columns to hold the most recent version of the
calibrated data (CORRECTED\_DATA), and the latest visibility plane
representation of the source or field model, $\vec{V}_{ij}^{IDEAL}$,
(MODEL\_DATA).  The latter two columns are filled in by the
\calibrater\ and \imager tools,
respectively.  The actual calibration information is stored in
separate calibration tables.  The observed DATA column does not
change during reduction, but the related columns do.  When plotting
the data using a tool such as \msplot, the data
sub-types ({\tt observed, corrected, and model}) can be selected
individually.

