\documentclass[11pt]{article}
\usepackage{html}
\include{ahrecipe}
\begin{document}
\begin{ahrecipe}

\begin{rawhtml}
<body bgcolor="#ffffff">
\end{rawhtml}

%\argoals
\begin{rawhtml}
<font face=helvetica size="+2" color="#cc0000">
Goals:</font>
\end{rawhtml}
{Make a multi-field clean of a confused field}

%\arusing
\begin{rawhtml}
<font face=helvetica size="+2" color="#cc0000">
Using:</font>
\end{rawhtml}
{imager}

%\arassume
\begin{rawhtml}
<font face=helvetica size="+2" color="#cc0000">
Assume:</font>
\end{rawhtml}
{You have a VLA observation of a weak source with a
confusing source nearby. You wish to remove the effect of the
confusing source and make a clean image of the weak source.
You can find the location of the confusing
source from a lower-resolution image.}

%\arscript
\begin{rawhtml}
<font face=helvetica size="+2" color="#cc0000">
Script:</font>
\end{rawhtml}
{\htmladdnormallink{multifield.g}{../multifield.g.txt}}

\begin{arexample}
\arline{include~'imager.g';
include~'general.g';}{Include the imager and general scripts.}
\arline{dowait:=T}{Tell the object system to wait for a function to finish}
\arline{if(!tableexists('uvdata.ms')) \{ \\
  m:=fitstoms('uvdata.ms', 'UVDATA.TB0');m.summary(verbose=T);m.close(); \\
\} }{If the MS doesn't exist, make it from the UVFITS file} 

\arline{imgr:=imager('uvdata.ms')}{Make the imager}
\arline{if(!tableexists('clean.restored')) \{ \\
  imgr.setimage(cellx='1.5arcsec', celly='1.5arcsec', nx=1024, ny=1024, stokes='I') \\
  imgr.weight('robust') \\
  imgr.uvrange(0, 60000); \\
  imgr.clean(model='clean') \\
  imgr.restore(model='clean')\\
\} }{Clean a low-resolution image from the inner third of the uv plane}

\arline{im:=image('clean.restored'); \\
center:=im.coordmeasures().direction \\
other :=im.coordmeasures([591,303]).direction \\
im.done() }{Find the measure for the direction of the phasecenter and of another bright source}

\arline{
imgr.setimage(cellx='0.15arcsec', celly='0.15arcsec', nx=256, ny=256, stokes='I',
              phasecenter=center);\\
imgr.make('center.clean')\\
imgr.weight('robust')\\
imgr.setimage(cellx='0.15arcsec', celly='0.15arcsec', nx=256, ny=256, stokes='I',\\
              phasecenter=other);\\
imgr.make('other.clean')
}{Make empty full resolution images at the two positions}
\arline{
imgr.clean(algorithm='mf', model=['center.clean', 'other.clean'])\\
imgr.restore(model=['center.clean', 'other.clean'])
}{Clean and restore the two images jointly using the mf algorithm.}
\arline{imgr.close()}{Close the imager}
\arline{imgr.done()}{Delete the imager object}
\end{arexample}

%\arconclusion
\begin{rawhtml}
<font face=helvetica size="+2" color="#cc0000">
Conclusion:</font>
\end{rawhtml}
{Note the use of the coordmeasures function in the
image object. This return both the value of the specified pixel
and the position in measures format which can be used elsewhere.

Once we have two images, we have to use the multi-field algorithm. The
multifield clean algorithm works as follows:

\begin{enumerate}
\item A first estimate image (usually zero) is made for each field.
\item For the current models, the predicted visibilities 
are calculated by applying the appropriate primary beam and then
Fourier transforming and de-gridding.
\item The residual visibilities are calculated and then
gridded and Fourier transformed to obtain the residual
image for each pointing.
\item Each residual image is then cleaned using the appropriate PSF down
to some level using the Clark Clean approach.
\item The incremental clean components are added to the current model
and the process is iterated from step 2 onwards.
\item The procedure stops at some threshold in residual.
\end{enumerate}

}
\end{ahrecipe}

\end{document}
